### 2024-09-20 manuscript/manuscript.Rmd
Update all datasets

---
title: "Flattering Advice: Avoiding Disappointment as a Driver of Gender Discrimination"
shorttitle        : "Flattering Advice"
author: 
  - name          : "Amanda Chen"
    affiliation   : "1"
    corresponding : yes    
    email         : "zchengj@connect.ust.hk"
    address       : "Department of Management, The Hong Kong University of Science and Technology, Hong Kong"
  - name          : "David Hagmann"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "The Hong Kong University of Science and Technology"
note : |
  `r format(Sys.time(), "%e %B, %Y")`
abstract: |
  The purpose of advice is to help recipients make better decisions and solve their problems. However, in this paper, we propose that advisors may also take recipients' responses into account and attempt to avoid disappointing them, leading to what we call "flattering advice." In two pre-registered experiments involving real interactions between advisors and advisees (n = 2,700), we show that individuals consider others' expectations when giving advice, even when such expectations do not provide informative insights for the underlying decision-making problem. As a result, men receive more aspirational advice than women, given their higher confidence, when expectations (but not gender) are shown to advisors. However, incorporating expectations leads to worse outcomes (Study 1). Individuals inflate their advice due to interpersonal considerations: people who are concerned about advisees' likability inflate their advice to a great extent. Advisees reward such flattering advice, finding the advisers more likable and no less trustworthy than those who provide more accurate advice (Study 2). We discuss practical and theoretical contributions and future research directions.
keywords          : "Gender difference, Advice, Interpresonal relationship"
wordcount         : "4,362"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa7"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
classoption       : "man"
header-includes:
   - \usepackage{setspace, appendix, placeins, booktabs, tikz, siunitx}
   - \usepackage{tabularx, epigraph, float, colortbl, tabu, pdflscape}
   - \usetikzlibrary{positioning}
   - \raggedbottom
   - \usepackage[maxfloats=256]{morefloats}
   - \maxdeadcycles=1000
bibliography      : bibliography.bib
output: 
  # papaja::apa6_docx
  papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
always_allow_html: true
---

```{r setup, include=FALSE}

library(magrittr)
library(ltm)

knitr::opts_chunk$set(fig.width = 8, fig.height = 4, cache = F, echo = F,
                      fig.align = 'center',warning = FALSE, message = FALSE)

study3_math50 <- read.csv(here::here("Data/Cleaned Data/S3_Math50_cleaned.csv"))
study3_time1All <- readRDS(here::here("Data/Study3_time1All.Rds"))
study3_time1 <- readRDS(here::here("Data/Study3_time1.Rds"))
study3_time2_giver <- readRDS(here::here("Data/Study3_time2_giver.Rds"))
study3_time2_giver50 <- readRDS(here::here("Data/Study3_time2_giver50.Rds"))
study3_advice <- readRDS(here::here("Data/Study3_time2_advice.Rds"))
study3_time3 <- readRDS(here::here("Data/Study3_time3.Rds"))

```

# Introduction

Seeking and giving advice can present conversational challenges that are
difficult to navigate. Advice seekers may be embarrassed to ask for help
[@Benabou2022], and they may fail to do so if the information they
receive could lead to disappointment [@Gill2012]. Indeed, people prefer
advisers who withhold unpleasant information [@Shalvi2019] and punish
those who tell them things they do not want to hear [@John2019]. People
appear to engage in strategic selection of advice to avoid unpleasant
information and protect the utility they get from positive beliefs about
themselves and the future they anticipate [@Loewenstein2018;
@Loewenstein1987]. But do advice givers anticipate and consider the
hedonic cost of honest information to the recipient? While previous work
has emphasized the instrumental value of advice (e.g., satisfaction with
a decision or accuracy of a prediction), we focus on the motivations of
the person providing advice.

Research on advice has largely focused on settings that do not tackle
belief utility. For example, participants may have been asked to
estimate the number of balls in an urn or choose from pairs of lotteries
[@Benjamin2015]. Advisers have been shown to be motivated to give
accurate advice [@Jonas2003], and engage in perspective-taking to meet
the preferences of the advisee [@Kray2000; @Kray1999]. When they are
uncertain, they even prefer that advisees do not rely (extensively) on
their recommendations [@Ache2020]. Thus, it seems that people are
concerned about providing good advice even when they are not
incentivized for the recipient's decision quality.

In many real settings, however, advice incorporates beliefs about the
recipient. Someone recommending an expensive restaurant conveys that
they think the advisee is wealthy enough to afford an expensive meal,
and someone recommending a student pursue a PhD in economics assumes
that they are intellectually capable. Conversely, telling a poorly
performing student to switch to a less-demanding major conveys a belief
that the student is not capable of improving. Thus, while the painful
truth may prevent someone from making a bad decision, it may come at a
hedonic cost to the recipient and damage the relationship between the
adviser and advisee. We propose that advisers take these costs into
account, even when the interactions are entirely anonymous and they are
incentivized for the quality of the advisee's decision.

Specifically, we focus on the signaling value of advice: what does the
advice imply about the adviser's perception of the advisee? We propose
that advisees have expectations about the advice they receive and treat
this as a reference point [@Koszegi2006]. Advice that falls short of
these expectations creates disappointment, while advice that exceeds
them is flattering and creates positive interpersonal benefits because
it creates a perception that the adviser holds the advisee in high
esteem. Such interpersonal considerations may be as important, if not
more important, to the adviser than the instrumental value of the
advice: the adviser directly experiences a souring relationship, but may
suffer little and only much delayed consequences if an advisee makes a
poor life decision.

Our account therefore predicts that people give and receive advice that
is more flattering and suggestive of inflated ability and optimistic
outcomes than would advice in the absence of such interpersonal
considerations. However, we propose that this is not true for all
advisees equally. Presenting flattering advice requires the formation of
beliefs about the advisee's expectations (reference point). @Exley2022
show that men are more overconfident than women, conditional on equal
performance, and that people anticipate this difference. An adviser who
seeks to avoid disappointment and incorporates expectations into their
advice would then present more favorable advice to male advisees than to
female advisees (see Figure 1 for an illustration). This provides a
novel account for a well-known finding that men receive more
aspirational advice than do women [@Kanze2018]. Notably, when it pays to
have accurate beliefs, such flattering advice would come at a cost:
while men would be more likely to aim higher, they would also suffer
greater costs from aiming too high and failing to meet the objective.

In a large, preregistered experiment (n = 2,000), we show that advisers
consider the expectations of advisees, even when such expectations do
not provide informative insights for the underlying decision-making
problem. We find that as a result, showing expectations (but not gender)
to advisers leads men to receive more aspirational advice than women.
Incorporating expectations leads to worse advice, even when advisers are
incentivized based on the outcome of the advisee's decision. In a second
preregistered experiment (n = 700), we show that advisers respond to
incentives for interpersonal considerations. They tell people that they
are more attractive than they believe them to be, and inflate to a
greater extent when incentivized for likability rather than accuracy.
Notably, they inflate even when only incentivized for accuracy. Advisees
indeed reward such flattering advice, rating the advisers more likeable
and no less trustworthy than those who provide more accurate advice.

# Open Science Statement

We report all sample sizes, data exclusions, manipulations, and measures
in the studies. Screen captures of the experimental materials are
available in the Supplemental Information. The complete data, code to
reproduce all statistical analyses and figures in the manuscript, as
well as the preregistration reports will be available via OSF. All our
studies were preregistered on AsPredicted.org.

<!-- ^[<https://osf.io/8r3d4/?view_only=5ad7bafcd16b4d4ba08bb28b0e2bd02d>] -->

<!-- Commenting this out for IACM -->

# Study 1

We begin by examining whether advisers take the expectations of advisees
into account, even when those expectations are not relevant to the
decision the advisee has to make. In a three-stage experiment, we first
recruit a sample of advisees and ask them to complete a mathematics
quiz. After answering ten multiple-choice questions, participants are
anchored to low or high performance expectations and asked to guess how
many questions they answered correctly. Next, we recruit advisers and
show them either only the true performance of the advisee, or the true
performance and how many questions they guessed they answered correctly.
Advisers are tasked with recommending whether the advisee should compete
against a group of high-performers or a group of low-performers on the
mathematics quiz, where a potential bonus depends on outperforming the
competitor and which group was chosen. Finally, we invite the advisees
back and show them the advice they have received. Advisees then pick
whether to compete against high or low performers, with the outcome
determined by their past performance on the quiz. Our key hypotheses are
that (1) advisees who were anchored to high expectations would be more
likely to receive advice to compete against the high performer group,
and (2) that men are more likely than women to be advised to compete
against the high-performer group when expectations are shown.

## Methods

```{r}

subgroup_averages <- data.frame(subgroup = numeric(), avg_score = numeric())
set.seed(807)
# loop through 1000 iterations of randomly selecting 10 participants and computing their average score
for (i in 1:1000) {
  # randomly select 10 participants
  subgroup <- sample(study3_math50$ProlificID, size = 5, replace = FALSE)
  # compute the average score for the selected participants
  subgroup_avg <- mean(study3_math50$Score[study3_math50$ProlificID %in% subgroup])
  # add the subgroup and its average score to the dataframe
  subgroup_averages <- rbind(subgroup_averages, data.frame(subgroup = i, avg_score = subgroup_avg))
}

p5 <- quantile(subgroup_averages$avg_score, probs = 0.05)
p95 <- quantile(subgroup_averages$avg_score, probs = 0.95)

sortedScore <- sort(study3_math50$Score, decreasing = TRUE)
LP_lower <- sortedScore[50]
LP_upper <- sortedScore[31]
HP_lower <- sortedScore[20]
HP_upper <- sortedScore[1]

```

As part of our main experiment, participants choose to compete against a
group of previous participants and have their expectations anchored
based on their performance. Thus, we begin by first recruiting a sample
of `r scales::comma(dim(study3_math50)[1])` participants from Prolific
to complete a 10-question multiple choice mathematics quiz. The
questions were taken from a paper-version of the ASVAB standardized
exam, such that answers were not available online. Participants had five
minutes to answer the quiz and were paid 10 cents for each correctly
answered question. We define the top 20 scorers as the "High Performers"
and the bottom 20 scorers as the "Low Performers." On average,
participants answered `r round(mean(study3_math50$Score),2)` questions
correctly, High Performers scored between `r HP_lower` and `r HP_upper`,
and Low Performers scored between `r LP_lower` and `r LP_upper`. In
order to anchor the expectations of participants in our main experiment,
we simulated 1,000 pairings of groups of five participants, with the
5<sup>th</sup> percentile of groups scoring an average of `r p5` and the
95 th percentile scoring an average of `r p95`. We report these averages
to participants in the Low Expectations and High Expectations treatment,
respectively.

We then recruited `r scales::comma(dim(study3_time1All)[1])`
participants for the role of advisees in Stage 1 of our main experiment.
To arrive at a gender-balanced sample, we dropped the last two male
participants to complete the survey, ending up with a sample of 500 men
and 500 women ($M_{\text{Age}}$ = `r mean(study3_time1$Age)`).
Participants completed the same 10-item mathematics quiz as the earlier
participants and were informed that their performance would affect their
bonus earnings in a follow-up stage to be conducted a few days later.
After completing the quiz, we informed them of the average score of a
group of five participants from the preliminary survey. We randomly
assigned them to learn about the 5<sup>th</sup> percentile of groups,
which scored `r p5`. ("Low Expectations" treatment) or the
95<sup>th</sup> percentile of groups, which scored `r p95` ("High
Expectations" treatment). Participants then made a guess
(unincentivized) about how many questions they think they answered
correctly. The survey concluded with basic demographic questions.

We then recruited `r scales::comma(dim(study3_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We began
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 had completed, and informed them of the
average score of all participants in the preliminary study. Advisers had
to recommend whether an advisee should compete against the Low
Performers or the High Performers (we used these terms in the survey).
We anticipated that being told to compete against High Performers is
more flattering and hence being told to compete against the Low
Performers would be disappointing if one had expected to do well.
Advisees would earn 30 cents if their score was equal to or higher than
a randomly selected member of the Low Performer group and they chose to
compete against this group, or 50 cents if they performed as well or
better than a randomly selected member of the High Performer group and
they chose to compete against that group. If they scored lower, they
would not receive any bonus.

Advisers were randomly assigned to one of two treatments. In the
"Baseline" treatment, they only observed the score of the advisee on the
mathematics quiz. In the "Expectation" treatment, they observed the
score as well as the advisee's guess for how many questions they
answered correctly. Note that since the outcome is determined only by
the past score on the quiz, the advisee's guess is immaterial to which
group they should compete against. Moreover, in neither treatment did
they receive any demographic information about their advisee.
Participants gave recommendations to 10 advisees, which unbeknownst to
them were five men and five women matched to have identical performance
on the test.[^1] They were informed that if their advice was shown to a
participant who returned for the follow-up survey, they would receive
the identical bonus as that participant. The survey then concluded with
basic demographic questions.

[^1]: We made this decision to account for the possibility of gender
    differences in performance.

Finally, we invited participants from Stage 1 back for the follow-up
survey. Following our preregistration, we kept the survey open for 7
days. In total, `r scales::comma(dim(study3_time3)[1])` participants
(`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Female',])[1])`
women) returned. The brief survey reminded them of the task they
completed in Stage 1, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally were reminded them of how
many questions they guessed they had answered correctly. Importantly,
they were not informed of their true score or the score of the groups
they could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

## Results

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants completed the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study3_time1[study3_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study3_time1, var.equal = T))$statistic`).
The expectations treatment did, however, affect how well they thought
they performed. Participants in the "Low Expectations" treatment guessed
a score of
`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study3_time1[study3_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study3_time1, var.equal = T))$statistic`),
showing that the manipulation was successful, albeit small. Contrary to
our expectations, we did find a gender difference in performance: men
scored
`r round(mean(study3_time1[study3_time1$Gender == "Male",]$count),2)` on
average, while women scored
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study3_time1, var.equal = T))$statistic`).
Consistent with this difference, men thought they answered more
questions correctly than did women
(`r round(mean(study3_time1[study3_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study3_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We
observe that women underestimate their performance by 0.76 points, while
men do so by only 0.27 points
`r papaja::apa_print(t.test(overconfidence ~ Gender, data = study3_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`.
Although neither gender is overconfident, men on average are more
confident in their performance than are women, as we had expected.

```{r study1regs}
study3_advice <- study3_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study3_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study3_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study3_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor.")

```

```{r study1genderdiff, fig.cap ="Advisers observed real performance and the participants' estimated performance, but not their gender. As a result of their higher expectations, men (blue) were advised to compete against the high performance group more often than women (red). The grey line shows advice absent expectations, which did not differ by gender."}
library(dplyr)
library(ggplot2)
library(magrittr)

ProbHG <- study3_advice %>%
  group_by(Condition, Performance, Gender) %>%
  dplyr::summarize(avg_hardGroup = mean(Advice))

df_exp <- subset(ProbHG, Condition == "E")
df_none <- subset(ProbHG, Condition == "None")

ggplot(df_exp, aes(x = Performance, y = avg_hardGroup, color = Gender)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),se = FALSE) +
  geom_smooth(data = df_none, aes(x = Performance, y = avg_hardGroup),
              method = "glm", method.args = list(family = "binomial"), color = "grey", se = FALSE) +
  scale_color_manual(values = c("#2A629A",'#FF7F3E', "grey"),
                     labels = c("Female + Expectations", "Male + Expectations", "Expectation Hidden")) +
  scale_y_continuous(name = "P(Compete Against High Performance)",
                     labels = scales::percent,
                     expand = c(0, 0)) +
  scale_x_continuous(name = "Performance",
                     expand = c(0, 0),
                     breaks = seq(0, 10, 2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20))
```

Next, we examine whether advisers took the advisee's expectations into
account. Column 1 of Table \@ref(tab:study1regs) reports a linear
probability model on advising to compete against the High Performer
group for advisers in the "Expectations" treatment, controlling for the
true performance of the advisee. Because each adviser made ten
recommendations, we cluster standard errors at the adviser level.
However, contrary to our expectations, we do not see a significant
effect of the expectations treatment. This may be because the induced
difference in expectations was too small.

However, recall that men were more confident in their performance than
were women. Our theory thus predicts that showing expectations should
lead to more flattering advice (i.e., a recommendation to compete
against the High Performer group) for men than for women. We report a
linear probability model with advice to compete against the high group
as the outcome measure, and the advisee's gender, whether expectations
were shown to the adviser, and the interaction of the two in Column 2 of
Table \@ref(tab:study1regs). As predicted, we find a significant
interaction effect: men are more likely to be advised to compete against
the High Performers when expectations are shown than when they are
hidden ($p < 0.001$). We show this result graphically in Figure
\@ref(fig:study1genderdiff)).

<!-- Amanda: The gender discrepancy in advice is more significant when performance is less decisive. When performance is extremely low, there is no chance to win high performers and only a slim chance when competing with low performers. When performance is extremely high, it makes no sense to forgo the high chance to win more and choose a low bonus. However, in the middle range of performance, either option has pros and cons, and neither is absolutely better. In such a situation, advisers would take participants' expectations into account, regarding them as a critical factor. -->

<!-- David: If we plotted on the x-axis performance and on the y-axis the difference -->

<!-- between recommendations for men & women, this would look inverse-U shaped, right? -->

<!-- Might be a nice way to illustrate what you describe. -->

<!-- Amanda: I think our primary purpose here is still showing the existence of gender difference -->

<!-- The current plot shows discrepancy between two lines, allowing us to demonstrate gender gap -->

To examine the quality of advice, we computed the expected bonus
earnings for someone who followed the recommendations. For example, if
an adviser suggested competing against the High Performer group, we
matched the advisee against all 20 members of that group and determined
how often their score matched or exceeded that of the member. We then
multiplied this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of Table
\@ref(tab:study1regs). Displaying expectations led men to be advised to
compete against the High Performer group more often, and this advise
turned out to be bad: men receive worse advise than do women when
expectations are displayed, but not in their absence.[^2]

[^2]: This analysis was not preregistered, and we note here that the
    reduction in expected earnings is small. However, it is interesting
    that expectations have a negative effect for men who underestimate
    their performance on average. One possibility is that advisers
    suggest the High Performer group more often than is optimal. We
    return to the possibility that advice is overall biased to be
    flattering in Study 2.

```{r study3time3}

study3_time3 <- study3_time3 %>% dplyr::mutate(AdviseeCondition = factor(AdviseeCondition, levels = c('Low', 'High'))) %>%
     dplyr::mutate(Adopt = ifelse(Bet == Advice,1,0))

modelsummary::modelsummary(list(`(1)` = lm(bonus ~ AdviseeCondition + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(2)` = lm(bonus ~ Gender + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(3)` = lm(Adopt ~ Gender + GiverCondition + Advice, data = study3_time3),
                                `(4)` = lm(Adopt ~ Gender + GiverCondition + Advice + Gender*GiverCondition + GiverCondition*Advice + Gender*Advice, data = study3_time3)),
                           vcov = c("classical", "classical","classical","classical"),
                            coef_map = c("AdviseeScore" = "Performance",
                                         "AdviseeConditionHigh" = "High Expectation",
                                        "GenderMale" = "Advisee Male",
                                        "GenderMale:GiverCondition" = "Expectation x Male",
                                        "GiverCondition" = "Expectation Shown",
                                        "Advice1" = "Advice: High Performer",
                                        "GenderMale:Advice1" = "Advice: High Performer x Male",
                                        "GiverCondition:Advice1" = "Expectation x Advice: High Performer",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the actual bonus based on the performance and whether the participant is primed with high expectations. Column 2 displays the actual bonus based on the performance and gender of the participant. Columns 3 and 4 display the chance of adopting the advice based on the gender of the participants, whether the advisee sees expectations, and whether the advice is to compete against high performers. The former considers only the main effect, while the latter also includes the interactive effect.")

```

To determine whether flattering advice is truly costly, however, we need
to examine the outcome of the advisees. In particular, they could ignore
flattering advice, recognizing it as such and thus failing to adhere to
it. Align with our prediction, participants who were primed with high
expectations earned less when their expectations were conveyed to the
advisers, although this result is only directional (see Column 1 of
Table \@ref(tab:study3time3)). Similarly, as shown in Column 2 of Table
\@ref(tab:study3time3), we also found that male participants received
less reward when their expectations were presented, even when their
performance was the same. These findings suggest that flattering advice
is not without consequences. Furthermore, our research indicates that
because males are more likely to follow such advice (see Column 4 of
Table \@ref(tab:study3time3), their tendency towards competition
intensifies the cost of such advice.

## Discussion

When the quality of a decision is not dependent on someone's
expectations, advice should not be moved by whether an adviser is aware
of the advisees expectations. However, consistent with our account that
advisers consider the belief utility of the advisee and want to flatter,
rather than disappoint them, we find that displaying expectations to
advisers does matter. Specifically, we found that men underestimate
their score on a mathematics test less than do women. When these
expectations are displayed to advisers, they are more likely to tell men
to compete against the High Performer group. Notably, this is bad
advice: men whose advisers observed their expectations got worse advice
as a result. Our findings suggest that males with the same scores as
their female counterparts earned less, although the evidence is only
directional. This may be due to males receiving more favorable advice
and being more likely to adopt it, resulting in even worse outcomes.

<!-- David: One more sentence about the earnings difference here, which does not -->

<!-- differ significantly. And one sentence about the new analysis of whether -->

<!-- people follow the advice or not. -->

# Study 2

<!-- David: Just commenting on what I tried to do here... a bit of a transition -->

<!-- for why this study follows from the last. -->

So far, we have shown that displaying expectations can induce gender
differences in advice-giving. Specifically, advisers were more likely to
recommend competing against a group of High Performers to men who had
higher expectations about their performance than did women. We now test
directly whether this is driven by interpersonal concerns. Specifically,
we examine whether more flattering advice is the result of a desire to
be liked and whether advisees indeed do like advisers more when they
give flattering advice.

We conducted a three-stage experiment in which we experimentally
manipulated the incentives for advicers. We began by inviting a group of
advisees to upload photos of themselves ("selfies") and informed them
that they would be rated on their attractiveness. We grouped them with
nine other participants of the same gender and recruited participants of
the opposing gender to rank them from most to least attractive and to
provide advice. Specifically, we asked them to advise the participant
they ranked as the 7th most attractive (4th least attractive) on what
rank they should bet they were ranked by a larger group of raters.
Advisers were randomly assigned to two treatments, receiving a bonus
payment either if the advisee guessed their rank accurately or if the
advisee evaluated the adviser as likeable as measured by an
unincentivized scale response. We hypothesize that advisers who want to
be liked by advisees will recommend that they bet on a lower rank, i.e.
that they are more attractive.

```{r}

study2_time1 <- readRDS(here::here("Data/Study2_time1.Rds"))
study2_time1_200 <- readRDS(here::here("Data/Study2_time1_200.Rds"))
study2_time2 <- readRDS(here::here("Data/Study2_time2.Rds"))
study2_finalRank <- readRDS(here::here("Data/Study2_finalRank.Rds"))
study2_time3 <- readRDS(here::here("Data/Study2_time3.Rds"))

```

## Methods

We recruited 300 participants from Prolific and, after asking
demographic questions, invited them to upload photos of themselves
(selfies) to be rated by other participants on attractiveness.
`r scales::comma(dim(study2_time1[study2_time1$gender == 'Male',])[1])`
men and
`r scales::comma(dim(study2_time1[study2_time1$gender == 'Female',])[1])`
women agreed to do so and uploaded pictures that adhered to our
instructions (e.g., did not include other people). We selected the first
100 photos from women to arrive at a gender-balanced sample
($M_{\text{Age}}$ = `r mean(study2_time1_200$age)` years). Participants
were informed that their selfies would be randomly grouped with those of
nine other participants of their gender and ranked in terms of
attractiveness by a group of new Prolific participants of the opposite
gender. Lastly, they guessed their rank (unincentivized).

Next, we recruited a new, gender-balanced sample for the role of
advisers. `r scales::comma(dim(study2_time2)[1])` participants from
Prolific ($M_{\text{Age}}$ = `r mean(study2_time2$age)` years;
`r round(mean(study2_time2$gender == "Female")*100, 2)`% Female) started
by providing demographic information, then were matched to a group of
the opposite gender. They then ranked participants from most to least
attractive using a drop-down menu next to each picture. Because of a
limitation with the survey software, participants were able to select
the same rank for multiple participants, rather than rank them uniquely.
We remove 115 participants who did not follow instructions and provide a
unique ranking.

After submitting their ratings, they saw the photo of the participant
they had ranked as the 7th most attractive (i.e., the 4th least
attractive). They were reminded of the rank they had given to that perso
and informed that this participant would be invited back and could earn
a \$1 bonus if they guessed their rank correctly. The rank was
determined by the aggregate ratings of all participants who had ranked
this group. Their task was to give advice to this participant about the
rank they should bet on based on having observed all ten selfies and
their own assessment. We randomly assigned participants to one of two
incentivization schemes. In the "Accuracy" treatment, they received a
bonus identical to the advisee: \$1 if they guessed their rank
correctly. In the "Likeability" treatment, we informed them that they
would be rated by the advisee on a 5-point Likert scale for how likeable
they thought they were. Each point on the scale would translate to a
bonus of 20 cents. They then selected a rank that they would recommend
the advisee to bet on.

Finally, we invited participants from Stage 1 and were ranked as the 7th
attractive by at least one adviser (so that they received advice) back
for the follow-up survey. Following our preregistration, we kept the
survey open for 7 days. In total,
`r scales::comma(dim(study2_time3)[1])` participants
(`r scales::comma(dim(study2_time3[study2_time3$gender == 'Male',])[1])`
men,
`r scales::comma(dim(study2_time3[study2_time3$gender == 'Female',])[1])`
women) returned. They were reminded of the selfie they uploaded in Stage
1 and informed that a group of 10 selfies, including theirs, had been
rated by other participants from Prolific. They then saw the advice from
a randomly selected adviser and made their estimate with a \$1 incentive
for guessing accurately. They then saw the rank they had been advised to
bet on one more time and were asked to rate the adviser's likability,
warmth, friendliness, good-naturedness, trustworthiness, and sincerity
on 5-point Likert scales (adapted from @Fiske2007).

## Results

```{r}
library(dplyr)

TargetsUnique <- study2_time2 %>%
  dplyr::select(Target) %>%
  unique()

TargetsUnique <- merge(TargetsUnique,study2_finalRank, by.x = 'Target', by.y = 'QID7_Id')

```

```{r study2advice, fig.cap="Advisers were asked to suggest what rank to bet on for someone they ranked as the 7th most attractive out of ten people. When incentivized for accuracy, most of them advised betting on 7, which was based on their evaluation of the recipient's attractiveness. However, when incentivized for likability, many participants gave flattering advice by suggesting betting on higher ranks that imply a higher evaluation of attractiveness."}

study2_time2 <- study2_time2 %>% dplyr::mutate(Advice = as.numeric(Advice))

ggplot(study2_time2, aes(x = Condition, y = Advice, fill = Condition)) +
    geom_violin(trim = FALSE) +
     scale_fill_manual(values = c("#2A629A",'#FF7F3E'), labels = c("Accurate", "Likable")) +
  geom_hline(yintercept = 7, col = 'grey', linetype = 'dashed', linewidth = 0.7) +
  scale_y_continuous(name = "Advised Rank", breaks = seq(1, 10)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20))

```

We begin by examining the prior beliefs of advisees who uploaded their
selfies. On average, men guessed they ranked
`r round(mean(study2_time1[study2_time1$gender == "Male",]$Guess),2)` in
their group of 10 and women guessed that they ranked
`r round(mean(study2_time1[study2_time1$gender == "Female",]$Guess),2)`.
Participants overall underestimate their attractiveness relative to the
benchmark average of 5.5. Moreover, women do so more than men
`r papaja::apa_print(t.test(Guess ~ gender, data = study2_time1, var.equal = T))$statistic`.
Notably, people's self-perceptions correlated strongly with the
aggregate ratings of the advisers
(`r round(cor(study2_finalRank$Guess, study2_finalRank$finalRank),2)`,
`r papaja::apa_print(cor.test(study2_finalRank$Guess, study2_finalRank$finalRank))$statistic`).
However, there was substantial heterogeneity in perceptions of
attractiveness. Of the 200 participants,
`r length(unique(study2_time2$Target))` were ranked as 7th most
attractive by at least one adviser. On average, men in this subset
estimated they were ranked
`r round(mean(TargetsUnique[TargetsUnique$gender == 'Male',]$Guess),2)`th
and women estimated they were ranked
`r round(mean(TargetsUnique[TargetsUnique$gender == 'Female',]$Guess),2)`th
(`r papaja::apa_print(t.test(Guess ~ gender, data = TargetsUnique, var.equal = T))$statistic`).

Next, we turn our attention to the advisers (Figure
\@ref(fig:study2advice)). In the Accuracy condition, those uploading
selfies were advised to bet on rank
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offered flattering advice.[^3] Importantly, and as predicted, we find
that advisers in the Likeable treatment recommend betting on a lower
rank, communicating that they think the participant in the selfie is
more attractive
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
This suggests that advisers inferred that advising someone that they
were more attractive would make the adviser appear more likeable and
therefore provided advice that communicated a more favorable impression
of the participants' attractiveness. The distribution shown in Figure
\@ref(fig:study2advice) shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

[^3]: This shading could be due to concerns of avoiding disappointment.
    However, it could also be that advisers are uncertain about the
    rankings they had given and make a recommendation that combines
    their own belief with a uniform prior.

<!-- David: Making a note here that we should be sure to discuss this at the end -->

<!-- as something for future research. -->

<!-- David: We should also report something on accuracy of the advice given across -->

<!-- the treatments. Even if no difference. -->

```{r study2regs}

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study2_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study2_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study2_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional.")

```

<!-- Amanda: How about this -->

<!-- For warmth it looks interesting, but I think the plot is hard -->

<!-- to read. Let's hold off on this for now. -->

Finally, we examine whether flattering advice indeed leads to more
positive evaluations of advisers, or whether flattering advice is
dismissed as insencere. Following our preregistration, we average the
ratings on likeability, warmth, friendliness, and good-naturedness to
create a scale of likeability ($\alpha$ =
`r cronbach.alpha(study2_time3[, 5:8])$alpha`); and we create a scale of
trustworthiness by averaging the ratings of trustworthiness and
sincerity ($\alpha$ = `r cronbach.alpha(study2_time3[, 9:10])$alpha`).

<!-- The Figure is not super clear. How about we try a series of box plots, -->

<!-- with the x-axis as the advised rank? Standard errors will be huge at the -->

<!-- extremes. -->

<!-- Also, we need to do something relative to the advisees' expectations. -->

<!-- I wonder what would be interesting here? Maybe averages for three groups: -->

<!-- more attractive than the participants' own guess, equally attractive, and less -->

<!-- attractive? -->

<!-- Amanda: How about the following picture? -->

```{r study2rating, fig.cap="Participants tend to rate advisors who suggest betting on a higher rank (implying greater attractiveness) as more likable and warm."}

library(ggplot2)
library(gridExtra)
library(dplyr)

study2_time3 <- study2_time3 %>% dplyr::mutate(Advice = as.factor(Advice))

# Create the first plot
plot1 <- ggplot(study2_time3, aes(x = Advice)) +
  geom_boxplot(aes(y = likable),position = position_dodge(width = 0.8), alpha = 0.8, color = "darkgrey", fill = "#2A629A") +
  scale_y_continuous(name = "Likability") +
  scale_x_discrete(name = "Advised Rank") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20))

# Create the second plot
plot2 <- ggplot(study2_time3, aes(x = Advice)) +
  geom_boxplot(aes(y = warmAvg),position = position_dodge(width = 0.8), alpha = 0.8, color = "darkgrey", fill = "#FF7F3E") +
  scale_y_continuous(name = "Warmth") +
  scale_x_discrete(name = "Advised Rank") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20))

# Combine the plots side by side
grid.arrange(plot1, plot2, ncol = 2)

```

```{r study2ratingByGroup, fig.cap="After participants received advice that suggested betting on a rank that was lower than, equal to, or higher than their self-expectations, they evaluated advisors based on their likability and warmth. Participants rated advisors who gave flattering advice and delivered a high evaluation of their attractiveness as more likable and warm than those whose advice implied a lower evaluation of their attractiveness."}

library(dplyr)
library(ggplot2)

study2_time3 <- study2_time3 %>% dplyr::mutate(Guess = as.numeric(Guess), Advice = as.numeric(Advice)) %>%
  dplyr::mutate(HorL = ifelse(Guess == Advice,0, ifelse(Guess > Advice, 1,-1))) %>%
  dplyr::mutate(HorL = as.factor(HorL)) %>% 
  dplyr::mutate(bonus = ifelse(bet == `Final Rank`,1,0))

ratings <- study2_time3 %>%
  group_by(HorL) %>%
  summarise(like = mean(likable), warmth = mean(warmAvg))

rating_table <- matrix(c(ratings$like,ratings$warmth), nrow = 2, byrow = TRUE, dimnames = list( c("Likable",'Warm'),c("Pessimistic", "Accurate", "Flattering")))

rating_df = as.data.frame.table(rating_table) %>%
  dplyr::rename( 'Condition'='Var1', 'Advice' ='Var2') %>%
  dplyr::mutate(Freq = round(Freq,2))

ggplot(rating_df, aes(x = Advice, y = Freq, fill = Condition)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#2A629A",'#FF7F3E')) +
  labs(x = "Advice", y = "Rating") +
  ggtitle('Response to Advice') +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),          
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20)) +
  geom_text(aes(label = paste0(Freq), y = 0.5), position = position_dodge(width = 0.9), size = 5, color = 'white', fontface = "bold")

```

As seen in Figure \@ref(fig:study2rating), we found that advisees who
suggested that the advisee was more attractive (lower rank) were indeed
rated as more likeable and warm (b = -0.109, p \< .05; b = -0.133, p \<
.001, respectively; See Columns 1 and 2 of Table \@ref(tab:study2regs)).
When using their self-expectations as a reference point, advisors who
suggest a rank lower than that point are perceived as less likable and
warm (see Figure \@ref(fig:study2ratingByGroup)). Interestingly, these
benefits are not at the cost of being hypocritical; advisers who
recommend a more favorable rank are viewed as no less trustworthy
(Column 3 of Table \@ref(tab:study2regs)).

We were not powered to do a comparison across the two experimental
groups and did not preregister such a difference. Indeed, we find no
difference in likeability and warmth across the two treatments
(`r papaja::apa_print(t.test(likable ~ Condition, data = study2_time3, var.equal = T))$statistic`,
and
`r papaja::apa_print(t.test(warmAvg ~ Condition, data = study2_time3, var.equal = T))$statistic`,
respectively).

## Discussion

In a context where advice communicates ego-relevant information (here,
people's attractiveness), we find that the advice people give is
contingent on their incentives. Specifically, when they get rewarded for
being more likeable, they recommend that the advisee bet on a more
favorable rank than when they are incentivized for accuracy.
Importantly, advisees do not discount flattering advice and instead
evaluate people who advise them to bet on a more attractive rank as
warmer and more likeable. These gains to interpersonal perceptions do
not come at the cost of trustworthiness, even when the advice is
substantially inflated (e.g., a recommendation that the advisee is the
most attractive in the group).

# General Discussion

Advice has the potential to shape people's career and personal outcomes.
Honest feedback, however, may be painful to learn if it falls short of
one's expectations. As prior work notes, this may motivate people to
avoid information and avoid seeking help [@Benabou2022; @Golman2017;
@Jaroszewicz2022]. We present evidence from two experiments that
advisers are also cognizant of this cost. As a result, they present
flattering advice that avoids disappointing the recipient, and correctly
anticipate that this boosts how advisees perceive them. However, this
flattering advice comes at a cost to advisees, who would do worse if
they followed it blindly, as we show in Study 1.

Moreover, a desire to avoid disappointment also means that advisers have
to take into account the expectations of the advisee. We document that
men are more optimistic about their performance than women are. As a
result, they receive more flattering advice and are more likely to be
told to aim higher. Notably, in the context of our experiment, this
turns out to be bad advice ex post.

Our findings have implications for organizational practice, where
mentoring and advice-giving may take into account an employees'
expectation. We document this as a novel source of gender bias.
Organizations could reduce this bias by calibrating employees'
expectations to reduce overconfidence.

Participants in our experiment were paired anonymously. Even so, we
document this desire to avoid disappointment. We anticipate that advice
would be more flattering in face-to-face communication and without
anonymity. Moreover, existing relationships might make it even more
difficult for advisers to be honest.

In our experiments, participants only received advice once and did not
evaluate the adviser after observing the outcome of their decision. For
example, advice that leads to bad outcomes may undermine the
interpersonal benefits of flattery. Alternatively, people may still like
the flattering advice and not fault the adviser for the bad outcome.
Moreover, our setting involved only a single piece of advice on one
task. Future research could examine whether people return to those who
gave them flattering advice, or if they prefer someone who gave them the
honest (but unpleasant) truth.

Advice has long been studied from the perspective of the receiver.
Similarly, research on belief utility has examined how recipients
respond to the valence of the information they receive. Here, we show
that advisers, too, take into account the psychological impact of the
information they convey. They may have even greater motivations to avoid
conveying bad news, because they incur the interpersonal costs of
delivering unfavorable information without reaping the benefits from
helping someone make a better choice.

# References

```{=tex}
\begingroup
\singlespacing
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\noindent
```
::: {#refs}
:::

```{=tex}
\endgroup
```