### 2024-12-15 manuscript/manuscript.Rmd
Edits up to study1regs

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle        : "Flattering Advice"
author: 
  - name          : "Amanda Chen"
    affiliation   : "1"
    corresponding : yes    
    email         : "zchengj@connect.ust.hk"
    address       : "Department of Management, The Hong Kong University of Science and Technology, Hong Kong"
  - name          : "David Hagmann"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "The Hong Kong University of Science and Technology"
note : |
  `r format(Sys.time(), "%e %B, %Y")`
abstract: |
  Good advice improves decision quality but often requires delivering unpleasant truths that may disappoint advisees. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (n = 3,900), we demonstrate that advisers prioritize avoiding disappointment, even at the expense of accuracy. In Study 1, advisers financially rewarded for accuracy still tailored recommendations to aspirational goals expressed by advisees, resulting in poorer advice. When incentivized to be liked, advisers provided even more flattering advice, and advisees rated these advisers as more likable, despite the advice being less honest and less accurate (Study 2). In Study 3, advisers incorporated advisees' subjective and uninformative performance expectations into recommendations, leading to systematically different advice for men and women. Because men expected to perform better than women, they received more aspirational advice, even when their gender was unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making and equity.

keywords          : "Gender difference, Advice, Interpresonal relationship"
wordcount         : "6,104"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa7"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
classoption       : "man"
header-includes:
   - \usepackage{setspace, appendix, placeins, booktabs, tikz, siunitx}
   - \usepackage{tabularx, epigraph, float, colortbl, tabu, pdflscape}
   - \usepackage{tabularray}
   - \usetikzlibrary{positioning}
   - \raggedbottom
   - \usepackage[maxfloats=256]{morefloats}
   - \maxdeadcycles=1000
bibliography      : bibliography.bib
output: 
  # papaja::apa6_docx
  papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
always_allow_html: true
---

```{r setup, include=FALSE}

library(magrittr)
library(ltm)
library(ggplot2)
library(gridExtra)
library(dplyr)

knitr::opts_chunk$set(fig.width = 8, fig.height = 4, cache = F, echo = F,
                      fig.align = 'center',warning = FALSE, message = FALSE)

study1_time1 <- readRDS(here::here("Data/Study1_time1.Rds"))
study1_ability50 <- read.csv(here::here("Data/Cleaned Data/S1_Ability50_clean.csv"))
study1_time2_giver <- readRDS(here::here("Data/Study1_time2_giver.Rds"))
study1_time2_advice <- readRDS(here::here("Data/Study1_time2_advice.Rds"))
study1_time3 <- readRDS(here::here("Data/Study1_time3.Rds"))

```

# Introduction
Advice plays a critical role in both professional and personal decision-making, 
often serving as a crucial tool for navigating complex challenges and improving 
outcomes [@Gino2007; @Harvey1997; @Soll2009]. While prior research has 
extensively examined the *recipients* of advice (e.g., how they evaluate or act 
on recommendations) far less attention has been devoted to understanding the 
motivations and behaviors of the *givers* of advice. In this paper, we shift the 
focus to advisers, proposing that their recommendations are shaped not only by 
accuracy considerations but also by interpersonal concerns. Specifically, we 
propose that advisers factor in advisees' subjective expectations, providing 
flattering advice that aligns with these expectations, even when doing so leads 
to objectively worse recommendations. Notably, this tendency persists in our 
experimental settings, where advisees' expectations are clearly uninformative 
and advisers are financially incentivized for accuracy.

Advice is often viewed as a mechanism to improve decision satisfaction or 
prediction accuracy, assuming both parties prioritize solving the problem at 
hand. However, prior research highlights that interpersonal dynamics can 
complicate these interactions. Advice seekers may avoid asking for help due to 
fear of rejection [@Benabou2022], and people generally have been shown to avoid
information that could lead to disappointment [@Gill2012; @Golman2017]. 
They are motivated to avoid unpleasant truths in an ethical dilemma [@Shalvi2019] 
and dislike those who deliver unwelcome feedback even when they had no role in
the outcome [@John2019]. These tendencies align with models of belief utility 
[@Loewenstein2018], which suggest that individuals experience 
hedonic costs when receiving bad news. A natural reference point for evaluating 
the valence of advice is the advisee's expectation in the absence of advice
[@Koszegi2006]. Consequently, advisers who are aware of expectations may 
strategically adapt their advice to avoid delivering bad news, even at the 
expense of accuracy.

This interpersonal dimension of advice-giving has significant implications for 
understanding systematic gender differences in advice quality. Men are more 
likely than women to receive aspirational advice and encouragement to pursue 
competitive or risky opportunities [@Kanze2018]. Conversely, women are often 
shielded from criticism [@Jampol2023], which, while intended to be kind, 
may ultimately limit their professional growth. We propose a novel mechanism to 
explain these patterns: advisers calibrate their recommendations to match 
advisees’ expectations. Because, in many tasks (including those in our experiments)
men are more confident and have higher expectations about their performance than
do women, they are more likely to receive flattering (but, in our setting, worse) 
advice.

We present evidence for this mechanism from three preregistered and incentivized 
experiments involving real interactions between advisers and advisees (N = 3,900). 
In Study 1, we experimentally manipulate advisee's expectations and show
that advisers account for those expectations, even when they offer no relevant
insight for the decision. In Study 2, we show that advisers put more weight on
interpersonal concerns when they are incentivized to be liked. Advisees reward
this flattery by rating advisers as more likeable and equally trustworthy compared
to those offering more honest (and accurate) feedback. In Study 3, we measure
expected performance on a math quiz and show that advisers take these expectations
into account even when they have information about actual performance. Because
men expect to score higher, they are more likely to be advised to compete against
a stronger group for a higher potential payment than are women, even when advisers
are unaware of the advisees' gender. Notably, this leads to worse advice for men.

# Open Science Statement

We report all manipulations, measures, and data exclusion in this and the following study. The preregistration reports, screenshots of all experimental materials, and the analysis code to replicate all statistical analyses and figures are available on the Open Science Framework
(<https://osf.io/8r3d4/?view_only=5ad7bafcd16b4d4ba08bb28b0e2bd02d>).

# Study 1
We first examine whether advisers take into account the expectations of advisees
when providing a recommendation. A group of participants ("Advisees") complete 
a quiz consisting of ten questions that draw on ego-relevant domains. They then
express a non-binding preference to compete against a group of high performers
or low performers on the same task, based on their past performance. We nudge
participants towards picking either the low or the high performer group by 
selecting one of the two by default. A second group of participants ("Advisers") 
observe Advisees' score on the quiz and, in one treatment, also which group
they had selected. We predicted that Advisees who were nudged toward the high
performer group would be more likely to receive advice to compete against this
high group when Advisers observed their non-binding choice.

## Methods
In a preliminary stage, we recruited `r scales::comma(dim(study1_ability50)[1])` 
participants from Prolific and gave them five minutes to complete a quiz 
consisting of ten items. The quiz included word puzzles,
identifying emotions from photos, and selecting the best responses for hypothetical
scenarios. These questions were adapted from surveys that measure problem-solving,
emotional intelligence, and communication skills---which we pre-tested as being
important in modern society and hence where participants might have a stake in
doing well. Participants received a bonus of five cents for each correctly
answered question. We then ranked them based on their score and label the 20
participants with the lowest (highest) scores as the Low (High) Performer Group.

Next, we recruited `r scales::comma(dim(study1_time1)[1])` participants who
completed the same 10-item quiz, also earning five cents for each correct answer.
We then asked them to express a non-binding preference for whether they would
like to compete against the High or Low Performer Group. We informed them that
they would be invited back at a later date when they could make a binding decision
and they could earn a bonus based on whether their score on this quiz was equal
to or higher than a randomly selected member from their comparison group.
If they picked the High Performer Group and had an equal or higher score than
a randomly selected member of that group, they would earn a bonus of 50 cents.
If they picked the Low Performer Group, they could earn a bonus of 20 cents.
If their score was lower in this comparison, they would not earn an additional
bonus. We randomized which of the two groups (High or Low Performers) was 
selected by default, and participants were free to select the other group.
Participants also made a guess as to their score, and the survey concluded with 
basic demographic questions.

We then recruited `r scales::comma(dim(study1_time2_giver)[1])`
participants for the role of advisers, and the focal part of our experiment. 
They were informed of the ability quiz that participants had completed, how the 
Low Performer Group and High Performer Groups were constructed, as well as the
choice and incentives for the other participants. We then asked participants
to give advice to ten participants on which group they should compete against.
Advisers were randomly assigned to one of two treatments. In the
"Performance" treatment, they observed the score of the other participant.
In the "Performance + Expectation" treatment, they additionally observed the
advisee's non-binding choice for which group to compete against.
However, since the outcome was based solely on the past quiz score and advisers,
but not advisees, know the score, this initial choice was not informative.
Advisers could earn the same bonus as one of the participants they had given
advice to and who returned to make a decision. The survey concluded with basic 
demographic questions.

Finally, we invited advisees back for the follow-up survey. Following our 
preregistration, we kept the survey open for 7
days. In total, `r scales::comma(dim(study1_time3)[1])` participants
returned. The short survey reminded them of the task they
completed in the previous survey, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally reminded them of how
many questions they guessed they had answered correctly. Importantly,
they were not informed of their true score or the score of the groups
they could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

## Results

```{r}
contingency_table <- table(study1_time1$Guess, study1_time1$default)

chi_test <- chisq.test(contingency_table)

```

We begin by examining the performance of the Stage 1 participants.
Because our default treatment takes place after participants completed the
ability quiz, we do not expect a difference in performance across
the Low and High Default treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study1_time1[study1_time1$default == "Low",]$Score),2)`
and
`r round(mean(study1_time1[study1_time1$default == "High",]$Score),2)`
for the Low Default and High Default treatments, respectively,
`r papaja::apa_print(t.test(Score ~ default, data = study1_time1, var.equal = T))$statistic`).
The default manipulation affected their initial choices. In the "Low Default" treatment, 
only `r round(mean(study1_time1[study1_time1$default == "Low",]$Guess),2)*100`% 
of participants initially preferred competing against the High Performer Group, 
compared to `r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`%
of participants who saw the High Performer Group highlighted ($\chi^2$(1, n = 201) = 36.62, p < .001).

```{r study1regs}

study1_advice <- study1_time2_advice %>%
  dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E', 1, 0))

modelsummary::modelsummary(
  list(`(1)` = lm(Advice ~ ExpShow * Treatment, data = study1_advice), 
       `(2)` = lm(Advice ~ Treatment + Score, data = study1_advice %>% dplyr::filter(ExpShow == 1))),
  vcov = ~GiverID,
  coef_map = c(
    "TreatmentHigh" = "High Default",
    "ExpShow" = "Expectation Shown",
    "ExpShow:TreatmentHigh" = "Expectation x High Default",
    "Score" = "Performance",
    "(Intercept)" = "Constant"
  ),
  stars = TRUE,
  gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
  caption = "Advice to compete against the High Performer Group in Study 1. Displaying the non-binding choice of those for whom the Low Performer Group was selected by default makes it more likely that they are advised to compete against the High Performer Group (Column 1). Column 2 restricts the analysis to advisers who observed advisees' initial non-binding choice and controls for advisees' score on the quiz. Standard errors are clustered at the adviser level."
)

```

We now examine whether advisers took the advisees' expectations into account. 
Recall that advisees who were defaulted to competing against the High Performer 
Group were indeed more likely to choose this group, which, from the advisers' 
perspective, could signal high confidence in their abilities. Our theory 
predicts that showing expectations should lead to more flattering advice—
specifically, a recommendation to compete against the High Performer Group—
for advisees with a high default choice compared to those with a low 
default choice. Column 1 of Table \@ref(tab:study1regs) reports a linear
probability model with advice to compete against the high group
as the outcome measure, incorporating the default choice advisees saw, whether expectations
were shown to the adviser, and the interaction of the two. Because each adviser made ten
recommendations, we cluster standard errors at the adviser level.
As predicted, we find a marginally significant interaction effect: advisees 
with a default choice to compete against the High Performer Group (nudged 
to have high expectations, though not necessarily holding them) were more 
likely to receive advice to compete against stronger competitors when 
their expectations were shown compared to when they were hidden.
We reported a linear probability model on advising to compete against the High Performer
group for advisers in the "Expectations" treatment, controlling for the
true performance of the advisee in Column 2 of Table \@ref(tab:study1regs).
We observe a significant main effect of the default treatment, where a high default
increased the likelihood of receiving flattering advice. It is worth noting that, 
since only `r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their default 
choice, this represents a conservative test of the effect of high expectations on 
the likelihood of receiving flattering advice.

```{r study1Advice, fig.cap ="Advisees who were nudged to express high expectations were more likely to receive advice to compete against the High Performer Group."}
study1_time2_advice %>%
  mutate(Treatment=factor(Treatment, levels=c('High','Low'), labels = c('High default', 'Low default'))) %>%
  group_by(Condition, Treatment) %>%
  summarize(avg_count = mean(Advice), ci = qnorm(0.975)*sd(Advice)/sqrt(dplyr::n())) %>%
  ggplot2::ggplot(ggplot2::aes(x = Condition, y = avg_count, fill = Treatment)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = avg_count - ci, ymax = avg_count + ci), 
                         position = ggplot2::position_dodge(0.9), linewidth = 0.2, width = 0.1) +
  ggtitle("Advice: Meeting Expectations vs. Absent Expectations") +
  scale_fill_manual(values = c("#2A629A",'#FF7F3E')) +
  theme_bw() +
  labs(y = "P(Compete Against High Performer)", x = NULL) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        axis.text.x = element_text(size = 12),
        legend.position = "none", # No legend
        plot.title = element_text(hjust = 0.5, size = 16)) +
  scale_x_discrete(labels = c("Expectation shown", "Expectation Absent")) +
   geom_text(aes(label=Treatment, y = -0.01), position = position_dodge(0.90))

```

## Discussion

When the quality of a decision does not depend on someone's expectations,
the adviser's awareness of the advisee's expectations should not influence
the advice given. However, in line with our argument that advisers take
into account the belief utility of the advisee and prefer to flatter 
rather than disappoint, we find that displaying the advisee's expectations
does make a difference. Specifically, advisees who expressed a preference
for stronger competitors were more likely to receive 'flattering' advice,
encouraging them to pursue that option. One possible explanation for this
finding is that, rather than trying to flatter, advisers may simply be 
telling advisees what they want to hear. In the next study, we aim to 
investigate the mechanism behind this discrepancy in advice-giving when
expectations are revealed, exploring the question: Why do people give 
flattering advice?

# Study 2

So far, we have shown that displaying expectations influences the advice given. 
Specifically, advisers were more likely to recommend competing against a group 
of high performers to advisees with higher performance expectations compared to 
those without. We now directly test whether this effect is driven by 
interpersonal concerns. In particular, we explore whether more flattering advice
stems from a desire to be liked, and whether advisees, in turn, like advisers 
more when they receive flattering advice.

We conducted a three-stage experiment in which we experimentally
manipulated the incentives for advisers. We began by inviting a group of
advisees to upload photos of themselves ("selfies") and informed them
that they would be rated on their attractiveness. We grouped them with
nine other participants of the same gender and recruited participants of
the opposing gender to rank them from most to least attractive and to
provide advice. Specifically, we asked them to advise the participant
they ranked as the 7th most attractive (4th least attractive) on what
rank they should bet they were ranked by a larger group of raters.
Advisers were randomly assigned to two treatments, receiving a bonus
payment either if the advisee guessed their rank accurately or if the
advisee evaluated the adviser as likeable as measured by an
unincentivized scale response. We hypothesize that advisers who want to
be liked by advisees will recommend that they bet on a lower rank, i.e.
that they are more attractive.

```{r}

study2_time1 <- readRDS(here::here("Data/Study2_time1.Rds"))
study2_time1_200 <- readRDS(here::here("Data/Study2_time1_200.Rds"))
study2_time2 <- readRDS(here::here("Data/Study2_time2.Rds"))
study2_finalRank <- readRDS(here::here("Data/Study2_finalRank.Rds"))
study2_time3 <- readRDS(here::here("Data/Study2_time3.Rds"))

```

## Methods

We recruited 300 participants from Prolific and, after asking
demographic questions, invited them to upload photos of themselves
(selfies) to be rated by other participants on attractiveness.
`r scales::comma(dim(study2_time1[study2_time1$gender == 'Male',])[1])`
men and
`r scales::comma(dim(study2_time1[study2_time1$gender == 'Female',])[1])`
women agreed to do so and uploaded pictures that adhered to our
instructions (e.g., did not include other people). We selected the first
100 photos from women to arrive at a gender-balanced sample
($M_{\text{Age}}$ = `r mean(study2_time1_200$age)` years). Participants
were informed that their selfies would be randomly grouped with those of
nine other participants of their gender and ranked in terms of
attractiveness by a group of new Prolific participants of the opposite
gender. Lastly, they guessed their rank (unincentivized).

Next, we recruited a new, gender-balanced sample for the role of
advisers. `r scales::comma(dim(study2_time2)[1])` participants from
Prolific ($M_{\text{Age}}$ = `r mean(study2_time2$age)` years;
`r round(mean(study2_time2$gender == "Female")*100, 2)`% Female) started
by providing demographic information, then were matched to a group of
the opposite gender. They then ranked participants from most to least
attractive using a drop-down menu next to each picture. Because of a
limitation with the survey software, participants were able to select
the same rank for multiple participants, rather than rank them uniquely.
We remove 115 participants who did not follow instructions and provide a
unique ranking.

After submitting their ratings, they saw the photo of the participant
they had ranked as the 7th most attractive (i.e., the 4th least
attractive). They were reminded of the rank they had given to that person
and informed that this participant would be invited back and could earn
a \$1 bonus if they guessed their rank correctly. The rank was
determined by the aggregate ratings of all participants who had ranked
this group. Their task was to give advice to this participant about the
rank they should bet on based on having observed all ten selfies and
their own assessment. We randomly assigned participants to one of two
incentivization schemes. In the "Accuracy" treatment, they received a
bonus identical to the advisee: \$1 if they guessed their rank
correctly. In the "Likeability" treatment, we informed them that they
would be rated by the advisee on a 5-point Likert scale for how likeable
they thought they were. Each point on the scale would translate to a
bonus of 20 cents. They then selected a rank that they would recommend
the advisee to bet on.

Finally, we invited participants from Stage 1 and were ranked as the 7th
attractive by at least one adviser (so that they received advice) back
for the follow-up survey. Following our preregistration, we kept the
survey open for 7 days. In total,
`r scales::comma(dim(study2_time3)[1])` participants
(`r scales::comma(dim(study2_time3[study2_time3$gender == 'Male',])[1])`
men,
`r scales::comma(dim(study2_time3[study2_time3$gender == 'Female',])[1])`
women) returned. They were reminded of the selfie they uploaded in Stage
1 and informed that a group of 10 selfies, including theirs, had been
rated by other participants from Prolific. They then saw the advice from
a randomly selected adviser and made their estimate with a \$1 incentive
for guessing accurately. They then saw the rank they had been advised to
bet on one more time and were asked to rate the adviser's likability,
warmth, friendliness, good-naturedness, trustworthiness, and sincerity
on 5-point Likert scales (adapted from @Fiske2007).

## Results

```{r}
library(dplyr)

TargetsUnique <- study2_time2 %>%
  dplyr::select(Target) %>%
  unique()

TargetsUnique <- merge(TargetsUnique,study2_finalRank, by.x = 'Target', by.y = 'QID7_Id')

```

```{r study2advice, fig.cap="Advisers were asked to suggest what rank to bet on for someone they ranked as the 7th most attractive out of ten people. When incentivized for accuracy, most of them advised betting on 7, which was based on their evaluation of the recipient's attractiveness. However, when incentivized for likability, many participants gave flattering advice by suggesting betting on higher ranks that imply a higher evaluation of attractiveness."}

library(ggplot2)
library(ggridges)

study2_time2 <- study2_time2 %>% dplyr::mutate(Advice = as.numeric(Advice))

ggplot(study2_time2, aes(x = Advice, y = Condition, fill = Condition)) +
    # Reduce vertical spacing with rel_min_height and scale
    geom_density_ridges(scale = 1.0, rel_min_height = 0.01, alpha = 0.7) +

    # Shading the left side of x = 7 in light grey
    geom_rect(aes(xmin = 7, xmax = Inf, ymin = -Inf, ymax = Inf), fill = "gray95", alpha = 0.01) +

    # Add the dashed line at x = 7
    geom_vline(xintercept = 7, col = 'black', linetype = 'dashed', linewidth = 0.7) +

    # Add the text "Flattering" on the left of the line
    annotate("text", x = 5, y = 0.6, label = "Flattering Ranks", hjust = 1, size = 8, fontface = "bold") +

    # Manual fill colors
    scale_fill_manual(values = c("#2A629A", '#FF7F3E'), labels = c("Accurate", "Likable")) +

    # X-axis adjustments
    scale_x_continuous(name = "Advised Rank", breaks = seq(1, 10)) +

    # Theme adjustments
    theme_bw() +
    theme(panel.grid = element_blank(),
          panel.border = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = "bottom",
          legend.title = element_blank(),
          legend.key.size = unit(1.2, "lines"),
          legend.text = element_text(size = 12),
          plot.title = element_text(hjust = 0.5, size = 20),
          axis.text.x = element_text(size = 14),  # X-axis text size
          axis.text.y = element_text(size = 14),
          axis.title.x = element_text(size = 16, face = "bold"),  # X-axis title size
          axis.title.y = element_text(size = 16, face = "bold"))

```

We begin by examining the prior beliefs of advisees who uploaded their
selfies. On average, men guessed they ranked
`r round(mean(study2_time1[study2_time1$gender == "Male",]$Guess),2)` in
their group of 10 and women guessed that they ranked
`r round(mean(study2_time1[study2_time1$gender == "Female",]$Guess),2)`.
Participants overall underestimate their attractiveness relative to the
benchmark average of 5.5. Moreover, women do so more than men
`r papaja::apa_print(t.test(Guess ~ gender, data = study2_time1, var.equal = T))$statistic`.
Notably, people's self-perceptions correlated strongly with the
aggregate ratings of the advisers
(`r round(cor(study2_finalRank$Guess, study2_finalRank$finalRank),2)`,
`r papaja::apa_print(cor.test(study2_finalRank$Guess, study2_finalRank$finalRank))$statistic`).
However, there was substantial heterogeneity in perceptions of
attractiveness. Of the 200 participants,
`r length(unique(study2_time2$Target))` were ranked as 7th most
attractive by at least one adviser. On average, men in this subset
estimated they were ranked
`r round(mean(TargetsUnique[TargetsUnique$gender == 'Male',]$Guess),2)`th
and women estimated they were ranked
`r round(mean(TargetsUnique[TargetsUnique$gender == 'Female',]$Guess),2)`th
(`r papaja::apa_print(t.test(Guess ~ gender, data = TargetsUnique, var.equal = T))$statistic`).

Next, we turn our attention to the advisers (Figure
\@ref(fig:study2advice)). In the Accuracy condition, those uploading
selfies were advised to bet on rank
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offered flattering advice.[^3] Importantly, and as predicted, we find
that advisers in the Likeable treatment recommend betting on a lower
rank, communicating that they think the participant in the selfie is
more attractive
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
This suggests that advisers inferred that advising someone that they
were more attractive would make the adviser appear more likeable and
therefore provided advice that communicated a more favorable impression
of the participants' attractiveness. The distribution shown in Figure
\@ref(fig:study2advice) shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

[^3]: This shading could be due to concerns of avoiding disappointment.
    However, it could also be that advisers are uncertain about the
    rankings they had given and make a recommendation that combines
    their own belief with a uniform prior.

<!-- David: Making a note here that we should be sure to discuss this at the end -->

<!-- as something for future research. -->

<!-- David: We should also report something on accuracy of the advice given across -->

<!-- the treatments. Even if no difference. -->

```{r study2regs}

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study2_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study2_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study2_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional.")

```

Finally, we examine whether flattering advice indeed leads to more
positive evaluations of advisers, or whether flattering advice is
dismissed as insencere. Following our preregistration, we average the
ratings on likeability, warmth, friendliness, and good-naturedness to
create a scale of likeability ($\alpha$ =
`r cronbach.alpha(study2_time3[, 5:8])$alpha`); and we create a scale of
trustworthiness by averaging the ratings of trustworthiness and
sincerity ($\alpha$ = `r cronbach.alpha(study2_time3[, 9:10])$alpha`).

```{r study2rating, fig.cap="Participants tend to rate advisors who suggest betting on a higher rank (implying greater attractiveness) as more likable and warm."}

library(ggplot2)
library(dplyr)
library(tidyr)

study2_time3 <- study2_time3 %>% dplyr::mutate(Advice = as.numeric(Advice))

# Reshape the data to long format
study2_long <- study2_time3 %>%
  pivot_longer(cols = c(likable, warmAvg), 
               names_to = "Evaluation", 
               values_to = "Value")

# Create the plot
ggplot(study2_long, aes(x = Advice, y = Value, color = Evaluation)) +  
  # Add points for each observation
  geom_point(size = 2, alpha = 0.7) +

  # Add linear regression lines for each metric
  geom_smooth(method = "lm", se = TRUE, aes(fill = Evaluation), alpha = 0.2, show.legend = FALSE) +

  # Customize axis labels
  scale_y_continuous(name = "Evaluation",limits = c(0.9, 5.1)) +  
  scale_x_continuous(name = "Advised Rank",limits = c(1, 10), breaks = c(1,2,3,4,5,6,7,8,9,10)) +

  # Apply a theme
  theme_bw() +  
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "right",
        plot.title = element_text(hjust = 0.5, size = 20)) +

  # Set custom colors for each metric
  scale_color_manual(values = c("#FF7F3E", "#2A629A"), 
                     labels = c("Likability", "Warm")) + 
  scale_fill_manual(values = c("#FF7F3E", "#2A629A"))

```

```{r study2ratingByGroup, fig.cap="After participants received advice that suggested betting on a rank that was lower than, equal to, or higher than their self-expectations, they evaluated advisors based on their likability and warmth. Participants rated advisors who gave flattering advice and delivered a high evaluation of their attractiveness as more likable and warm than those whose advice implied a lower evaluation of their attractiveness."}

library(dplyr)
library(ggplot2)
library(reshape2)

study2_time3 <- study2_time3 %>% dplyr::mutate(Guess = as.numeric(Guess), Advice = as.numeric(Advice)) %>%
  dplyr::mutate(HorL = ifelse(Guess == Advice,0, ifelse(Guess > Advice, 1,-1))) %>%
  dplyr::mutate(HorL = as.factor(HorL)) %>% 
  dplyr::mutate(bonus = ifelse(bet == `Final Rank`,1,0)) %>%
  dplyr::mutate(accuracy = `Final Rank`-Advice)

ratings <- study2_time3 %>%
  group_by(HorL) %>%
  summarise(like = round(mean(likable),2), warmth = round(mean(warmAvg),2))

rating_table <- matrix(c(ratings$like,ratings$warmth), nrow = 2, byrow = TRUE, dimnames = list( c("Likable",'Warm'),c("Lower than", "Equal to", "Higher than")))

rating_df = as.data.frame.table(rating_table) %>%
  dplyr::rename( 'Condition'='Var1', 'Advice' ='Var2') %>%
  dplyr::mutate(Freq = round(Freq,2))

# Create a data frame similar to your structure
data <- rating_df

# Reshape the data to wide format for the heatmap
heatmap_data <- dcast(data, Condition ~ Advice, value.var = "Freq")

# Melt the data for ggplot
melted_data <- melt(heatmap_data, id.vars = "Condition", variable.name = "Advice", value.name = "Rating")

# Create the heatmap
ggplot(melted_data, aes(x = Advice, y = Condition, fill = Rating)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "grey80", high = "grey30", name = "Rating") +  # Grey gradient from light to dark
  geom_text(aes(label = sprintf("%.2f", Rating)), color = "white", size = 8) +
  theme_minimal() +
  labs(x = "Attractiveness (Advice Relative to Self-Expectations)", y = "Evaluation") +
  theme(
    axis.text.x = element_text(size = 12),  # X-axis text size
          axis.text.y = element_text(size = 12),
          axis.title.x = element_text(size = 12),  # X-axis title size
          axis.title.y = element_text(size = 12))

```

As seen in Figure \@ref(fig:study2rating), we found that advisees who
suggested that the advisee was more attractive (lower rank) were indeed
rated as more likeable and warm (b = -0.109, p \< .05; b = -0.133, p \<
.001, respectively; See Columns 1 and 2 of Table \@ref(tab:study2regs)).
When using their self-expectations as a reference point, advisors who
suggest a rank lower than that point are perceived as less likable and
warm (see Figure \@ref(fig:study2ratingByGroup)). Interestingly, these
benefits are not at the cost of being hypocritical; advisers who
recommend a more favorable rank are viewed as no less trustworthy
(Column 3 of Table \@ref(tab:study2regs)).

We were not powered to do a comparison across the two experimental
groups and did not preregister such a difference. Indeed, we find no
difference in likeability and warmth across the two treatments
(`r papaja::apa_print(t.test(likable ~ Condition, data = study2_time3, var.equal = T))$statistic`,
and
`r papaja::apa_print(t.test(warmAvg ~ Condition, data = study2_time3, var.equal = T))$statistic`,
respectively). We also assess the quality of advice by measuring the 
discrepancy between the advised rank and the final rank. Our analysis
revealed no significant difference between the two treatments in terms 
of how helpful the advice was for advisees in making accurate guesses,
`r papaja::apa_print(t.test(accuracy ~ Condition, data = study2_time3, var.equal = T))$statistic`.

## Discussion

In another context where advice communicates ego-relevant information (here,
people's attractiveness), we find that the advice people give is
contingent on their incentives. Specifically, when they get rewarded for
being more likeable, they recommend that the advisee bet on a more
favorable rank than when they are incentivized for accuracy.
Importantly, advisees do not discount flattering advice and instead
evaluate people who advise them to bet on a more attractive rank as
warmer and more likeable. These gains to interpersonal perceptions do
not come at the cost of trustworthiness, even when the advice is
substantially inflated (e.g., a recommendation that the advisee is the
most attractive in the group).

# Study 3

```{r}

study3_math50 <- read.csv(here::here("Data/Cleaned Data/S3_Math50_cleaned.csv"))
study3_time1All <- readRDS(here::here("Data/Study3_time1All.Rds"))
study3_time1 <- readRDS(here::here("Data/Study3_time1.Rds"))
study3_time2_giver <- readRDS(here::here("Data/Study3_time2_giver.Rds"))
study3_time2_giver50 <- readRDS(here::here("Data/Study3_time2_giver50.Rds"))
study3_advice <- readRDS(here::here("Data/Study3_time2_advice.Rds"))
study3_time3 <- readRDS(here::here("Data/Study3_time3.Rds"))

```

In Study 3, we replicate the main effect in a different ego-relevant domain: 
intelligence. The experimental design is conceptually identical to Study 1, 
but this time we utilize a more direct measure of expectation by asking 
participants to estimate their scores.

This three-stage experiment begins with measuring actual performance. 
Participants are then anchored to low or high performance expectations 
and asked to guess how many questions they answered correctly. As in 
previous studies, we recruit advisers who are presented with either 
just the advisee's true performance or both the true performance and 
the number of questions the advisee believes they answered correctly. 
Advisers are then asked to provide advice on choosing a competitor. 
Finally, we present the advice to the advisees and ask them to make 
their final choice. Our key hypotheses are: (1) advisees anchored to 
high expectations will be more likely to receive advice to compete 
against the high-performing group, and (2) men will be more likely 
than women to be advised to compete against the high-performing group
when expectations are disclosed.

## Methods

```{r}

subgroup_averages <- data.frame(subgroup = numeric(), avg_score = numeric())
set.seed(807)
# loop through 1000 iterations of randomly selecting 10 participants and computing their average score
for (i in 1:1000) {
  # randomly select 10 participants
  subgroup <- sample(study3_math50$ProlificID, size = 5, replace = FALSE)
  # compute the average score for the selected participants
  subgroup_avg <- mean(study3_math50$Score[study3_math50$ProlificID %in% subgroup])
  # add the subgroup and its average score to the dataframe
  subgroup_averages <- rbind(subgroup_averages, data.frame(subgroup = i, avg_score = subgroup_avg))
}

p5 <- quantile(subgroup_averages$avg_score, probs = 0.05)
p95 <- quantile(subgroup_averages$avg_score, probs = 0.95)

sortedScore <- sort(study3_math50$Score, decreasing = TRUE)
LP_lower <- sortedScore[50]
LP_upper <- sortedScore[31]
HP_lower <- sortedScore[20]
HP_upper <- sortedScore[1]

```

We begin by first recruiting a sample
of `r scales::comma(dim(study3_math50)[1])` participants from Prolific
to complete a 10-question multiple choice mathematics quiz. The
questions were taken from a paper-version of the ASVAB standardized
exam, such that answers were not available online. Participants had five
minutes to answer the quiz and were paid 10 cents for each correctly
answered question. Like before, we define top 20 scores as "High Performers"
and the bottom 20 scorers as the "Low Performers." On average,
participants answered `r round(mean(study3_math50$Score),2)` questions
correctly, High Performers scored between `r HP_lower` and `r HP_upper`,
and Low Performers scored between `r LP_lower` and `r LP_upper`. In
order to anchor the expectations of participants in our main experiment,
we simulated 1,000 pairings of groups of five participants, with the
5<sup>th</sup> percentile of groups scoring an average of `r p5` and the
95 th percentile scoring an average of `r p95`. We report these averages
to participants in the Low Expectations and High Expectations treatment,
respectively.

We then recruited `r scales::comma(dim(study3_time1All)[1])`
participants for the role of advisees in Stage 1 of our main experiment.
To arrive at a gender-balanced sample, we dropped the last two male
participants to complete the survey, ending up with a sample of 500 men
and 500 women ($M_{\text{Age}}$ = `r mean(study3_time1$Age)`).
Participants completed the same 10-item mathematics quiz as the earlier
participants and were informed that their performance would affect their
bonus earnings in a follow-up stage to be conducted a few days later.
After completing the quiz, we informed them of the average score of a
group of five participants from the preliminary survey. We randomly
assigned them to learn about the 5<sup>th</sup> percentile of groups,
which scored `r p5`. ("Low Expectations" treatment) or the
95<sup>th</sup> percentile of groups, which scored `r p95` ("High
Expectations" treatment). Participants then made a guess
(unincentivized) about how many questions they think they answered
correctly. The survey concluded with basic demographic questions.

We then recruited `r scales::comma(dim(study3_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We began
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 had completed, and informed them of the
average score of all participants in the preliminary study. Advisers had
to recommend whether an advisee should compete against the Low
Performers or the High Performers (we used these terms in the survey).
We anticipated that being told to compete against High Performers is
more flattering and hence being told to compete against the Low
Performers would be disappointing if one had expected to do well. Advisees
could earn a bonus if their score was equal to or higher than that of 
a randomly selected member from their chosen group. Competing against 
the High Performers was more challenging, resulting in an average bonus
of 50 cents, compared to 30 cents when competing against other groups.

Like before, advisers were randomly assigned to one of two treatments. In the
"Baseline" treatment, they only observed the score of the advisee on the
mathematics quiz. In the "Expectation" treatment, they observed the
score as well as the advisee's guess for how many questions they
answered correctly. We want to emphasize that sice the outcome is determined only by
the past score on the quiz, the advisee's guess is immaterial to which
group they should compete against. Moreover, in neither treatment did
they receive any demographic information about their advisee.
Participants gave recommendations to 10 advisees, which unbeknownst to
them were five men and five women matched to have identical performance
on the test.[^1] They were informed that if their advice was shown to a
participant who returned for the follow-up survey, they would receive
the identical bonus as that participant. The survey then concluded with
basic demographic questions.

[^1]: We made this decision to account for the possibility of gender
    differences in performance.

Finally, we invited participants from Stage 1 back for the follow-up
survey. Following our preregistration, we kept the survey open for 7
days. In total, `r scales::comma(dim(study3_time3)[1])` participants
(`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Female',])[1])`
women) returned. The brief survey reminded them of the task they
completed in Stage 1, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally were reminded them of how
many questions they guessed they had answered correctly. Importantly,
they were not informed of their true score or the score of the groups
they could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

## Results

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants completed the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study3_time1[study3_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study3_time1, var.equal = T))$statistic`).
The expectations treatment did, however, affect how well they thought
they performed. Participants in the "Low Expectations" treatment guessed
a score of
`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study3_time1[study3_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study3_time1, var.equal = T))$statistic`),
showing that the manipulation was successful, albeit small. Contrary to
our expectations, we did find a gender difference in performance: men
scored
`r round(mean(study3_time1[study3_time1$Gender == "Male",]$count),2)` on
average, while women scored
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study3_time1, var.equal = T))$statistic`).
Consistent with this difference, men thought they answered more
questions correctly than did women
(`r round(mean(study3_time1[study3_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study3_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We
observe that women underestimate their performance by 0.76 points, while
men do so by only 0.27 points
`r papaja::apa_print(t.test(overconfidence ~ Gender, data = study3_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`.
Although neither gender is overconfident, men on average are more
confident in their performance than are women, as we had expected.

```{r study3regs}
study3_advice <- study3_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study3_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study3_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study3_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor.")

```

```{r study3genderdiff, fig.cap ="Advisers observed real performance and the participants' estimated performance, but not their gender. As a result of their higher expectations, men (blue) were advised to compete against the high performance group more often than women (red). The grey line shows advice absent expectations, which did not differ by gender."}
library(dplyr)
library(ggplot2)

# Calculate averages for the performance, grouped by condition and gender
ProbHG <- study3_advice %>%  
  group_by(Condition, Performance, Gender) %>%  
  summarise(avg_hardGroup = mean(Advice), .groups = 'drop')

# Create a data frame for the NONE condition as the baseline
baseline <- ProbHG %>%  
  filter(Condition == "None") %>%  
  select(Performance, baseline_avg = avg_hardGroup)

# Merge with the original data to get comparison values
comparison_data <- ProbHG %>%  
  filter(Condition != "None") %>%  
  left_join(baseline, by = "Performance") %>%  
  mutate(diff = avg_hardGroup - baseline_avg)  # Calculate the difference

# Create a scatter plot comparing conditions with NONE as baseline
ggplot() +
  # Step lines for the NONE condition
  geom_segment(data = baseline,
               aes(x = Performance, 
                   xend = lead(Performance, 1),  # Create horizontal lines to the next performance point
                   y = baseline_avg, 
                   yend = baseline_avg,
                   linetype = "None"), 
               color = "grey70", size = 0.5, 
               na.rm = TRUE,
               show.legend = TRUE) + 
  geom_point(data = comparison_data %>% filter(Gender == "Male"), 
             aes(x = Performance + 0.5, y = avg_hardGroup, shape = Gender, color = Gender), 
             size = 4) +  # Males with an offset to the left
  geom_point(data = comparison_data %>% filter(Gender == "Female"), 
             aes(x = Performance + 0.5, y = avg_hardGroup, shape = Gender, color = Gender), 
             size = 4) +  # Females with an offset to the right
  scale_shape_manual(values = c(16, 17), labels = c("Female", "Male")) +  # Circle for Females, Triangle for Males
  scale_color_manual(values = c( "#FF7F3E","#2A629A"), 
                     labels = c("Female", "Male")) +  
  scale_y_continuous(name = "P(Compete Against High Performance)",
                     labels = scales::percent,
                     limits = c(0, 1.1),
                     breaks = c(0,1)) +  # Set limits for y-axis
  scale_x_continuous(name = "Performance",
                     limits = c(0, 11),expand = c(0, 0), 
                     breaks = seq(0, 10, 1)) +  
  theme_bw() +  
  theme(panel.grid = element_blank(), 
        panel.border = element_blank(), 
        axis.line = element_line(colour = "grey"), 
        legend.position = "bottom", 
        legend.title = element_blank(), 
        legend.key.size = unit(1.2, "lines"), 
        legend.text = element_text(size = 12), 
        plot.title = element_text(hjust = 0.5, size = 20))

```

Next, we examine whether advisers took the advisee's expectations into
account. Column 1 of Table \@ref(tab:study3regs) reports a linear
probability model on advising to compete against the High Performer
group for advisers in the "Expectations" treatment, controlling for the
true performance of the advisee. Because each adviser made ten
recommendations, we cluster standard errors at the adviser level.
However, contrary to our expectations, we do not see a significant
effect of the expectations treatment. This may be because the induced
difference in expectations was too small.

However, recall that men were more confident in their performance than
were women. Our theory thus predicts that showing expectations should
lead to more flattering advice (i.e., a recommendation to compete
against the High Performer group) for men than for women. We report a
linear probability model with advice to compete against the high group
as the outcome measure, and the advisee's gender, whether expectations
were shown to the adviser, and the interaction of the two in Column 2 of
Table \@ref(tab:study3regs). As predicted, we find a significant
interaction effect: men are more likely to be advised to compete against
the High Performers when expectations are shown than when they are
hidden ($p < 0.001$). We show this result graphically in Figure
\@ref(fig:study3genderdiff)).

To examine the quality of advice, we computed the expected bonus
earnings for someone who followed the recommendations. For example, if
an adviser suggested competing against the High Performer group, we
matched the advisee against all 20 members of that group and determined
how often their score matched or exceeded that of the member. We then
multiplied this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of Table
\@ref(tab:study3regs). Displaying expectations led men to be advised to
compete against the High Performer group more often, and this advise
turned out to be bad: men receive worse advise than do women when
expectations are displayed, but not in their absence.[^2]

[^2]: This analysis was not preregistered, and we note here that the
    reduction in expected earnings is small. However, it is interesting
    that expectations have a negative effect for men who underestimate
    their performance on average. One possibility is that advisers
    suggest the High Performer group more often than is optimal. We
    return to the possibility that advice is overall biased to be
    flattering in Study 2.

```{r study3time3}

study3_time3 <- study3_time3 %>% dplyr::mutate(AdviseeCondition = factor(AdviseeCondition, levels = c('Low', 'High'))) %>%
     dplyr::mutate(Adopt = ifelse(Bet == Advice,1,0))

modelsummary::modelsummary(list(`(1)` = lm(bonus ~ AdviseeCondition + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(2)` = lm(bonus ~ Gender + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(3)` = lm(Adopt ~ Gender + GiverCondition + Advice, data = study3_time3),
                                `(4)` = lm(Adopt ~ Gender + GiverCondition + Advice + Gender*GiverCondition + GiverCondition*Advice + Gender*Advice, data = study3_time3)),
                           vcov = c("classical", "classical","classical","classical"),
                            coef_map = c("AdviseeScore" = "Performance",
                                         "AdviseeConditionHigh" = "High Expectation",
                                        "GenderMale" = "Advisee Male",
                                        "GenderMale:GiverCondition" = "Expectation x Male",
                                        "GiverCondition" = "Expectation Shown",
                                        "Advice1" = "Advice: High Performer",
                                        "GenderMale:Advice1" = "Advice: High Performer x Male",
                                        "GiverCondition:Advice1" = "Expectation x Advice: High Performer",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the actual bonus based on the performance and whether the participant is primed with high expectations. Column 2 displays the actual bonus based on the performance and gender of the participant. Columns 3 and 4 display the chance of adopting the advice based on the gender of the participants, whether the advisee sees expectations, and whether the advice is to compete against high performers. The former considers only the main effect, while the latter also includes the interactive effect.")

```

To determine whether flattering advice is truly costly, however, we need
to examine the outcome of the advisees. In particular, they could ignore
flattering advice, recognizing it as such and thus failing to adhere to
it. Align with our prediction, participants who were primed with high
expectations earned less when their expectations were conveyed to the
advisers, although this result is only directional (see Column 1 of
Table \@ref(tab:study3time3)). Similarly, as shown in Column 2 of Table
\@ref(tab:study3time3), we also found that male participants received
less reward when their expectations were presented, even when their
performance was the same. These findings suggest that flattering advice
is not without consequences. Furthermore, our research indicates that
because males are more likely to follow such advice (see Column 4 of
Table \@ref(tab:study3time3), their tendency towards competition
intensifies the cost of such advice.

## Discussion

We found supportive evidence for our argument in a different domain. 
Advisors tend to consider advisees' self-expectations and match 
their advice accordingly, aiming to avoid disappointment. Consistent
with the confidence gap narrative, men tend to underestimate their 
mathematics test scores less than women. When these expectations are
displayed to advisors, they are more likely to advise men to compete
against high performers. Notably, this turns out to be poor advice: 
men whose advisors were aware of their expectations received worse 
advice. Our findings suggest that men with the same scores as their
female counterparts ended up earning less, although this result is 
only directional. This discrepancy may be due to men receiving more
favorable advice, being more likely to follow it, and ultimately 
facing worse outcomes.

<!-- David: One more sentence about the earnings difference here, which does not -->

<!-- differ significantly. And one sentence about the new analysis of whether -->

<!-- people follow the advice or not. -->

# General Discussion

Advice has the potential to shape people's career and personal outcomes.
Honest feedback, however, may be painful to learn if it falls short of
one's expectations. As prior work notes, this may motivate people to
avoid information and avoid seeking help [@Benabou2022; @Golman2017;
@Jaroszewicz2022]. We present evidence from two experiments that
advisers are also cognizant of this cost. As a result, they present
flattering advice that avoids disappointing the recipient, and correctly
anticipate that this boosts how advisees perceive them. However, this
flattering advice comes at a cost to advisees, who would do worse if
they followed it blindly, as we show in Study 1.

Moreover, a desire to avoid disappointment also means that advisers have
to take into account the expectations of the advisee. We document that
men are more optimistic about their performance than women are. As a
result, they receive more flattering advice and are more likely to be
told to aim higher. Notably, in the context of our experiment, this
turns out to be bad advice ex post.

Our findings have implications for organizational practice, where
mentoring and advice-giving may take into account an employees'
expectation. We document this as a novel source of gender bias.
Organizations could reduce this bias by calibrating employees'
expectations to reduce overconfidence.

Participants in our experiment were paired anonymously. Even so, we
document this desire to avoid disappointment. We anticipate that advice
would be more flattering in face-to-face communication and without
anonymity. Moreover, existing relationships might make it even more
difficult for advisers to be honest.

In our experiments, participants only received advice once and did not
evaluate the adviser after observing the outcome of their decision. For
example, advice that leads to bad outcomes may undermine the
interpersonal benefits of flattery. Alternatively, people may still like
the flattering advice and not fault the adviser for the bad outcome.
Moreover, our setting involved only a single piece of advice on one
task. Future research could examine whether people return to those who
gave them flattering advice, or if they prefer someone who gave them the
honest (but unpleasant) truth.

Advice has long been studied from the perspective of the receiver.
Similarly, research on belief utility has examined how recipients
respond to the valence of the information they receive. Here, we show
that advisers, too, take into account the psychological impact of the
information they convey. They may have even greater motivations to avoid
conveying bad news, because they incur the interpersonal costs of
delivering unfavorable information without reaping the benefits from
helping someone make a better choice.

# References

```{=tex}
\begingroup
\singlespacing
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\noindent
```
::: {#refs}
:::

```{=tex}
\endgroup
```