### 2025-01-09 manuscript/manuscript.tex

\documentclass[
  man,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{Flattering Advice: Avoiding Disappointment in Advice-Giving}

\shorttitle{Flattering Advice}

\usepackage{etoolbox}

\authorsnames{Amanda Chen,David Hagmann}

\authorsaffiliations{
{Department of Management, The Hong Kong University of Science and
Technology},{}}

\leftheader{Chen and Hagmann}

\date{Invalid Date}

\abstract{Good advice improves decision quality but often requires
delivering unpleasant truths that may disappoint advisees. Across three
pre-registered and incentivized experiments involving real
adviser-advisee interactions (N = 3,900), we show that advisers
prioritize avoiding disappointment at the expense of accuracy and their
own earnings. In Study 1, advisers financially rewarded for accuracy
still tailor recommendations to aspirational goals expressed by
advisees, resulting in worse advice. When incentivized to be liked,
advisers provide even more flattering advice, and advisees reward this
by rating these advisers as more likable, despite the advice being less
honest and less accurate (Study 2). The desire to avoid disappointment
may lead to inequities if there are differences in expectations across
social groups. In Study 3, we examine a setting in which men expect to
perform better than women. We show that advisers take into account these
expectations, leading to systematically different advice for men and
women even when their gender is unknown to advisers. Advisers' efforts
to avoid disappointment may thus contribute to systematic gender
disparities in advice, with implications for downstream
decision-making.}

\keywords{Gender difference, Advice, Interpresonal relationship}

\authornote{\par{\addORCIDlink{Amanda
Chen}{0000-0002-4428-627X}}\par{\addORCIDlink{David
Hagmann}{0000-0002-2080-997X}}

\par{     The authors are grateful to XXX for their helpful comments.  }
\par{Correspondence concerning this article should be addressed
to Amanda Chen, Email: zchengj@connect.ust.hk}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

\usepackage{float}
\usepackage{tabularray}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\UseTblrLibrary{booktabs}
\UseTblrLibrary{rotating}
\UseTblrLibrary{siunitx}
\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

%% Disable longtable counter
%% https://tex.stackexchange.com/a/248395/211326

\usepackage{etoolbox}

\makeatletter
\patchcmd{\LT@caption}
  {\bgroup}
  {\bgroup\global\LTpatch@captiontrue}
  {}{}
\patchcmd{\longtable}
  {\par}
  {\par\global\LTpatch@captionfalse}
  {}{}
\apptocmd{\endlongtable}
  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
  {}{}
\newif\ifLTpatch@caption
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{Introduction}\label{introduction}

Career trajectories for men and women often differ substantially, even
among those with similar qualifications. One source of such differences
may be the advice that people receive. Prior work has documented that
men receive more aspirational advice and women more risk-averse advice,
which may explain why women are less likely to apply for and ultimately
obtain more rewarding positions (Kanze et al., 2018). Research on gender
discrimination has focused on the role of unconscious bias (Greenwald et
al., 2009; Greenwald \& Banaji, 1995), but interventions that try to
reduce this bias have been largely unsuccessful (Chang et al., 2019;
Paluck \& Green, 2009). Recent work has proposed cognitive, rather than
unconscious, biases that can lead to gender discrimination through the
formation of false beliefs (Hagmann et al., 2024). Here, we propose a
novel source of differences in advice by drawing on recent work on
belief-based utility, expectation-based reference points (Köszegi, 2006;
Kőszegi \& Rabin, 2009), and the hedonic consequences of information
(Golman et al., 2017).

Advice figures prominently in workplace and personal decision-making,
shaping choices about promotions, career paths, and everyday dilemmas
(Gino \& Moore, 2007; Harvey \& Fischer, 1997; Soll \& Larrick, 2009).
Its fundamental premise is to improve outcomes by enabling
decision-makers to learn from others' expertise. However, while honest
advice can lead to better decisions, it can also lead to disappointment
when it suggests a desired outcome is unlikely to materialize. Advisers
may be motivated to offer flattering recommendations that align with
advisees' hopes, sparing them the disappointment of a more sobering
forecast.

Advice with the potential for disappointment appears in diverse
organizational settings: supervisors deliver feedback on employee
performance, lawyers convey likely outcomes of litigation to their
clients, and academics offer thoughts on colleagues' manuscripts.
Ideally, honest information would help people identify their strengths
and weaknesses, make informed decisions about whether to pursue legal
cases, and strengthen papers prior to submission to academic journals,
thus improving long-term outcomes. In practice, however, interpersonal
considerations may prevent candor. When someone's expectations are (too)
high, an adviser might hesitate to deliver information that undermines
it---particularly if doing so risks negative interpersonal
repercussions.

People often go to great lengths to avoid unpleasant information,
deriving belief-based utility from maintaining favorable views of
themselves (Golman et al., 2017; Loewenstein \& Molnar, 2018;
\textbf{Ho2018?}). Moreover, they often prefer advisers who withhold bad
news from them (Shalvi et al., 2019) and punish those who communicate
bad news (John et al., 2019). Advisers, anticipating these preferences
and consequences face a dilemma: Should they provide the accurate but
potentially painful truth, or shade their recommendations to preserve
the advisee's mood and the adviser's own standing? Advisers who fear
they will be blamed or disliked for undermining someone's confidence may
reasonably avoid candid feedback, particularly if they don't incur a
cost when the advisee makes a mistake.

In our experiments, however, advisees have no opportunity to punish the
adviser, and advisers' financial incentives are linked to the decisions
the advisees make. We propose that advisers nonetheless have reason to
provide flattering advise because they recognize the psychological toll
of disappointing news. As a result, they may similarly incur a hedonic
cost for delivering unpleasant information that they anticipate may
distress the advisee. Thus, even in the anonymized context of an online
experiment, we hypothesize (and find) that advisers are relucant to
disappoint advisees and advice is thus biased upward, away from accuracy
and toward confirming advisees' priors. We show that this reduces the
quality of advice---but that advisees nonetheless view flattering
advisers more favorably.

We present the results of three preregistered experiments in which
participants are paired anonymously as advisers and advisees. Advisers
are incentivized to offer accurate advice, with bonus earnings depending
on the outcome of an advisee's decision. We study a setting in which
advisees can communicate their expectations to an adviser, and
experimentally manipulate whether advisers observe this expectation. Our
design simulates organizational contexts where mentors and managers
often have some sense of an employee's confidence or aspirations and
thus could adjust their feedback to avoid disappointing them.
Interpersonal considerations would likely be stronger in situations in
which the adviser and advisee have an existing relationship and when the
advice is delivered face-to-face. Moreover, advisers usually do not
suffer any direct costs when their flattering advice leads to a poor
outcome for the recipient, particularly when it is not clear what would
have happened under a counterfactual.\^{}{[}In some cases, however,
repeated interaction might also increase incentives for honest feedback.
Someone who is known to persistently give overoptimistic advice may be
viewed as less trustworthy in the long run.{]}

In Study 1, we nudge advisees into a preference for competing against
either a group of top-performers or a group of low-performers, through
the use of a default option. Advisers who observe the actual performance
of advisees and therefore have the relevant information needed to make a
recommendation nonetheless take into account the initial preference when
recommending which group they should compete against. In Study 2, we ask
advisees to upload photos of themselves and advisers rank them on their
attractiveness. Advisers then recommend which rank the advisee should
bet on, receiving an incentive either for the advisee's accuracy or for
being evaluated favorably. Both groups of advisers recommend a rank that
is more attractive than what they themselves have evaluated the advisee,
and those incentivized for likability further inflate the ranking.
Advisees evaluate advisers more favorably when they recommend betting on
a more attractive rank, including viewing them as more trustworthy.

In Study 3, we examine a setting in which participants report their
expected performance on a mathematics quiz. Advisers observe the
test-taker's true score and, in one treatment, the score they guessed
they would receive. They then recommend whether the advisee should
compete against a group of high performers or low performers.
Importantly, this ``competition'' is based on the past performance
observed by the adviser and thus the advisee's expectation does not
provide instrumental information. However, we find that (1) men expect
to perform better than women, given identical performance, and (2)
advisers take these expectations into account. As a result, when
expectations (but not gender) are known to advisers, the advice given to
men and women differs, and men are more often advised to compete against
the group of high performers. Notably, this leads to worse advice for
men than for women.

\section{Open Science Statement}\label{open-science-statement}

We report all manipulations, measures, and data exclusion in our
experiments. The preregistration reports, screenshots of all
experimental materials, and the analysis code to replicate all
statistical analyses and figures are available on the Open Science
Framework
(\url{https://osf.io/8r3d4/?view_only=5ad7bafcd16b4d4ba08bb28b0e2bd02d}).

\section{Study 1}\label{study-1}

We first examine whether advisers take into account the expectations of
advisees when providing a recommendation. A group of participants
(``Advisees'') complete a quiz consisting of ten questions that draw on
ego-relevant domains. They then express a non-binding preference to
compete against a group of high performers or low performers on the same
task, based on their past performance. We nudge participants towards
picking either the low or the high performer group by selecting one of
the two by default. A second group of participants (``Advisers'')
observe Advisees' score on the quiz and, in one treatment, also which
group they had selected. We predicted that Advisees who were nudged
toward the high performer group would be more likely to receive advice
to compete against this high group when Advisers observed their
non-binding choice.

\subsection{Methods}\label{methods}

In a preliminary stage, we recruited 50 participants from Prolific and
gave them five minutes to complete a quiz consisting of ten items. The
quiz included word puzzles, identifying emotions from photos, and
selecting the best responses for hypothetical scenarios. These questions
were adapted from surveys that measure problem-solving, emotional
intelligence, and communication skills---which we pre-tested as being
important in modern society and hence where participants might have a
stake in doing well. Participants received a bonus of five cents for
each correctly answered question. We then ranked them based on their
score and label the 20 participants with the lowest (highest) scores as
the Low (High) Performer Group.

Next, we recruited 201 participants for the role of Advisees. They
completed the same 10-item quiz, also earning five cents for each
correct answer. We then asked them to express a non-binding preference
for whether they would like to compete against the High or Low Performer
Group. We informed them that they would be invited back at a later date
when they could make a binding decision and they could earn a bonus
based on whether their score on this quiz was equal to or higher than a
randomly selected member from their comparison group. If they picked the
High Performer Group and had an equal or higher score than a randomly
selected member of that group, they would earn a bonus of 50 cents. If
they picked the Low Performer Group and outperformed a randomly selected
member of that group, they could earn a bonus of 20 cents. If their
score was lower in this comparison, they would not earn an additional
bonus. We randomized which of the two groups (High or Low Performers)
was selected by default, and Advisees were free to select the other
group. They also made a guess as to their score, and the survey
concluded with basic demographic questions.

We then recruited 1,000 participants for the role of Advisers, and the
focal part of our experiment. They were informed of the ability quiz
that the Advisees had completed, how the Low Performer Group and High
Performer Groups were constructed, as well as the choice and incentives
for the Advisees. We then asked Advisers to give advice to ten
participants on which group they should compete against. Advisers were
randomly assigned to one of two treatments. In the ``Performance''
treatment, they observed only the score of the other participant. In the
``Performance + Expectation'' treatment, they additionally observed the
advisee's non-binding choice of which group to compete against. However,
since the outcome was based solely on the past quiz score and advisers,
but not advisees, know the score, this initial choice was not
informative for the purpose of recommending competing against one group
or the other. Advisers could earn the same bonus as one of the
participants they had given advice to and who returned to make a
decision. The survey concluded with basic demographic questions.

Finally, we invited Advisees back for the follow-up survey. Following
our preregistration, we kept the survey open for 7 days. In total, 176
Advisees returned. The short survey reminded them of the task they
completed in the previous survey, informed them that other participants
from Prolific had observed their real score and given them advice on
which group to compete against, and finally reminded them of how many
questions they guessed they had answered correctly. They were not
informed of their true score or the score of the groups they could
compete against. Participants then observed the advice from a randomly
selected adviser and made their decision.

\subsection{Results}\label{results}

We begin by examining the performance of the Stage 1 participants.
Because our default treatment takes place after participants completed
the ability quiz, we do not expect a difference in performance across
the Low and High Default treatments. Indeed, the two groups scored no
different from one another (5.1 and 5.18 for the Low Default and High
Default treatments, respectively, \(t(199) = 0.31\), \(p = .756\)). The
default manipulation affected their initial choices. In the ``Low
Default'' treatment, only 22\% of participants initially preferred
competing against the High Performer Group, compared to 65\% of
participants who saw the High Performer Group highlighted (\(\chi^2\)(1,
n = 201) = 36.62, p \textless{} .001).

\begin{table}

{\caption{{Advice to compete against the High Performer Group in Study
1. Displaying the non-binding choice of those for whom the Low Performer
Group was selected by default makes it more likely that they are advised
to compete against the High Performer Group (Column 1). Column 2
restricts the analysis to advisers who observed advisees' initial
non-binding choice and controls for advisees' score on the quiz.
Standard errors are clustered at the adviser
level.}{\label{tbl-study1regs}}}
\vspace{-20pt}}

\centering
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={+ p \num{< 0.1}, * p \num{< 0.05}, ** p \num{< 0.01}, *** p \num{< 0.001}},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]},
column{2,3}={}{halign=c,},
column{1}={}{halign=l,},
hline{12}={1,2,3}{solid, black, 0.05em},
}                     %% tabularray inner close
\toprule
& (1) & (2) \\ \midrule %% TinyTableHeader
High Default               & \num{0.011}    & \num{0.025}*    \\
& (\num{0.013})  & (\num{0.011})   \\
Expectation Shown          & \num{-0.041}*  &                  \\
& (\num{0.016})  &                  \\
Expectation x High Default & \num{0.037}+   &                  \\
& (\num{0.019})  &                  \\
Performance                &                 & \num{0.168}***  \\
&                 & (\num{0.004})   \\
Constant                   & \num{0.389}*** & \num{-0.503}*** \\
& (\num{0.012})  & (\num{0.020})   \\
N                          & \num{10000}    & \num{5000}      \\
\bottomrule
\end{talltblr}

\end{table}

We now examine whether advisers took the advisees' expectations into
account. Recall that advisees who were defaulted to competing against
the High Performer Group were indeed more likely to choose this group,
which, from the advisers' perspective, could signal high confidence in
their abilities. Our theory predicts that showing expectations should
lead to more flattering advice--- specifically, a recommendation to
compete against the High Performer Group--- for advisees with a high
default choice compared to those with a low default choice. Column 1 of
Table~\ref{tbl-study1regs} reports a linear probability model with
advice to compete against the high group as the outcome measure,
incorporating the default choice advisees saw, whether expectations were
shown to the adviser, and the interaction of the two. Because each
adviser made ten recommendations, we cluster standard errors at the
adviser level. As predicted, we find a marginally significant
interaction effect: advisees with a default choice to compete against
the High Performer Group (nudged to have high expectations, though not
necessarily holding them) were more likely to receive advice to compete
against stronger competitors when their expectations were shown compared
to when they were hidden. We reported a linear probability model on
advising to compete against the High Performer group for advisers in the
``Expectations'' treatment, controlling for the true performance of the
advisee in Column 2 of Table~\ref{tbl-study1regs}. We observe a
significant main effect of the default treatment, where a high default
increased the likelihood of receiving flattering advice. It is worth
noting that, since only 65 \% of advisees in the high default group
actually stuck with their default choice, this represents a conservative
test of the effect of high expectations on the likelihood of receiving
flattering advice.

\begin{figure}

\caption{\label{fig-study1Advice}Advisees who were nudged to express
high expectations were more likely to receive advice to compete against
the High Performer Group.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-study1Advice-1.pdf}}

}

\end{figure}%

\subsection{Discussion}\label{discussion}

When the quality of a decision does not depend on someone's
expectations, the adviser's awareness of the advisee's expectations
should not influence the advice given. However, in line with our
argument that advisers take into account the belief utility of the
advisee and prefer to flatter rather than disappoint, we find that
displaying the advisee's expectations does make a difference.
Specifically, advisees who expressed a preference for stronger
competitors were more likely to receive `flattering' advice, encouraging
them to pursue that option. One possible explanation for this finding is
that, rather than trying to flatter, advisers may simply be telling
advisees what they want to hear. In the next study, we aim to
investigate the mechanism behind this discrepancy in advice-giving when
expectations are revealed, exploring the question: Why do people give
flattering advice?

\section{Study 2}\label{study-2}

So far, we have shown that displaying expectations influences the advice
given. Specifically, advisers were more likely to recommend competing
against a group of high performers to advisees with higher performance
expectations compared to those without. We now directly test whether
this effect is driven by interpersonal concerns. In particular, we
explore whether more flattering advice stems from a desire to be liked,
and whether advisees, in turn, like advisers more when they receive
flattering advice.

We conducted a three-stage experiment in which we experimentally
manipulated the incentives for advisers. We began by inviting a group of
advisees to upload photos of themselves (``selfies'') and informed them
that they would be rated on their attractiveness. We grouped them with
nine other participants of the same gender and recruited participants of
the opposing gender to rank them from most to least attractive and to
provide advice. Specifically, we asked them to advise the participant
they ranked as the 7th most attractive (4th least attractive) on what
rank they should bet they were ranked by a larger group of raters.
Advisers were randomly assigned to two treatments, receiving a bonus
payment either if the advisee guessed their rank accurately or if the
advisee evaluated the adviser as likeable as measured by an
unincentivized scale response. We hypothesize that advisers who want to
be liked by advisees will recommend that they bet on a lower rank,
i.e.~that they are more attractive.

\subsection{Methods}\label{methods-1}

We recruited 300 participants from Prolific and, after asking
demographic questions, invited them to upload photos of themselves
(selfies) to be rated by other participants on attractiveness. 100 men
and 107 women agreed to do so and uploaded pictures that adhered to our
instructions (e.g., did not include other people). We selected the first
100 photos from women to arrive at a gender-balanced sample
(\(M_{\text{Age}}\) = 39.37 years). Participants were informed that
their selfies would be randomly grouped with those of nine other
participants of their gender and ranked in terms of attractiveness by a
group of new Prolific participants of the opposite gender. Lastly, they
guessed their rank (unincentivized).

Next, we recruited a new, gender-balanced sample for the role of
advisers. 472 participants from Prolific (\(M_{\text{Age}}\) =
41.0275424 years; 49.79\% Female) started by providing demographic
information, then were matched to a group of the opposite gender. They
then ranked participants from most to least attractive using a drop-down
menu next to each picture. Because of a limitation with the survey
software, participants were able to select the same rank for multiple
participants, rather than rank them uniquely. We remove 115 participants
who did not follow instructions and provide a unique ranking.

After submitting their ratings, they saw the photo of the participant
they had ranked as the 7th most attractive (i.e., the 4th least
attractive). They were reminded of the rank they had given to that
person and informed that this participant would be invited back and
could earn a \$1 bonus if they guessed their rank correctly. The rank
was determined by the aggregate ratings of all participants who had
ranked this group. Their task was to give advice to this participant
about the rank they should bet on based on having observed all ten
selfies and their own assessment. We randomly assigned participants to
one of two incentivization schemes. In the ``Accuracy'' treatment, they
received a bonus identical to the advisee: \$1 if they guessed their
rank correctly. In the ``Likeability'' treatment, we informed them that
they would be rated by the advisee on a 5-point Likert scale for how
likeable they thought they were. Each point on the scale would translate
to a bonus of 20 cents. They then selected a rank that they would
recommend the advisee to bet on.

Finally, we invited participants from Stage 1 and were ranked as the 7th
attractive by at least one adviser (so that they received advice) back
for the follow-up survey. Following our preregistration, we kept the
survey open for 7 days. In total, 146 participants (77 men, 69 women)
returned. They were reminded of the selfie they uploaded in Stage 1 and
informed that a group of 10 selfies, including theirs, had been rated by
other participants from Prolific. They then saw the advice from a
randomly selected adviser and made their estimate with a \$1 incentive
for guessing accurately. They then saw the rank they had been advised to
bet on one more time and were asked to rate the adviser's likability,
warmth, friendliness, good-naturedness, trustworthiness, and sincerity
on 5-point Likert scales (adapted from Fiske et al. (2007)).

\subsection{Results}\label{results-1}

\begin{figure}

\caption{\label{fig-study2advice}Advisers were asked to suggest what
rank to bet on for someone they ranked as the 7th most attractive out of
ten people. When incentivized for accuracy, most of them advised betting
on 7, which was based on their evaluation of the recipient's
attractiveness. However, when incentivized for likability, many
participants gave flattering advice by suggesting betting on higher
ranks that imply a higher evaluation of attractiveness.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-study2advice-1.pdf}}

}

\end{figure}%

We begin by examining the prior beliefs of advisees who uploaded their
selfies. On average, men guessed they ranked 5.76 in their group of 10
and women guessed that they ranked 6.37. Participants overall
underestimate their attractiveness relative to the benchmark average of
5.5. Moreover, women do so more than men \(t(205) = 2.15\),
\(p = .033\). Notably, people's self-perceptions correlated strongly
with the aggregate ratings of the advisers (0.43, \(t(193) = 6.70\),
\(p < .001\)). However, there was substantial heterogeneity in
perceptions of attractiveness. Of the 200 participants, 149 were ranked
as 7th most attractive by at least one adviser. On average, men in this
subset estimated they were ranked 5.94th and women estimated they were
ranked 6.59th (\(t(143) = 1.94\), \(p = .054\)).

Next, we turn our attention to the advisers
(Figure~\ref{fig-study2advice}). In the Accuracy condition, those
uploading selfies were advised to bet on rank 6.19. Notably, this is
significantly more attractive than the 7th rank those advisers had
themselves guessed just on the prior screen (\(t(234) = -8.82\),
\(p < .001\)). This suggests that even when incentivized for accuracy,
participants offered flattering advice.\footnote{This shading could be
  due to concerns of avoiding disappointment. However, it could also be
  that advisers are uncertain about the rankings they had given and make
  a recommendation that combines their own belief with a uniform prior.}
Importantly, and as predicted, we find that advisers in the Likeable
treatment recommend betting on a lower rank, communicating that they
think the participant in the selfie is more attractive (5.38,
\(t(470) = 5.44\), \(p < .001\)). This suggests that advisers inferred
that advising someone that they were more attractive would make the
adviser appear more likeable and therefore provided advice that
communicated a more favorable impression of the participants'
attractiveness. The distribution shown in Figure~\ref{fig-study2advice}
shows that participants do not simply tell participants that they are
the most attractive person in the group. They may infer that flattering
advise needs to be somewhat realistic to be believable. We return to
this in the general discussion.

\begin{table}

{\caption{{When individuals receive advice that implies a high level of
attractiveness (lower rank), they tend to perceive the advice giver as
more likable (Column 1), warm (Column 2). Column 3 shows that advisors
are rated as more trustworthy when they advice lower ranks, but this
relationship is only directional.}{\label{tbl-study2regs}}}
\vspace{-20pt}}

\centering
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={+ p \num{< 0.1}, * p \num{< 0.05}, ** p \num{< 0.01}, *** p \num{< 0.001}},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]},
column{2,3,4}={}{halign=c,},
column{1}={}{halign=l,},
hline{6}={1,2,3,4}{solid, black, 0.05em},
}                     %% tabularray inner close
\toprule
& (1) & (2) & (3) \\ \midrule %% TinyTableHeader
Advised Rank & \num{-0.109}*  & \num{-0.133}** & \num{-0.054}   \\
& (\num{0.046})  & (\num{0.044})  & (\num{0.042})  \\
Constant     & \num{3.777}*** & \num{3.903}*** & \num{3.397}*** \\
& (\num{0.279})  & (\num{0.266})  & (\num{0.257})  \\
N            & \num{146}      & \num{146}      & \num{146}      \\
\bottomrule
\end{talltblr}

\end{table}

Finally, we examine whether flattering advice indeed leads to more
positive evaluations of advisers, or whether flattering advice is
dismissed as insencere. Following our preregistration, we average the
ratings on likeability, warmth, friendliness, and good-naturedness to
create a scale of likeability (\(\alpha\) = 0.9243334); and we create a
scale of trustworthiness by averaging the ratings of trustworthiness and
sincerity (\(\alpha\) = 0.8740347).

\begin{figure}

\caption{\label{fig-study2rating}Participants tend to rate advisors who
suggest betting on a higher rank (implying greater attractiveness) as
more likable and warm.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-study2rating-1.pdf}}

}

\end{figure}%

\begin{figure}

\caption{\label{fig-study2ratingByGroup}After participants received
advice that suggested betting on a rank that was lower than, equal to,
or higher than their self-expectations, they evaluated advisors based on
their likability and warmth. Participants rated advisors who gave
flattering advice and delivered a high evaluation of their
attractiveness as more likable and warm than those whose advice implied
a lower evaluation of their attractiveness.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-study2ratingByGroup-1.pdf}}

}

\end{figure}%

As seen in Figure~\ref{fig-study2rating}, we found that advisees who
suggested that the advisee was more attractive (lower rank) were indeed
rated as more likeable and warm (b = -0.109, p \textless{} .05; b =
-0.133, p \textless{} .001, respectively; See Columns 1 and 2 of
Table~\ref{tbl-study2regs}). When using their self-expectations as a
reference point, advisors who suggest a rank lower than that point are
perceived as less likable and warm (see
Figure~\ref{fig-study2ratingByGroup}). Interestingly, these benefits are
not at the cost of being hypocritical; advisers who recommend a more
favorable rank are viewed as no less trustworthy (Column 3 of
Table~\ref{tbl-study2regs}).

We were not powered to do a comparison across the two experimental
groups and did not preregister such a difference. Indeed, we find no
difference in likeability and warmth across the two treatments
(\(t(144) = 1.09\), \(p = .275\), and \(t(144) = 0.58\), \(p = .562\),
respectively). We also assess the quality of advice by measuring the
discrepancy between the advised rank and the final rank. Our analysis
revealed no significant difference between the two treatments in terms
of how helpful the advice was for advisees in making accurate guesses,
\(t(144) = -0.29\), \(p = .776\).

\subsection{Discussion}\label{discussion-1}

In another context where advice communicates ego-relevant information
(here, people's attractiveness), we find that the advice people give is
contingent on their incentives. Specifically, when they get rewarded for
being more likeable, they recommend that the advisee bet on a more
favorable rank than when they are incentivized for accuracy.
Importantly, advisees do not discount flattering advice and instead
evaluate people who advise them to bet on a more attractive rank as
warmer and more likeable. These gains to interpersonal perceptions do
not come at the cost of trustworthiness, even when the advice is
substantially inflated (e.g., a recommendation that the advisee is the
most attractive in the group).

\section{Study 3}\label{study-3}

In Study 3, we replicate the main effect in a different ego-relevant
domain: intelligence. The experimental design is conceptually identical
to Study 1, but this time we utilize a more direct measure of
expectation by asking participants to estimate their scores.

This three-stage experiment begins with measuring actual performance.
Participants are then anchored to low or high performance expectations
and asked to guess how many questions they answered correctly. As in
previous studies, we recruit advisers who are presented with either just
the advisee's true performance or both the true performance and the
number of questions the advisee believes they answered correctly.
Advisers are then asked to provide advice on choosing a competitor.
Finally, we present the advice to the advisees and ask them to make
their final choice. Our key hypotheses are: (1) advisees anchored to
high expectations will be more likely to receive advice to compete
against the high-performing group, and (2) men will be more likely than
women to be advised to compete against the high-performing group when
expectations are disclosed.

\subsection{Methods}\label{methods-2}

We begin by first recruiting a sample of 50 participants from Prolific
to complete a 10-question multiple choice mathematics quiz. The
questions were taken from a paper-version of the ASVAB standardized
exam, such that answers were not available online. Participants had five
minutes to answer the quiz and were paid 10 cents for each correctly
answered question. Like before, we define top 20 scores as ``High
Performers'' and the bottom 20 scorers as the ``Low Performers.'' On
average, participants answered 4.42 questions correctly, High Performers
scored between 5 and 10, and Low Performers scored between 0 and 3. In
order to anchor the expectations of participants in our main experiment,
we simulated 1,000 pairings of groups of five participants, with the 5th
percentile of groups scoring an average of 2.6 and the 95 th percentile
scoring an average of 6.4. We report these averages to participants in
the Low Expectations and High Expectations treatment, respectively.

We then recruited 1,002 participants for the role of advisees in Stage 1
of our main experiment. To arrive at a gender-balanced sample, we
dropped the last two male participants to complete the survey, ending up
with a sample of 500 men and 500 women (\(M_{\text{Age}}\) = 42.16).
Participants completed the same 10-item mathematics quiz as the earlier
participants and were informed that their performance would affect their
bonus earnings in a follow-up stage to be conducted a few days later.
After completing the quiz, we informed them of the average score of a
group of five participants from the preliminary survey. We randomly
assigned them to learn about the 5th percentile of groups, which scored
2.6. (``Low Expectations'' treatment) or the 95th percentile of groups,
which scored 6.4 (``High Expectations'' treatment). Participants then
made a guess (unincentivized) about how many questions they think they
answered correctly. The survey concluded with basic demographic
questions.

We then recruited 1,000 participants for Stage 2, placing them in the
role of advisers. We began by informing them of the mathematics quiz
that participants in the preliminary study and Stage 1 had completed,
and informed them of the average score of all participants in the
preliminary study. Advisers had to recommend whether an advisee should
compete against the Low Performers or the High Performers (we used these
terms in the survey). We anticipated that being told to compete against
High Performers is more flattering and hence being told to compete
against the Low Performers would be disappointing if one had expected to
do well. Advisees could earn a bonus if their score was equal to or
higher than that of a randomly selected member from their chosen group.
Competing against the High Performers was more challenging, resulting in
an average bonus of 50 cents, compared to 30 cents when competing
against other groups.

Like before, advisers were randomly assigned to one of two treatments.
In the ``Baseline'' treatment, they only observed the score of the
advisee on the mathematics quiz. In the ``Expectation'' treatment, they
observed the score as well as the advisee's guess for how many questions
they answered correctly. We want to emphasize that sice the outcome is
determined only by the past score on the quiz, the advisee's guess is
immaterial to which group they should compete against. Moreover, in
neither treatment did they receive any demographic information about
their advisee. Participants gave recommendations to 10 advisees, which
unbeknownst to them were five men and five women matched to have
identical performance on the test.\footnote{We made this decision to
  account for the possibility of gender differences in performance.}
They were informed that if their advice was shown to a participant who
returned for the follow-up survey, they would receive the identical
bonus as that participant. The survey then concluded with basic
demographic questions.

Finally, we invited participants from Stage 1 back for the follow-up
survey. Following our preregistration, we kept the survey open for 7
days. In total, 951 participants (481 men, 470 women) returned. The
brief survey reminded them of the task they completed in Stage 1,
informed them that other participants from Prolific had observed their
real score and given them advice against which group to compete against,
and finally were reminded them of how many questions they guessed they
had answered correctly. Importantly, they were not informed of their
true score or the score of the groups they could compete against.
Participants then observed the advice from a randomly selected adviser
and made their decision.

\subsection{Results}\label{results-2}

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants completed the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups scored
no different from one another (4.79 and 4.55 for the Low Expectations
and High Expectations treatments, respectively, \(t(998) = -1.64\),
\(p = .102\)). The expectations treatment did, however, affect how well
they thought they performed. Participants in the ``Low Expectations''
treatment guessed a score of 4.79 vs.~4.42 in the High Expectations
treatment (\(t(998) = 3.78\), \(p < .001\)), showing that the
manipulation was successful, albeit small. Contrary to our expectations,
we did find a gender difference in performance: men scored 5.02 on
average, while women scored 4.33 (\(t(998) = -4.73\), \(p < .001\)).
Consistent with this difference, men thought they answered more
questions correctly than did women (4.75 vs 3.56, \(t(998) = -8.60\),
\(p < .001\)). This difference, however, does not affect the
interpretation of our findings, which will rely on an interaction of
gender with an experimental treatment for Stage 2 participants.

We define a measure of ``overconfidence'' as (Performance - Estimate).
We observe that women underestimate their performance by 0.76 points,
while men do so by only 0.27 points \(\Delta M = -0.49\), 95\% CI
\([-0.73, -0.25]\), \(t(997.87) = -4.03\), \(p < .001\). Although
neither gender is overconfident, men on average are more confident in
their performance than are women, as we had expected.

\begin{table}

{\caption{{Column 1 displays the advice given to compete against a High
Performance group based on whether the advisees were primed with low or
high expectations and the expectation level shown to the advisor. Column
2 displays the advice given to compete against a High Performance group
based on the gender of the targets and the expectation level shown to
the advisor. Column 3 displays the expected bonus of the advice received
based on the gender of the targets and the expectation level shown to
the advisor. All standard errors clustered at the level of the
advisor.}{\label{tbl-study3regs}}}
\vspace{-20pt}}

\centering
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={+ p \num{< 0.1}, * p \num{< 0.05}, ** p \num{< 0.01}, *** p \num{< 0.001}},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]},
column{2,3,4}={}{halign=c,},
column{1}={}{halign=l,},
hline{14}={1,2,3,4}{solid, black, 0.05em},
}                     %% tabularray inner close
\toprule
& (1) & (2) & (3) \\ \midrule %% TinyTableHeader
High Expectation   & \num{0.012}     &                 &                 \\
& (\num{0.011})   &                 &                 \\
Performance        & \num{0.128}***  &                 &                 \\
& (\num{0.003})   &                 &                 \\
Expectation Shown  &                  & \num{-0.034}*  & \num{0.000}    \\
&                  & (\num{0.013})  & (\num{0.003})  \\
Advisee Male       &                  & \num{0.002}    & \num{0.000}    \\
&                  & (\num{0.005})  & (\num{0.001})  \\
Expectation x Male &                  & \num{0.032}*** & \num{-0.005}** \\
&                  & (\num{0.009})  & (\num{0.002})  \\
Constant           & \num{-0.263}*** & \num{0.339}*** & \num{0.258}*** \\
& (\num{0.015})   & (\num{0.010})  & (\num{0.002})  \\
N                  & \num{5000}      & \num{10000}    & \num{10000}    \\
\bottomrule
\end{talltblr}

\end{table}

\begin{figure}

\caption{\label{fig-study3genderdiff}Advisers observed real performance
and the participants' estimated performance, but not their gender. As a
result of their higher expectations, men (blue) were advised to compete
against the high performance group more often than women (red). The grey
line shows advice absent expectations, which did not differ by gender.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-study3genderdiff-1.pdf}}

}

\end{figure}%

Next, we examine whether advisers took the advisee's expectations into
account. Column 1 of Table~\ref{tbl-study3regs} reports a linear
probability model on advising to compete against the High Performer
group for advisers in the ``Expectations'' treatment, controlling for
the true performance of the advisee. Because each adviser made ten
recommendations, we cluster standard errors at the adviser level.
However, contrary to our expectations, we do not see a significant
effect of the expectations treatment. This may be because the induced
difference in expectations was too small.

However, recall that men were more confident in their performance than
were women. Our theory thus predicts that showing expectations should
lead to more flattering advice (i.e., a recommendation to compete
against the High Performer group) for men than for women. We report a
linear probability model with advice to compete against the high group
as the outcome measure, and the advisee's gender, whether expectations
were shown to the adviser, and the interaction of the two in Column 2 of
Table~\ref{tbl-study3regs}. As predicted, we find a significant
interaction effect: men are more likely to be advised to compete against
the High Performers when expectations are shown than when they are
hidden (\(p < 0.001\)). We show this result graphically in
Figure~\ref{fig-study3genderdiff}.

To examine the quality of advice, we computed the expected bonus
earnings for someone who followed the recommendations. For example, if
an adviser suggested competing against the High Performer group, we
matched the advisee against all 20 members of that group and determined
how often their score matched or exceeded that of the member. We then
multiplied this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of
Table~\ref{tbl-study3regs}. Displaying expectations led men to be
advised to compete against the High Performer group more often, and this
advise turned out to be bad: men receive worse advise than do women when
expectations are displayed, but not in their absence.\footnote{This
  analysis was not preregistered, and we note here that the reduction in
  expected earnings is small. However, it is interesting that
  expectations have a negative effect for men who underestimate their
  performance on average. One possibility is that advisers suggest the
  High Performer group more often than is optimal. We return to the
  possibility that advice is overall biased to be flattering in Study 2.}

\begin{table}

{\caption{{Column 1 displays the actual bonus based on the performance
and whether the participant is primed with high expectations. Column 2
displays the actual bonus based on the performance and gender of the
participant. Columns 3 and 4 display the chance of adopting the advice
based on the gender of the participants, whether the advisee sees
expectations, and whether the advice is to compete against high
performers. The former considers only the main effect, while the latter
also includes the interactive effect.}{\label{tbl-study3time3}}}
\vspace{-20pt}}

\centering
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={+ p \num{< 0.1}, * p \num{< 0.05}, ** p \num{< 0.01}, *** p \num{< 0.001}},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]Q[]},
column{2,3,4,5}={}{halign=c,},
column{1}={}{halign=l,},
hline{20}={1,2,3,4,5}{solid, black, 0.05em},
}                     %% tabularray inner close
\toprule
& (1) & (2) & (3) & (4) \\ \midrule %% TinyTableHeader
Performance                          & \num{0.038}*** & \num{0.039}*** &                  &                  \\
& (\num{0.003})  & (\num{0.003})  &                  &                  \\
High Expectation                     & \num{-0.020}   &                 &                  &                  \\
& (\num{0.013})  &                 &                  &                  \\
Advisee Male                         &                 & \num{-0.002}   & \num{-0.045}+   & \num{-0.144}*** \\
&                 & (\num{0.013})  & (\num{0.026})   & (\num{0.041})   \\
Expectation x Male                   &                 &                 &                  & \num{0.113}*    \\
&                 &                 &                  & (\num{0.053})   \\
Expectation Shown                    &                 &                 & \num{-0.002}    & \num{-0.057}    \\
&                 &                 & (\num{0.026})   & (\num{0.040})   \\
Advice: High Performer               &                 &                 & \num{-0.138}*** & \num{-0.207}*** \\
&                 &                 & (\num{0.028})   & (\num{0.051})   \\
Advice: High Performer x Male        &                 &                 &                  & \num{0.131}*    \\
&                 &                 &                  & (\num{0.056})   \\
Expectation x Advice: High Performer &                 &                 &                  & \num{-0.009}    \\
&                 &                 &                  & (\num{0.056})   \\
Constant                             & \num{0.087}*** & \num{0.076}*** & \num{0.857}***  & \num{0.906}***  \\
& (\num{0.016})  & (\num{0.015})  & (\num{0.024})   & (\num{0.030})   \\
N                                    & \num{483}      & \num{483}      & \num{951}       & \num{951}       \\
\bottomrule
\end{talltblr}

\end{table}

To determine whether flattering advice is truly costly, however, we need
to examine the outcome of the advisees. In particular, they could ignore
flattering advice, recognizing it as such and thus failing to adhere to
it. Align with our prediction, participants who were primed with high
expectations earned less when their expectations were conveyed to the
advisers, although this result is only directional (see Column 1 of
Table~\ref{tbl-study3time3}). Similarly, as shown in Column 2 of
Table~\ref{tbl-study3time3}, we also found that male participants
received less reward when their expectations were presented, even when
their performance was the same. These findings suggest that flattering
advice is not without consequences. Furthermore, our research indicates
that because males are more likely to follow such advice (see Column 4
of Table~\ref{tbl-study3time3}), their tendency towards competition
intensifies the cost of such advice.

\subsection{Discussion}\label{discussion-2}

We found supportive evidence for our argument in a different domain.
Advisors tend to consider advisees' self-expectations and match their
advice accordingly, aiming to avoid disappointment. Consistent with the
confidence gap narrative, men tend to underestimate their mathematics
test scores less than women. When these expectations are displayed to
advisors, they are more likely to advise men to compete against high
performers. Notably, this turns out to be poor advice: men whose
advisors were aware of their expectations received worse advice. Our
findings suggest that men with the same scores as their female
counterparts ended up earning less, although this result is only
directional. This discrepancy may be due to men receiving more favorable
advice, being more likely to follow it, and ultimately facing worse
outcomes.

\section{General Discussion}\label{general-discussion}

Advice has the potential to shape people's career and personal outcomes.
Honest feedback, however, may be painful to learn if it falls short of
one's expectations. As prior work notes, this may motivate people to
avoid information and avoid seeking help (Bénabou et al., 2022; Golman
et al., 2017; Jaroszewicz et al., 2021). We present evidence from two
experiments that advisers are also cognizant of this cost. As a result,
they present flattering advice that avoids disappointing the recipient,
and correctly anticipate that this boosts how advisees perceive them.
However, this flattering advice comes at a cost to advisees, who would
do worse if they followed it blindly, as we show in Study 1.

Moreover, a desire to avoid disappointment also means that advisers have
to take into account the expectations of the advisee. We document that
men are more optimistic about their performance than women are. As a
result, they receive more flattering advice and are more likely to be
told to aim higher. Notably, in the context of our experiment, this
turns out to be bad advice ex post.

Our findings have implications for organizational practice, where
mentoring and advice-giving may take into account an employees'
expectation. We document this as a novel source of gender bias.
Organizations could reduce this bias by calibrating employees'
expectations to reduce overconfidence.

Participants in our experiment were paired anonymously. Even so, we
document this desire to avoid disappointment. We anticipate that advice
would be more flattering in face-to-face communication and without
anonymity. Moreover, existing relationships might make it even more
difficult for advisers to be honest.

In our experiments, participants only received advice once and did not
evaluate the adviser after observing the outcome of their decision. For
example, advice that leads to bad outcomes may undermine the
interpersonal benefits of flattery. Alternatively, people may still like
the flattering advice and not fault the adviser for the bad outcome.
Moreover, our setting involved only a single piece of advice on one
task. Future research could examine whether people return to those who
gave them flattering advice, or if they prefer someone who gave them the
honest (but unpleasant) truth.

Advice has long been studied from the perspective of the receiver.
Similarly, research on belief utility has examined how recipients
respond to the valence of the information they receive. Here, we show
that advisers, too, take into account the psychological impact of the
information they convey. They may have even greater motivations to avoid
conveying bad news, because they incur the interpersonal costs of
delivering unfavorable information without reaping the benefits from
helping someone make a better choice.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Benabou2022}
Bénabou, R., Jaroszewicz, A., \& Loewenstein, G. (2022). \emph{It
{Hurts} to {Ask}}. National Bureau of Economic Research.

\bibitem[\citeproctext]{ref-Fiske2007}
Fiske, S. T., Cuddy, A. J., \& Glick, P. (2007). Universal dimensions of
social cognition: Warmth and competence. \emph{Trends in Cognitive
Sciences}, \emph{11}(2), 77--83.

\bibitem[\citeproctext]{ref-Gino2007}
Gino, F., \& Moore, D. A. (2007). Effects of task difficulty on use of
advice. \emph{Journal of Behavioral Decision Making}, \emph{20}(1),
21--35.

\bibitem[\citeproctext]{ref-Golman2017}
Golman, R., Hagmann, D., \& Loewenstein, G. (2017). Information
avoidance. \emph{Journal of Economic Literature}, \emph{55}(1), 96--135.
\url{https://doi.org/10.1257/jel.20151245}

\bibitem[\citeproctext]{ref-Harvey1997}
Harvey, N., \& Fischer, I. (1997). Taking advice: {Accepting} help,
improving judgment, and sharing responsibility. \emph{Organizational
Behavior and Human Decision Processes}, \emph{70}(2), 117--133.

\bibitem[\citeproctext]{ref-Jaroszewicz2022}
Jaroszewicz, A., Loewenstein, G., \& Bénabou, R. (2021). \emph{The pain
of asking and being asked for informal help}.

\bibitem[\citeproctext]{ref-John2019}
John, L. K., Jeong, M., Gino, F., \& Huang, L. (2019). The
self-presentational consequences of upholding one's stance in spite of
the evidence. \emph{Organizational Behavior and Human Decision
Processes}, \emph{154}, 1--14.

\bibitem[\citeproctext]{ref-Loewenstein2018}
Loewenstein, G., \& Molnar, A. (2018). The renaissance of belief-based
utility in economics. \emph{Nature Human Behaviour}, \emph{2}(3),
166--167.

\bibitem[\citeproctext]{ref-Shalvi2019}
Shalvi, S., Soraperra, I., Weele, J. J. van der, \& Villeval, M. C.
(2019). \emph{Shooting the messenger? Supply and demand in markets for
willful ignorance}.

\bibitem[\citeproctext]{ref-Soll2009}
Soll, J. B., \& Larrick, R. P. (2009). Strategies for revising judgment:
How (and how well) people use others' opinions. \emph{Journal of
Experimental Psychology: Learning, Memory, and Cognition}, \emph{35}(3),
780.

\end{CSLReferences}

\end{document}