### 2025-01-28 manuscript/manuscript.qmd
Study 2

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle: "Flattering Advice"
date: yes
author: 
  - name: "Amanda Chen"
    corresponding: true
    orcid: 0000-0002-4428-627X
    email: zchengj@connect.ust.hk
    affiliations:
      - id: id1
        name: "The Hong Kong University of Science and Technology"
        department: "Department of Management"
        city: "Hong Kong"
  - name: "David Hagmann"
    orcid: 0000-0002-2080-997X
    affiliations:
    - id: id1
author-note:
  disclosures:
    gratitude: The authors are grateful to XXX for their helpful comments.
abstract: Good advice improves decision quality but often requires delivering unpleasant truths that may disappoint advisees. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers prioritize avoiding disappointment at the expense of accuracy and their own earnings. In Study 1, advisers financially rewarded for accuracy still tailor recommendations to aspirational goals expressed by advisees, resulting in worse advice. When incentivized to be liked, advisers provide even more flattering advice, and advisees reward this by rating these advisers as more likable, despite the advice being less honest and less accurate (Study 2). The desire to avoid disappointment may lead to inequities if there are differences in expectations across social groups. In Study 3, we examine a setting in which men expect to perform better than women. We show that advisers take into account these expectations, leading to systematically different advice for men and women even when their gender is unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making.

keywords: [Gender difference, Advice, Interpresonal relationship]
bibliography: bibliography.bib
floatsintext: true
mask: false
link-citations: false
format:
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    prefer-html: true
  apaquarto-docx:
    default: true
    prefer-html: true
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| include: false

library(magrittr)
library(ltm)
library(ggplot2)
library(gridExtra)
library(dplyr)

knitr::opts_chunk$set(fig.width = 8, fig.height = 4, cache = F, echo = F,
                      fig.align = 'center',warning = FALSE, message = FALSE)

study1_time1 <- readRDS(here::here("Data/Study1_time1.Rds"))
study1_ability50 <- read.csv(here::here("Data/Cleaned Data/S1_Ability50_clean.csv"))
study1_time2_giver <- readRDS(here::here("Data/Study1_time2_giver.Rds"))
study1_time2_advice <- readRDS(here::here("Data/Study1_time2_advice.Rds"))
study1_time3 <- readRDS(here::here("Data/Study1_time3.Rds"))

```

# Introduction

Career trajectories for men and women often differ substantially, even
among those with similar qualifications. One source of such differences
may be the advice that people receive. Prior work has documented that
men receive more aspirational advice and women more risk-averse advice,
which may explain why women are less likely to apply for and ultimately
obtain more rewarding positions [@Kanze2018]. Research on gender
discrimination has focused on the role of unconscious bias
[@Greenwald2009; @Greenwald1995], but interventions that try to reduce
this bias have been largely unsuccessful [@Chang2019; @Paluck2009].
Recent work has proposed cognitive, rather than unconscious, biases that
can lead to gender discrimination through the formation of false beliefs
[@Hagmann2024]. Here, we propose a novel source of differences in advice
by drawing on recent work on belief-based utility, expectation-based
reference points [@Koszegi2006; @Koszegi2009], and the hedonic
consequences of information [@Golman2017].

Advice figures prominently in workplace and personal decision-making,
shaping choices about promotions, career paths, and everyday dilemmas
[@Gino2007; @Harvey1997; @Soll2009]. Its fundamental premise is to
improve outcomes by enabling decision-makers to learn from others'
expertise. However, while honest advice can lead to better decisions, it
can also lead to disappointment when it suggests a desired outcome is
unlikely to materialize. Advisers may be motivated to offer flattering
recommendations that align with advisees' hopes, sparing them the
disappointment of a more sobering forecast.

Advice with the potential for disappointment appears in diverse
organizational settings: supervisors deliver feedback on employee
performance, lawyers convey likely outcomes of litigation to their
clients, and academics offer thoughts on colleagues' manuscripts.
Ideally, honest information would help people identify their strengths
and weaknesses, make informed decisions about whether to pursue legal
cases, and strengthen papers prior to submission to academic journals,
thus improving long-term outcomes. In practice, however, interpersonal
considerations may prevent candor. When someone's expectations are (too)
high, an adviser might hesitate to deliver information that undermines
it---particularly if doing so risks negative interpersonal
repercussions.

People often go to great lengths to avoid unpleasant information,
deriving belief-based utility from maintaining favorable views of
themselves [@Golman2017; @Ho2018; @Loewenstein2018]. Moreover, they
often prefer advisers who withhold bad news from them [@Shalvi2019] and
punish those who communicate bad news [@John2019]. Advisers,
anticipating these preferences and consequences face a dilemma: Should
they provide the accurate but potentially painful truth, or shade their
recommendations to preserve the advisee's mood and the adviser's own
standing? Advisers who fear they will be blamed or disliked for
undermining someone's confidence may reasonably avoid candid feedback,
particularly if they don't incur a cost when the advisee makes a
mistake.

In our experiments, however, advisees have no opportunity to punish the
adviser, and advisers' financial incentives are linked to the decisions
the advisees make. We propose that advisers nonetheless have reason to
provide flattering advise because they recognize the psychological toll
of disappointing news. As a result, they may similarly incur a hedonic
cost for delivering unpleasant information that they anticipate may
distress the advisee. Thus, even in the anonymized context of an online
experiment, we hypothesize (and find) that advisers are relucant to
disappoint advisees and advice is thus biased upward, away from accuracy
and toward confirming advisees' priors. We show that this reduces the
quality of advice—but that advisees nonetheless view flattering advisers
more favorably.

We present the results of three preregistered experiments in which
participants are paired anonymously as advisers and advisees. Advisers
are incentivized to offer accurate advice, with bonus earnings depending
on the outcome of an advisee's decision. We study a setting in which
advisees can communicate their expectations to an adviser, and
experimentally manipulate whether advisers observe this expectation. Our
design simulates organizational contexts where mentors and managers
often have some sense of an employee's confidence or aspirations and
thus could adjust their feedback to avoid disappointing them.
Interpersonal considerations would likely be stronger in situations in
which the adviser and advisee have an existing relationship and when the
advice is delivered face-to-face. Moreover, advisers usually do not
suffer any direct costs when their flattering advice leads to a poor
outcome for the recipient, particularly when it is not clear what would
have happened under a counterfactual.\^\[In some cases, however,
repeated interaction might also increase incentives for honest feedback.
Someone who is known to persistently give overoptimistic advice may be
viewed as less trustworthy in the long run.\]

In Study 1, we nudge advisees into a preference for competing against
either a group of top-performers or a group of low-performers, through
the use of a default option. Advisers who observe the actual performance
of advisees and therefore have the relevant information needed to make a
recommendation nonetheless take into account the initial preference when
recommending which group they should compete against. In Study 2, we ask
advisees to upload photos of themselves and advisers rank them on their
attractiveness. Advisers then recommend which rank the advisee should
bet on, receiving an incentive either for the advisee's accuracy or for
being evaluated favorably. Both groups of advisers recommend a rank that
is more attractive than what they themselves have evaluated the advisee,
and those incentivized for likability further inflate the ranking.
Advisees evaluate advisers more favorably when they recommend betting on
a more attractive rank, including viewing them as more trustworthy.

In Study 3, we examine a setting in which participants report their
expected performance on a mathematics quiz. Advisers observe the
test-taker's true score and, in one treatment, the score they guessed
they would receive. They then recommend whether the advisee should
compete against a group of high performers or low performers.
Importantly, this "competition" is based on the past performance
observed by the adviser and thus the advisee's expectation does not
provide instrumental information. However, we find that (1) men expect
to perform better than women, given identical performance, and (2)
advisers take these expectations into account. As a result, when
expectations (but not gender) are known to advisers, the advice given to
men and women differs, and men are more often advised to compete against
the group of high performers. Notably, this leads to worse advice for
men than for women.

# Open Science Statement

We report all manipulations, measures, and data exclusion in our
experiments. The preregistration reports, screenshots of all
experimental materials, and the analysis code to replicate all
statistical analyses and figures are available on the Open Science
Framework
(<https://osf.io/8r3d4/?view_only=5ad7bafcd16b4d4ba08bb28b0e2bd02d>).

# Study 1

We first examine whether advisers take into account the expectations of
advisees when providing a recommendation. A group of participants
("Advisees") complete a quiz consisting of ten questions that draw on
ego-relevant domains. They then express a non-binding preference to
compete against a group of high performers or low performers on the same
task, based on their past performance. We nudge participants towards
picking either the low or the high performer group by selecting one of
the two by default. A second group of participants ("Advisers") observe
Advisees' score on the quiz and, in one treatment, also which group they
had selected. We predicted that Advisees who were nudged toward the high
performer group would be more likely to receive advice to compete
against this high group when Advisers observed their non-binding choice.

## Methods

In a preliminary stage, we recruited
`r scales::comma(dim(study1_ability50)[1])` participants from Prolific
and gave them five minutes to complete a quiz consisting of ten items.
The quiz included word puzzles, identifying emotions from photos, and
selecting the best responses for hypothetical scenarios. These questions
were adapted from surveys that measure problem-solving, emotional
intelligence, and communication skills---which we pre-tested as being
important in modern society and hence where participants might have a
stake in doing well. Participants received a bonus of five cents for
each correctly answered question. We then ranked them based on their
score and label the 20 participants with the lowest (highest) scores as
the Low (High) Performer Group.

Next, we recruited `r scales::comma(dim(study1_time1)[1])` participants
for the role of Advisees. They completed the same 10-item quiz, also
earning five cents for each correct answer. We then asked them to
express a non-binding preference for whether they would like to compete
against the High or Low Performer Group. We informed them that they
would be invited back at a later date when they could make a binding
decision and they could earn a bonus based on whether their score on
this quiz was equal to or higher than a randomly selected member from
their comparison group. If they picked the High Performer Group and had
an equal or higher score than a randomly selected member of that group,
they would earn a bonus of 50 cents. If they picked the Low Performer
Group and outperformed a randomly selected member of that group, they
could earn a bonus of 20 cents. If their score was lower in this
comparison, they would not earn an additional bonus. We randomized which
of the two groups (High or Low Performers) was selected by default, and
Advisees were free to select the other group. They also made a guess as
to their score, and the survey concluded with basic demographic
questions.

We then recruited `r scales::comma(dim(study1_time2_giver)[1])`
participants for the role of Advisers, and the focal part of our
experiment. They were informed of the ability quiz that the Advisees had
completed, how the Low Performer Group and High Performer Groups were
constructed, as well as the choice and incentives for the Advisees. We
then asked Advisers to give advice to ten participants on which group
they should compete against. Advisers were randomly assigned to one of
two treatments. In the "Performance" treatment, they observed only the
score of the other participant. In the "Performance + Expectation"
treatment, they additionally observed the advisee's non-binding choice
of which group to compete against. However, since the outcome was based
solely on the past quiz score and advisers, but not advisees, know the
score, this initial choice was not informative for the purpose of
recommending competing against one group or the other. Advisers could
earn the same bonus as one of the participants they had given advice to
and who returned to make a decision. The survey concluded with basic
demographic questions.

Finally, we invited Advisees back for the follow-up survey. Following
our preregistration, we kept the survey open for 7 days. In total,
`r scales::comma(dim(study1_time3)[1])` Advisees returned. The short
survey reminded them of the task they completed in the previous survey,
informed them that other participants from Prolific had observed their
real score and given them advice on which group to compete against, and
finally reminded them of how many questions they guessed they had
answered correctly. They were not informed of their true score or the
score of the groups they could compete against. Participants then
observed the advice from a randomly selected adviser and made their
decision.

## Results

```{r}
contingency_table <- table(study1_time1$Guess, study1_time1$default)

chi_test <- chisq.test(contingency_table)

```

We begin by examining the performance of the Stage 1 participants.
Because our default treatment takes place after participants completed
the ability quiz, we do not expect a difference in performance across
the Low and High Default treatments. Indeed, the two groups score no
different from one another
(`r papaja::printnum(round(mean(study1_time1[study1_time1$default == "Low",]$Score),2))`
and
`r round(mean(study1_time1[study1_time1$default == "High",]$Score),2)`
for the Low Default and High Default treatments, respectively,
`r papaja::apa_print(t.test(Score ~ default, data = study1_time1, var.equal = T))$statistic`).
As expected, the default manipulation changes their initial choices. In the "Low
Default" treatment, only
`r round(mean(study1_time1[study1_time1$default == "Low",]$Guess),2)*100`%
of participants prefer to compete against the High Performer
Group, compared to
`r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`%
of participants in the "High Default" treatment
($\chi^2$(1, n = 201) = 36.62, p \< .001).

```{r}
#| label: tbl-study1regs
#| tbl-cap: "Advice to compete against the High Performer Group in Study 1. Displaying the non-binding choice of those for whom the Low Performer Group was selected by default makes it less likely that they are advised to compete against the High Performer Group (Column 1). Column 2 restricts the analysis to advisers who observe advisees' initial non-binding choice and controls for advisees' score on the quiz. Standard errors are clustered at the adviser level."

study1_advice <- study1_time2_advice %>%
  dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E', 1, 0))

modelsummary::modelsummary(
  list(`All Advice` = lm(Advice ~ ExpShow * Treatment, data = study1_advice), 
       `Expectations Shown` = lm(Advice ~ Treatment + Score, data = study1_advice %>% dplyr::filter(ExpShow == 1))),
  vcov = ~GiverID,
  coef_map = c(
    "TreatmentHigh" = "High Default",
    "ExpShow" = "Expectation Shown",
    "ExpShow:TreatmentHigh" = "Expectation x High Default",
    "Score" = "Score",
    "(Intercept)" = "Constant"
  ),
  stars = TRUE,
  gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0))
) %>%
  tinytable::group_tt(
   j = list("Advice to Compete Against High Performers" = 2:3))

```

We now examine whether advisers take the advisees' expectations into
account. Advisers observe the true performance of the advisee, such that
expectations about performance add no additional information. However,
because participants who score higher are also more likely to pick the
High Performer Group, we use our Default treatment as an instrument that reflect
higher expectations.
We show linear probability models for the decision of the Advisers, which is
equal to 1 if they recommend competing against the High Performer Group and
0 if they recommend the Low Performer Group. In Column 1 of @tbl-study1regs,
we show our main specificiation with the experimental assignment for Advisees
(High Performer Default vs. Low Performer Default) and Advisers
(Expectation Shown vs. Expectation Hidden), as well as their interaction.
Because each adviser made ten
recommendations, we cluster standard errors at the adviser level. We find that
displaying expectations of those in the Low Performer Default reduces recommendations
to compete against the Top group by 4 percentage points, or about 10\% relative
to when expectations are hidden. This decrease is marginally smaller for advisees
with a default choice to compete against the High Performer Group
(nudged to have high expectations, though not necessarily holding them), and whose
advice did not differ based on whether their expectations were shown or hidden.
We reported a linear probability model on advising to
compete against the High Performer group for advisers in the
"Expectations" treatment, controlling for the true performance of the
advisee in Column 2 of @tbl-study1regs. We observe a significant main
effect of the default treatment, where a high default increased the
likelihood of receiving flattering advice. It is worth noting that,
since only
`r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their
default choice, this represents a conservative test of the effect of
high expectations on the likelihood of receiving flattering advice.

```{r}
#| label: fig-study1Advice
#| fig-cap: "Advice given in Study 1. Advisees who were nudged to express high expectations were more likely to receive advice to compete against the High Performer Group."

study1_time2_advice %>%
  mutate(Treatment=factor(Treatment, levels=c('High','Low'), labels = c('High default', 'Low default'))) %>%
  group_by(Condition, Treatment) %>%
  summarize(avg_count = mean(Advice), ci = qnorm(0.975)*sd(Advice)/sqrt(dplyr::n())) %>%
  ggplot2::ggplot(ggplot2::aes(x = Condition, y = avg_count, fill = Treatment)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = avg_count - ci, ymax = avg_count + ci), 
                         position = ggplot2::position_dodge(0.9), linewidth = 0.2, width = 0.1) +
  # ggtitle("Advice: Meeting Expectations vs. Absent Expectations") +
  scale_fill_manual(values = c("#2A629A",'#FF7F3E')) +
  scale_y_continuous("P(Compete Against High Performer)",
                     labels = scales::percent) +
  geom_text(aes(x = Condition, y = 0.02,
                label = paste0(round(avg_count*100), "%")),
            position = position_dodge(0.9),
            color = "white") +
  papaja::theme_apa() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        axis.text.x = element_text(size = 12),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none", # No legend
        plot.title = element_text(hjust = 0.5, size = 16)) +
  scale_x_discrete(labels = c("Expectation Shown", "Expectation Hidden")) +
   geom_text(aes(label=Treatment, y = -0.01), position = position_dodge(0.90))

```

## Discussion

When an adviser knows an outcome that an advisee can only estimate, learning
this estimate should not change the advice that is given. However, in line with our
argument that advisers take into account the belief utility of the
advisee and prefer to avoid disappointing them, we find that providing information
about expectations makes a difference.
Specifically, advisees who expressed a preference for stronger
competitors were more likely to receive flattering advice, encouraging
them to pursue that option. This occurs even though advisers are incentivized
to give good advice.

# Study 2

We next examine if advisers believe that flattering advice will improve
how they are perceived, and if so, whether this perception is true. We do this
by manipulating the incentives for advisers, who receive a bonus either based
on how they are evaluated by an advisee, or a bonus for the advisee's accuracy.

The study takes place in three stages. First, a group of
advisees upload photos of themselves ("selfies") and are grouped with 
nine other participants of the same sex We then recruit participants of
the opposing sex to rank them from most to least attractive and to
provide advice. Specifically, they advise the participant
they ranked as the 7th most attractive (i.e., 4th least attractive) on what
rank they should bet they were ranked by a larger group of raters.
Advisers were randomly assigned to two treatments, receiving a bonus
payment either if the advisee guessed their rank accurately or if the
advisee evaluated the adviser as likeable as measured by a scale response. 
We hypothesize that advisers incentivized to be liked will recommend betting on
a lower rank, i.e. that the advisee is more attractive.

```{r}
#| include: false

study2_time1 <- readRDS(here::here("Data/Study2_time1.Rds"))
study2_time1_200 <- readRDS(here::here("Data/Study2_time1_200.Rds"))
study2_time2 <- readRDS(here::here("Data/Study2_time2.Rds"))
study2_finalRank <- readRDS(here::here("Data/Study2_finalRank.Rds"))
study2_time3 <- readRDS(here::here("Data/Study2_time3.Rds"))

```

## Methods

We recruit 300 participants from Prolific and, after asking
demographic questions, invite them to upload photos of themselves
(selfies) to be rated by other participants on attractiveness.
We obtain selfies from `r scales::comma(dim(study2_time1[study2_time1$gender == 'Male',])[1])`
men and
`r scales::comma(dim(study2_time1[study2_time1$gender == 'Female',])[1])`
women adhering to our
instructions (e.g., did not include other people). In line with our preregistration,
we select the first 100 photos from women to arrive at a gender-balanced sample
($M_{\text{Age}}$ = `r mean(study2_time1_200$age)` years). Participants
are informed that their selfies will be randomly grouped with those of
nine other participants of their sex and ranked in terms of
attractiveness by a group of new Prolific participants of the opposite
sex. They then make an unincentivized guess of their rank.

Next, we recruit `r scales::comma(dim(study2_time2)[1])` participants from Prolific
for the role of advisers ($M_{\text{Age}}$ = `r mean(study2_time2$age)` years;
`r round(mean(study2_time2$gender == "Female")*100, 2)`% Female). We first
collect demographic information, then match them to a group of
the opposite sex. They rank selfies from ten participants from most to least
attractive using a drop-down menu next to each picture. Because of a
limitation with the survey software, we could not validate that each rank is
given only once, and we remove 115 participants who failed to follow instructions 
and did not provide a complete ranking.^[Moreover, we could not collect data from any participant who did not select any participant as the 7th most attractive participant, as that image would be shown on subsequent pages.]

After submitting their ratings, participants see the photo of the participant
they ranked as the 7th most attractive (i.e., the 4th least
attractive). We remind them of the rank they have just given to that
person and inform them that this participant will be invited back and
can earn a \$1 bonus if they correctly guess their rank. The rank
is determined by the aggregate ratings of all participants who have
ranked this group. Because the participant does not observe the other nine
people in the group, they would depend on the adviser's recommendation along with
their own assessment. We randomly assign advisers to
one of two incentivization schemes. In the "Accuracy" treatment, they
receive a bonus identical to the advisee: \$1 if they guess their
rank correctly. In the "Likeability" treatment, we inform them that
the advisee will rate them on a 5-point Likert scale on how
likeable they thought they are. Each point on the scale would translate
to a bonus of 20 cents. Advisers then select a rank (from 1 to 10) that they
recommend the advisee to bet on.

Finally, we invite participants from Stage 1 who were ranked as the 7th most
attractive participant by at least one adviser (so that they received advice) back
for the follow-up survey. Following our preregistration, we keep the
survey open for 7 days. In total,
`r scales::comma(dim(study2_time3)[1])` participants
(`r scales::comma(dim(study2_time3[study2_time3$gender == 'Male',])[1])`
men,
`r scales::comma(dim(study2_time3[study2_time3$gender == 'Female',])[1])`
women) complete the follow-up survey. We remind them of the selfie they uploaded in Stage
1 and that a group of 10 selfies, including theirs, had been
rated by other participants from Prolific. They then see the advice from
a randomly selected adviser and make their own guess with a \$1 incentive
for accuracy. Finally, they rate the adviser's likability,
warmth, friendliness, good-naturedness, trustworthiness, and sincerity
on 5-point Likert scales [adapted from @Fiske2007].

## Results

```{r}
#| include: false
library(dplyr)

TargetsUnique <- study2_time2 %>%
  dplyr::select(Target) %>%
  unique()

TargetsUnique <- merge(TargetsUnique,study2_finalRank, by.x = 'Target', by.y = 'QID7_Id')

```

```{r}
#| label: fig-study2advice
#| fig-cap: "Advisers were asked to suggest what rank to bet on for someone they ranked as the 7th most attractive out of ten people. When incentivized for accuracy, most of them advised betting on 7, which was based on their evaluation of the recipient's attractiveness. However, when incentivized for likability, many participants gave flattering advice by suggesting betting on higher ranks that imply a higher evaluation of attractiveness."
#| eval: false

library(ggplot2)
library(ggridges)

study2_time2 <- study2_time2 %>% dplyr::mutate(Advice = as.numeric(Advice))

ggplot(study2_time2, aes(x = Advice, y = Condition, fill = Condition)) +
    # Reduce vertical spacing with rel_min_height and scale
    geom_density_ridges(scale = 1.0, rel_min_height = 0.01, alpha = 0.7) +

    # Shading the left side of x = 7 in light grey
    geom_rect(aes(xmin = 7, xmax = Inf, ymin = -Inf, ymax = Inf), fill = "gray95", alpha = 0.01) +

    # Add the dashed line at x = 7
    geom_vline(xintercept = 7, col = 'black', linetype = 'dashed', linewidth = 0.7) +

    # Add the text "Flattering" on the left of the line
    annotate("text", x = 5, y = 0.6, label = "Flattering Ranks", hjust = 1, size = 8, fontface = "bold") +

    # Manual fill colors
    scale_fill_manual(values = c("#2A629A", '#FF7F3E'), labels = c("Accurate", "Likable")) +

    # X-axis adjustments
    scale_x_continuous(name = "Advised Rank", breaks = seq(1, 10)) +

    # Theme adjustments
    theme_bw() +
  coord_flip() +
    theme(panel.grid = element_blank(),
          panel.border = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = "none",
          legend.title = element_blank(),
          legend.key.size = unit(1.2, "lines"),
          legend.text = element_text(size = 12),
          plot.title = element_text(hjust = 0.5, size = 20),
          axis.text.x = element_text(size = 14),  # X-axis text size
          axis.text.y = element_text(size = 14),
          axis.title.x = element_text(size = 16, face = "bold"),  # X-axis title size
          axis.title.y = element_text(size = 16, face = "bold"))

```
```{r}
#| label: fig-study2advice
#| fig-cap: "Participants in Study 2 give advice to the person they rank 7th most attractive out of 10. When incentivized for accuracy, the majority recommends betting on that rank. However, when incentivized for likeability, participants inflate their advice to flatter the recipient as more attractive."
study2_time2 %>%
  dplyr::group_by(Condition) %>%
  dplyr::mutate(total_per_condition = dplyr::n()) %>%
  dplyr::group_by(Condition, Advice, total_per_condition) %>%
  dplyr::summarize(n = dplyr::n()) %>%
  dplyr::mutate(pct = ifelse(Condition == "Accurate", -n/total_per_condition, n/total_per_condition),
                Advice = factor(Advice)) %>%
  ggplot(aes(x = pct, y = Advice, fill = Condition)) +
  geom_col() +
  scale_x_continuous("Share of Advice",
                     limits = c(-0.6, 0.6),
                     labels = \(x) scales::percent(abs(x))) +
  scale_y_discrete(limits = rev) +
  # Dashed 'horizontal' line at Advice = 7
  geom_hline(yintercept = 4, col = 'black', linetype = 'dashed', linewidth = 0.7) +
  # Example text annotation (adjust x/y for positioning)
  annotate("text", x = -0.3, y = 9, label = "Accuracy Incentive", hjust = 1, size = 4, fontface = "bold") +
  annotate("text", x = 0.6, y = 9, label = "Likeable Incentive", hjust = 1, size = 4, fontface = "bold") +
  # scale_y_continuous(name = "Advised Rank", breaks = seq(1, 10)) +
  scale_fill_manual(values = c("#2A629A", "#FF7F3E")) +
  papaja::theme_apa() +
  theme(
    legend.position  = "none",
    axis.text.x  = element_text(size = 14),
    axis.text.y  = element_text(size = 14),
    axis.title.x = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 16, face = "bold")
  )
```

We begin by examining the prior beliefs of advisees who uploaded their
selfies. On average, men guess they rank
`r round(mean(study2_time1[study2_time1$gender == "Male",]$Guess),2)` in
their group of 10 and women guess that they rank
`r round(mean(study2_time1[study2_time1$gender == "Female",]$Guess),2)`
(`r papaja::apa_print(t.test(Guess ~ gender, data = study2_time1, var.equal = T))$statistic`).
Notably, people's self-perceptions correlate strongly with the
aggregate ratings of the advisers
(`r round(cor(study2_finalRank$Guess, study2_finalRank$finalRank),2)`,
`r papaja::apa_print(cor.test(study2_finalRank$Guess, study2_finalRank$finalRank))$statistic`).
However, there is substantial heterogeneity in perceptions of
attractiveness. Of the 200 participants,
`r length(unique(study2_time2$Target))` are ranked as 7th most
attractive by at least one adviser. On average, men in this subset
estimate they rank
`r round(mean(TargetsUnique[TargetsUnique$gender == 'Male',]$Guess),2)`th
and women estimate they rank
`r round(mean(TargetsUnique[TargetsUnique$gender == 'Female',]$Guess),2)`th
(`r papaja::apa_print(t.test(Guess ~ gender, data = TargetsUnique, var.equal = T))$statistic`).

Next, we turn our attention to the advisers (see @fig-study2advice). In the
Accuracy condition, those uploading selfies are on average advised to bet on rank
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offer flattering advice.[^1] Importantly, and as predicted, we find
that advisers in the Likeable treatment recommend betting on a lower
rank, communicating that they think the participant in the selfie is
more attractive
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
This suggests that advisers infer flattering someone with pleasant advice
would make the adviser appear more likeable. The distribution shown in
@fig-study2advice shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

[^1]: This shading could be due to concerns of avoiding disappointment.
    However, it could also be that advisers are uncertain about the
    rankings they have given and make a recommendation that combines
    their own belief with a uniform prior. Therefore, our analyses focus
    on the difference between the two conditions.

<!-- David: Making a note here that we should be sure to discuss this at the end -->

<!-- as something for future research. -->

<!-- David: We should also report something on accuracy of the advice given across -->

<!-- the treatments. Even if no difference. -->

```{r}
#| label: tbl-study2regs
#| tbl-cap: "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional."

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study2_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study2_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study2_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

```

Finally, we examine whether flattering advice indeed leads to more
positive evaluations of advisers, or whether flattering advice is
dismissed as insencere. Following our preregistration, we average the
ratings on likeability, warmth, friendliness, and good-naturedness to
create a scale of likeability ($\alpha$ =
`r cronbach.alpha(study2_time3[, 5:8])$alpha`); and we create a scale of
trustworthiness by averaging the ratings of trustworthiness and
sincerity ($\alpha$ = `r cronbach.alpha(study2_time3[, 9:10])$alpha`).

```{r}
#| label: fig-study2rating
#| fig-cap: "Participants tend to rate advisors who suggest betting on a higher rank (implying greater attractiveness) as more likable and warm."

library(ggplot2)
library(dplyr)
library(tidyr)

study2_time3 <- study2_time3 %>% dplyr::mutate(Advice = as.numeric(Advice))

# Reshape the data to long format
study2_long <- study2_time3 %>%
  pivot_longer(cols = c(likable, warmAvg), 
               names_to = "Evaluation", 
               values_to = "Value")

# Create the plot
ggplot(study2_long, aes(x = Advice, y = Value, color = Evaluation)) +  
  # Add points for each observation
  geom_point(size = 2, alpha = 0.7,
             position = position_jitter()) +

  # Add linear regression lines for each metric
  geom_smooth(method = "lm", se = TRUE, aes(fill = Evaluation), alpha = 0.2, show.legend = FALSE) +

  # Customize axis labels
  scale_y_continuous(name = "Evaluation",limits = c(0.9, 5.1)) +  
  scale_x_continuous(name = "Advised Rank",limits = c(1, 10), breaks = c(1,2,3,4,5,6,7,8,9,10)) +

  # Apply a theme
  theme_bw() +  
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "right",
        plot.title = element_text(hjust = 0.5, size = 20)) +

  # Set custom colors for each metric
  scale_color_manual(values = c("#FF7F3E", "#2A629A"), 
                     labels = c("Likability", "Warm")) + 
  scale_fill_manual(values = c("#FF7F3E", "#2A629A"))

```

```{r}
#| label: fig-study2ratingByGroup
#| fig-cap: "After participants receive advice that suggests betting on a rank that was lower than, equal to, or higher than their own guess, they evaluate advisors based on their likability and warmth. Participants rate advisors who give flattering advice and deliver a high evaluation of their attractiveness as more likable and warm than those whose advice implied a lower evaluation of their attractiveness."

library(dplyr)
library(ggplot2)
library(reshape2)

study2_time3 <- study2_time3 %>% dplyr::mutate(Guess = as.numeric(Guess), Advice = as.numeric(Advice)) %>%
  dplyr::mutate(HorL = ifelse(Guess == Advice,0, ifelse(Guess > Advice, 1,-1))) %>%
  dplyr::mutate(HorL = as.factor(HorL)) %>% 
  dplyr::mutate(bonus = ifelse(bet == `Final Rank`,1,0)) %>%
  dplyr::mutate(accuracy = `Final Rank`-Advice)

ratings <- study2_time3 %>%
  group_by(HorL) %>%
  summarise(like = round(mean(likable),2), warmth = round(mean(warmAvg),2))

rating_table <- matrix(c(ratings$like,ratings$warmth), nrow = 2, byrow = TRUE, dimnames = list( c("Likable",'Warm'),c("Lower than", "Equal to", "Higher than")))

rating_df = as.data.frame.table(rating_table) %>%
  dplyr::rename( 'Condition'='Var1', 'Advice' ='Var2') %>%
  dplyr::mutate(Freq = round(Freq,2))

# Create a data frame similar to your structure
data <- rating_df

# Reshape the data to wide format for the heatmap
heatmap_data <- dcast(data, Condition ~ Advice, value.var = "Freq")

# Melt the data for ggplot
melted_data <- melt(heatmap_data, id.vars = "Condition", variable.name = "Advice", value.name = "Rating")

# Create the heatmap
ggplot(melted_data, aes(x = Advice, y = Condition, fill = Rating)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "darkblue", high = "orange", name = "Rating") +  # Grey gradient from light to dark
  geom_text(aes(label = sprintf("%.2f", Rating)), color = "white", size = 8) +
  theme_minimal() +
  labs(x = "Attractiveness (Advice Relative to Self-Expectations)", y = "Evaluation") +
  theme(
    axis.text.x = element_text(size = 12),  # X-axis text size
          axis.text.y = element_text(size = 12),
          axis.title.x = element_text(size = 12),  # X-axis title size
          axis.title.y = element_text(size = 12))

```

As seen in @fig-study2rating, we advisers who suggested that
the advisee was more attractive (lower rank) were indeed rated as more
likeable and warm (b = -0.109, p \< .05; b = -0.133, p \< .001,
respectively; See Columns 1 and 2 of @tbl-study2regs). When using their
own initial guess as a reference point, advisers who suggest a worse rank
are perceived as less likable and warm (see
@fig-study2ratingByGroup). Interestingly, these benefits are not at the
cost of sincerity; advisers who recommend a more favorable rank
are viewed as no less trustworthy (Column 3 of @tbl-study2regs).

We are not powered to do a comparison across the two experimental
groups and did not preregister such a difference. Indeed, we find no
difference in likeability and warmth across the two treatments
(`r papaja::apa_print(t.test(likable ~ Condition, data = study2_time3, var.equal = T))$statistic`,
and
`r papaja::apa_print(t.test(warmAvg ~ Condition, data = study2_time3, var.equal = T))$statistic`,
respectively). We also assess the quality of advice by measuring the
discrepancy between the advised rank and the correct rank. We find
no difference in accuracy across the advice treatments
(`r papaja::apa_print(t.test(accuracy ~ Condition, data = study2_time3, var.equal = T))$statistic`).

## Discussion

In another context where advice communicates ego-relevant information
(here, people's attractiveness), we find that the advice people give is
contingent on their incentives. Specifically, when they get rewarded for
being more likeable, they recommend that the advisee bet on a more
favorable rank than when they are incentivized for accuracy.
Importantly, advisees do not discount flattering advice and instead
evaluate people who advise them to bet on a more attractive rank as
warmer and more likeable. These gains to interpersonal perceptions do
not come at the cost of trustworthiness, even when the advice is
inflated.

# Study 3

```{r}
#| include: false

study3_math50 <- read.csv(here::here("Data/Cleaned Data/S3_Math50_cleaned.csv"))
study3_time1All <- readRDS(here::here("Data/Study3_time1All.Rds"))
study3_time1 <- readRDS(here::here("Data/Study3_time1.Rds"))
study3_time2_giver <- readRDS(here::here("Data/Study3_time2_giver.Rds"))
study3_time2_giver50 <- readRDS(here::here("Data/Study3_time2_giver50.Rds"))
study3_advice <- readRDS(here::here("Data/Study3_time2_advice.Rds"))
study3_time3 <- readRDS(here::here("Data/Study3_time3.Rds"))

```

In Study 3, we replicate the main effect in a different ego-relevant
domain: intelligence. The experimental design is conceptually identical
to Study 1, but this time we utilize a more direct measure of
expectation by asking participants to estimate their scores.

This three-stage experiment begins with measuring actual performance.
Participants are then anchored to low or high performance expectations
and asked to guess how many questions they answered correctly. As in
previous studies, we recruit advisers who are presented with either just
the advisee's true performance or both the true performance and the
number of questions the advisee believes they answered correctly.
Advisers are then asked to provide advice on choosing a competitor.
Finally, we present the advice to the advisees and ask them to make
their final choice. Our key hypotheses are: (1) advisees anchored to
high expectations will be more likely to receive advice to compete
against the high-performing group, and (2) men will be more likely than
women to be advised to compete against the high-performing group when
expectations are disclosed.

## Methods

```{r}
#| include: false

subgroup_averages <- data.frame(subgroup = numeric(), avg_score = numeric())
set.seed(807)
# loop through 1000 iterations of randomly selecting 10 participants and computing their average score
for (i in 1:1000) {
  # randomly select 10 participants
  subgroup <- sample(study3_math50$ProlificID, size = 5, replace = FALSE)
  # compute the average score for the selected participants
  subgroup_avg <- mean(study3_math50$Score[study3_math50$ProlificID %in% subgroup])
  # add the subgroup and its average score to the dataframe
  subgroup_averages <- rbind(subgroup_averages, data.frame(subgroup = i, avg_score = subgroup_avg))
}

p5 <- quantile(subgroup_averages$avg_score, probs = 0.05)
p95 <- quantile(subgroup_averages$avg_score, probs = 0.95)

sortedScore <- sort(study3_math50$Score, decreasing = TRUE)
LP_lower <- sortedScore[50]
LP_upper <- sortedScore[31]
HP_lower <- sortedScore[20]
HP_upper <- sortedScore[1]

```

We begin by first recruiting a sample of
`r scales::comma(dim(study3_math50)[1])` participants from Prolific to
complete a 10-question multiple choice mathematics quiz. The questions
were taken from a paper-version of the ASVAB standardized exam, such
that answers were not available online. Participants had five minutes to
answer the quiz and were paid 10 cents for each correctly answered
question. Like before, we define top 20 scores as "High Performers" and
the bottom 20 scorers as the "Low Performers." On average, participants
answered `r round(mean(study3_math50$Score),2)` questions correctly,
High Performers scored between `r HP_lower` and `r HP_upper`, and Low
Performers scored between `r LP_lower` and `r LP_upper`. In order to
anchor the expectations of participants in our main experiment, we
simulated 1,000 pairings of groups of five participants, with the
5<sup>th</sup> percentile of groups scoring an average of `r p5` and the
95 th percentile scoring an average of `r p95`. We report these averages
to participants in the Low Expectations and High Expectations treatment,
respectively.

We then recruited `r scales::comma(dim(study3_time1All)[1])`
participants for the role of advisees in Stage 1 of our main experiment.
To arrive at a gender-balanced sample, we dropped the last two male
participants to complete the survey, ending up with a sample of 500 men
and 500 women ($M_{\text{Age}}$ = `r mean(study3_time1$Age)`).
Participants completed the same 10-item mathematics quiz as the earlier
participants and were informed that their performance would affect their
bonus earnings in a follow-up stage to be conducted a few days later.
After completing the quiz, we informed them of the average score of a
group of five participants from the preliminary survey. We randomly
assigned them to learn about the 5<sup>th</sup> percentile of groups,
which scored `r p5`. ("Low Expectations" treatment) or the
95<sup>th</sup> percentile of groups, which scored `r p95` ("High
Expectations" treatment). Participants then made a guess
(unincentivized) about how many questions they think they answered
correctly. The survey concluded with basic demographic questions.

We then recruited `r scales::comma(dim(study3_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We began
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 had completed, and informed them of the
average score of all participants in the preliminary study. Advisers had
to recommend whether an advisee should compete against the Low
Performers or the High Performers (we used these terms in the survey).
We anticipated that being told to compete against High Performers is
more flattering and hence being told to compete against the Low
Performers would be disappointing if one had expected to do well.
Advisees could earn a bonus if their score was equal to or higher than
that of a randomly selected member from their chosen group. Competing
against the High Performers was more challenging, resulting in an
average bonus of 50 cents, compared to 30 cents when competing against
other groups.

Like before, advisers were randomly assigned to one of two treatments.
In the "Baseline" treatment, they only observed the score of the advisee
on the mathematics quiz. In the "Expectation" treatment, they observed
the score as well as the advisee's guess for how many questions they
answered correctly. We want to emphasize that sice the outcome is
determined only by the past score on the quiz, the advisee's guess is
immaterial to which group they should compete against. Moreover, in
neither treatment did they receive any demographic information about
their advisee. Participants gave recommendations to 10 advisees, which
unbeknownst to them were five men and five women matched to have
identical performance on the test.[^2] They were informed that if their
advice was shown to a participant who returned for the follow-up survey,
they would receive the identical bonus as that participant. The survey
then concluded with basic demographic questions.

[^2]: We made this decision to account for the possibility of gender
    differences in performance.

Finally, we invited participants from Stage 1 back for the follow-up
survey. Following our preregistration, we kept the survey open for 7
days. In total, `r scales::comma(dim(study3_time3)[1])` participants
(`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Female',])[1])`
women) returned. The brief survey reminded them of the task they
completed in Stage 1, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally were reminded them of how
many questions they guessed they had answered correctly. Importantly,
they were not informed of their true score or the score of the groups
they could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

## Results

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants completed the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study3_time1[study3_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study3_time1, var.equal = T))$statistic`).
The expectations treatment did, however, affect how well they thought
they performed. Participants in the "Low Expectations" treatment guessed
a score of
`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study3_time1[study3_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study3_time1, var.equal = T))$statistic`),
showing that the manipulation was successful, albeit small. Contrary to
our expectations, we did find a gender difference in performance: men
scored
`r round(mean(study3_time1[study3_time1$Gender == "Male",]$count),2)` on
average, while women scored
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study3_time1, var.equal = T))$statistic`).
Consistent with this difference, men thought they answered more
questions correctly than did women
(`r round(mean(study3_time1[study3_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study3_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We
observe that women underestimate their performance by 0.76 points, while
men do so by only 0.27 points
`r papaja::apa_print(t.test(overconfidence ~ Gender, data = study3_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`.
Although neither gender is overconfident, men on average are more
confident in their performance than are women, as we had expected.

```{r}
#| label: tbl-study3regs
#| tbl-cap: "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor."

study3_advice <- study3_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study3_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study3_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study3_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

```

```{r}
#| label: fig-study3genderdiff
#| fig-cap: "Advisers observed real performance and the participants' estimated performance, but not their gender. As a result of their higher expectations, men (blue) were advised to compete against the high performance group more often than women (red). The grey line shows advice absent expectations, which did not differ by gender."

library(dplyr)
library(ggplot2)

# Calculate averages for the performance, grouped by condition and gender
ProbHG <- study3_advice %>%  
  group_by(Condition, Performance, Gender) %>%  
  summarise(avg_hardGroup = mean(Advice), .groups = 'drop')

# Create a data frame for the NONE condition as the baseline
baseline <- ProbHG %>%  
  filter(Condition == "None") %>%  
  select(Performance, baseline_avg = avg_hardGroup)

# Merge with the original data to get comparison values
comparison_data <- ProbHG %>%  
  filter(Condition != "None") %>%  
  left_join(baseline, by = "Performance") %>%  
  mutate(diff = avg_hardGroup - baseline_avg)  # Calculate the difference

# Create a scatter plot comparing conditions with NONE as baseline
ggplot() +
  # Step lines for the NONE condition
  geom_segment(data = baseline,
               aes(x = Performance, 
                   xend = lead(Performance, 1),  # Create horizontal lines to the next performance point
                   y = baseline_avg, 
                   yend = baseline_avg,
                   linetype = "None"), 
               color = "grey70", size = 0.5, 
               na.rm = TRUE,
               show.legend = TRUE) + 
  geom_point(data = comparison_data %>% filter(Gender == "Male"), 
             aes(x = Performance + 0.5, y = avg_hardGroup, shape = Gender, color = Gender), 
             size = 4) +  # Males with an offset to the left
  geom_point(data = comparison_data %>% filter(Gender == "Female"), 
             aes(x = Performance + 0.5, y = avg_hardGroup, shape = Gender, color = Gender), 
             size = 4) +  # Females with an offset to the right
  scale_shape_manual(values = c(16, 17), labels = c("Female", "Male")) +  # Circle for Females, Triangle for Males
  scale_color_manual(values = c( "#FF7F3E","#2A629A"), 
                     labels = c("Female", "Male")) +  
  scale_y_continuous(name = "P(Compete Against High Performance)",
                     labels = scales::percent,
                     limits = c(0, 1.1),
                     breaks = c(0,1)) +  # Set limits for y-axis
  scale_x_continuous(name = "Performance",
                     limits = c(0, 11),expand = c(0, 0), 
                     breaks = seq(0, 10, 1)) +  
  theme_bw() +  
  theme(panel.grid = element_blank(), 
        panel.border = element_blank(), 
        axis.line = element_line(colour = "grey"), 
        legend.position = "bottom", 
        legend.title = element_blank(), 
        legend.key.size = unit(1.2, "lines"), 
        legend.text = element_text(size = 12), 
        plot.title = element_text(hjust = 0.5, size = 20))

```

Next, we examine whether advisers took the advisee's expectations into
account. Column 1 of @tbl-study3regs reports a linear probability model
on advising to compete against the High Performer group for advisers in
the "Expectations" treatment, controlling for the true performance of
the advisee. Because each adviser made ten recommendations, we cluster
standard errors at the adviser level. However, contrary to our
expectations, we do not see a significant effect of the expectations
treatment. This may be because the induced difference in expectations
was too small.

However, recall that men were more confident in their performance than
were women. Our theory thus predicts that showing expectations should
lead to more flattering advice (i.e., a recommendation to compete
against the High Performer group) for men than for women. We report a
linear probability model with advice to compete against the high group
as the outcome measure, and the advisee's gender, whether expectations
were shown to the adviser, and the interaction of the two in Column 2 of
@tbl-study3regs. As predicted, we find a significant interaction effect:
men are more likely to be advised to compete against the High Performers
when expectations are shown than when they are hidden ($p < 0.001$). We
show this result graphically in @fig-study3genderdiff.

To examine the quality of advice, we computed the expected bonus
earnings for someone who followed the recommendations. For example, if
an adviser suggested competing against the High Performer group, we
matched the advisee against all 20 members of that group and determined
how often their score matched or exceeded that of the member. We then
multiplied this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of
@tbl-study3regs. Displaying expectations led men to be advised to
compete against the High Performer group more often, and this advise
turned out to be bad: men receive worse advise than do women when
expectations are displayed, but not in their absence.[^3]

[^3]: This analysis was not preregistered, and we note here that the
    reduction in expected earnings is small. However, it is interesting
    that expectations have a negative effect for men who underestimate
    their performance on average. One possibility is that advisers
    suggest the High Performer group more often than is optimal. We
    return to the possibility that advice is overall biased to be
    flattering in Study 2.

```{r}
#| label: tbl-study3time3
#| tbl-cap: "Column 1 displays the actual bonus based on the performance and whether the participant is primed with high expectations. Column 2 displays the actual bonus based on the performance and gender of the participant. Columns 3 and 4 display the chance of adopting the advice based on the gender of the participants, whether the advisee sees expectations, and whether the advice is to compete against high performers. The former considers only the main effect, while the latter also includes the interactive effect."

study3_time3 <- study3_time3 %>% dplyr::mutate(AdviseeCondition = factor(AdviseeCondition, levels = c('Low', 'High'))) %>%
     dplyr::mutate(Adopt = ifelse(Bet == Advice,1,0))

modelsummary::modelsummary(list(`(1)` = lm(bonus ~ AdviseeCondition + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(2)` = lm(bonus ~ Gender + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(3)` = lm(Adopt ~ Gender + GiverCondition + Advice, data = study3_time3),
                                `(4)` = lm(Adopt ~ Gender + GiverCondition + Advice + Gender*GiverCondition + GiverCondition*Advice + Gender*Advice, data = study3_time3)),
                           vcov = c("classical", "classical","classical","classical"),
                            coef_map = c("AdviseeScore" = "Performance",
                                         "AdviseeConditionHigh" = "High Expectation",
                                        "GenderMale" = "Advisee Male",
                                        "GenderMale:GiverCondition" = "Expectation x Male",
                                        "GiverCondition" = "Expectation Shown",
                                        "Advice1" = "Advice: High Performer",
                                        "GenderMale:Advice1" = "Advice: High Performer x Male",
                                        "GiverCondition:Advice1" = "Expectation x Advice: High Performer",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

```

To determine whether flattering advice is truly costly, however, we need
to examine the outcome of the advisees. In particular, they could ignore
flattering advice, recognizing it as such and thus failing to adhere to
it. Align with our prediction, participants who were primed with high
expectations earned less when their expectations were conveyed to the
advisers, although this result is only directional (see Column 1 of
@tbl-study3time3). Similarly, as shown in Column 2 of @tbl-study3time3,
we also found that male participants received less reward when their
expectations were presented, even when their performance was the same.
These findings suggest that flattering advice is not without
consequences. Furthermore, our research indicates that because males are
more likely to follow such advice (see Column 4 of @tbl-study3time3),
their tendency towards competition intensifies the cost of such advice.

## Discussion

We found supportive evidence for our argument in a different domain.
Advisors tend to consider advisees' self-expectations and match their
advice accordingly, aiming to avoid disappointment. Consistent with the
confidence gap narrative, men tend to underestimate their mathematics
test scores less than women. When these expectations are displayed to
advisors, they are more likely to advise men to compete against high
performers. Notably, this turns out to be poor advice: men whose
advisors were aware of their expectations received worse advice. Our
findings suggest that men with the same scores as their female
counterparts ended up earning less, although this result is only
directional. This discrepancy may be due to men receiving more favorable
advice, being more likely to follow it, and ultimately facing worse
outcomes.

<!-- David: One more sentence about the earnings difference here, which does not -->

<!-- differ significantly. And one sentence about the new analysis of whether -->

<!-- people follow the advice or not. -->

# General Discussion

Advice has the potential to shape people's career and personal outcomes.
Honest feedback, however, may be painful to learn if it falls short of
one's expectations. As prior work notes, this may motivate people to
avoid information and avoid seeking help [@Benabou2022; @Golman2017;
@Jaroszewicz2022]. We present evidence from two experiments that
advisers are also cognizant of this cost. As a result, they present
flattering advice that avoids disappointing the recipient, and correctly
anticipate that this boosts how advisees perceive them. However, this
flattering advice comes at a cost to advisees, who would do worse if
they followed it blindly, as we show in Study 1.

Moreover, a desire to avoid disappointment also means that advisers have
to take into account the expectations of the advisee. We document that
men are more optimistic about their performance than women are. As a
result, they receive more flattering advice and are more likely to be
told to aim higher. Notably, in the context of our experiment, this
turns out to be bad advice ex post.

Our findings have implications for organizational practice, where
mentoring and advice-giving may take into account an employees'
expectation. We document this as a novel source of gender bias.
Organizations could reduce this bias by calibrating employees'
expectations to reduce overconfidence.

Participants in our experiment were paired anonymously. Even so, we
document this desire to avoid disappointment. We anticipate that advice
would be more flattering in face-to-face communication and without
anonymity. Moreover, existing relationships might make it even more
difficult for advisers to be honest.

In our experiments, participants only received advice once and did not
evaluate the adviser after observing the outcome of their decision. For
example, advice that leads to bad outcomes may undermine the
interpersonal benefits of flattery. Alternatively, people may still like
the flattering advice and not fault the adviser for the bad outcome.
Moreover, our setting involved only a single piece of advice on one
task. Future research could examine whether people return to those who
gave them flattering advice, or if they prefer someone who gave them the
honest (but unpleasant) truth.

Advice has long been studied from the perspective of the receiver.
Similarly, research on belief utility has examined how recipients
respond to the valence of the information they receive. Here, we show
that advisers, too, take into account the psychological impact of the
information they convey. They may have even greater motivations to avoid
conveying bad news, because they incur the interpersonal costs of
delivering unfavorable information without reaping the benefits from
helping someone make a better choice.

# References

::: {#refs}
:::