### 2025-03-12 manuscript/manuscript.qmd
manuscript_0311

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle: "Flattering Advice"
date: yes
author: 
  - name: "Amanda Chen"
    corresponding: true
    orcid: 0000-0002-4428-627X
    email: zchengj@connect.ust.hk
    affiliations:
      - id: id1
        name: "The Hong Kong University of Science and Technology"
        department: "Department of Management"
        city: "Hong Kong"
  - name: "David Hagmann"
    orcid: 0000-0002-2080-997X
    affiliations:
    - id: id1
# author-note:
  # disclosures:
    # gratitude: The authors are grateful to XXX for their helpful comments.
abstract: Advice that improves decision-making does not always feel good. While advice has the potential to improve decision-making by helping advisees more accurately assess feasibility and predict potential outcomes, it can also lead to disappointment when it challenges the advisee’s overly optimistic expectations. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers prioritize avoiding disappointment at the expense of accuracy and their own earnings. In study 1, advisers use flattering advice to create more positive impressions, and advisees reward this by rating these advisers as more likable, despite the advice being less honest and less accurate. Even advisers incentivised for accuracy still tailor recommendations to aspirational goals expressed by advisees, resulting in financial loss for them and the advisees (Study 2). The concern for interpersonal consequences may lead to inequities if there are differences in expectations across social groups. In Study 3, we examine a setting in which men expect to perform better than women. We show that advisers take into account these expectations, leading to systematically different advice for men and women even when their gender is unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making.

keywords: [Advice, Interpresonal relationship, Gender difference]
bibliography: bibliography.bib
floatsintext: true
mask: false
link-citations: false
format:
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    prefer-html: true
  apaquarto-docx:
    default: true
    prefer-html: true
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| include: false

library(magrittr)
library(ltm)
library(ggplot2)
library(gridExtra)
library(dplyr)

knitr::opts_chunk$set(fig.width = 8, fig.height = 4, cache = F, echo = F,
                      fig.align = 'center',warning = FALSE, message = FALSE)

```

<!--# Amanda: I replaced all 'advisers' with 'adviser' for consistency. -->

# Introduction

Advice figures prominently in workplace and personal decision-making,
shaping choices about promotions, career paths, and everyday dilemmas
[@Gino2007; @Harvey1997; @Soll2009]. While others can provide
information in the form of opinions [@Trouche2018], feedback
[@Kluger1996], or personal observations [@Yoon2021] to influence the
decision-maker, this project focuses specifically on advice—defined as
"information that is perceived as intended to help an advisee deal with
their issue of deliberation" [@Milyavsky2024]. Advice is distinct from
other forms of input because its primary purpose is to improve
decision-making by leveraging others' expertise.

Advice has the power to enhance decisions, yet it also carries the risk
of disappointment. For example, a career consultant might advise an
employee to enroll in a management program to strengthen their skill set
before competing for a leadership role. Ideally, such honest advice
helps the employee identify areas for personal growth. In practice,
however, the employee might feel overlooked and frustrated that their
ambitions are not immediately supported. As a result, advisers may be
inclined to offer flattering recommendations that align with—or even
exceed—advisees' hopes, such as encouraging the employee to pursue the
leadership role despite concerns about their readiness. In this way,
interpersonal dynamics may discourage full candor in giving advice.

The objective of this research is to examine the dilemma advisers face:
Should they provide accurate but potentially painful advice, or soften
their recommendations to preserve the advisee's mood and protect their
own standing? Given that advisees tend to prefer advisers who withhold
bad news [@Shalvi2019] nd penalize those who deliver it [@John2019],
advisers may reasonably avoid full candor—especially when they bear no
cost for the advisee’s mistakes. Specifically, when someone's
expectations are (too) high, an adviser may hesitate to provide advice
that challenges them, particularly if doing so risks straining the
relationship or causing negative interpersonal consequences.

Our research contributes to the growing literature on advisers' motives
and intentions in advice-giving. This line of work examines how
advisers' perceived dishonesty or helping intentions influence advice
utilization [@Haran2020; @Mackinger2017; @Milyavsky2024; @Van2009]. By
examining advice within its interactional and communicative context, we
extend this research by incorporating another dimension of advisers'
motives—interpersonal concerns. We demonstrate that advisers adjust
their recommendations to avoid hurting the advisee’s feelings and
self-image. Second, we contribute to the belief-based utility
literature, which suggests that individuals derive satisfaction from
their beliefs about the world and themselves [@Loewenstein2018]. We show
that advisers also place value on preserving advisees' self-perceptions,
often tailoring their advice to align with the advisee’s expectations.
Furthermore, we reveal that this advice-giving strategy has unintended
consequences—it transforms the confidence gap [where women tend to be
less confident than men; @Lundeberg1994] into an advice gap, in which
women are less likely than men to receive advice to aim higher, thereby
reinforcing existing gender inequalities.

## The Motives Behind Advice-Giving

While much of the literature on advice assumes that advisers aim to
provide the most helpful advice, their motivations are often more
complex [@Yaniv2012; @Dalal2010; @Rader2017]. Other than the accuracy
and usefulness of the advice, interpersonal concerns—such as maintaining
a positive image, avoiding conflict, or preserving social harmony—can
shape what advice is delivered and how [@Sniezek2001; @Levine2020]. For
example, research has shown how people often avoid providing negative
feedback directly, opting instead for ambiguity or indirect phrasing to
minimize discomfort [@Levine2020]. These interpersonal dynamics suggest
that advice is not always optimized for the recipient’s best interest
but is instead filtered through social and psychological considerations.

In particular, advisers may strategically soften their messages or
withhold unfavorable evaluations to manage how they are perceived or to
prevent discomfort for both themselves and the advisee [@Van2009;
@Schaerer2018; @Chentsova2012]. When advisers anticipate disagreement or
resistance, they may soften their recommendations or frame them more
diplomatically, sometimes at the expense of accuracy. This suggests that
advice is not merely a means of sharing information but also a tool for
managing social interactions and reducing potential friction [@Van2009].
These interpersonal considerations become even more pronounced when
there is a power imbalance between the adviser and advisee, as
lower-power advisers may be more hesitant to offer direct or critical
guidance to avoid negative repercussions [@Schaerer2018].

Just as advisers may hesitate to provide unwelcome truths, advisees
themselves are often resistant to hearing them. People often go to great
lengths to avoid unpleasant information, deriving belief-based utility
from maintaining favorable views of themselves [@Golman2017; @Ho2018;
@Loewenstein2018]. Moreover, they often prefer advisers who withhold bad
news from them [@Shalvi2019] and punish those who communicate bad news
[@John2019]. Advisers who fear they will be blamed or disliked for
undermining someone's confidence may reasonably avoid candid advice,
particularly if they don't incur a cost when the advisee makes a
mistake.

Providing flattering advice, however, should not be seen as maliciously
misleading the advisee. Rather than being irrational, it tends to be
more optimistic, and this optimism is not necessarily misplaced. Often,
the advisee possesses insights about themselves—such as their motivation
and past experiences—that the adviser may overlook or have no access to.
Take, for example, an employee deciding whether to compete for a
leadership role. While the adviser might focus on the employee’s lack of
management knowledge, the employee may have strong leadership
characteristics, excellent communication skills, and persuasive
abilities. In this way, flattering advice can encourage the advisee to
set ambitious goals, fostering a self-fulfilling prophecy where
confidence drives success [@Jussim1986; @Locke2006].

Of course, if the advice is overly inflated, it may become unhelpful.
However, this is unlikely to happen, as advisers typically aim to boost
the advisee’s confidence while maintaining credibility. Suggesting an
unrealistic option—such as encouraging an entry-level employee to
compete for a CEO position—would come across as insincere or even
mocking rather than flattering.

Beyond protecting the advisee's beliefs and fostering their confidence
and motivation, advisers may also offer flattering advice to shield
themselves from blame if the ultimate decision leads to a poor outcome.
Although both the adviser and advisee recognize the inherent uncertainty
between the advice, the decision, and the final outcome, people tend to
attribute poor outcomes to the adviser [@Weiner1985; @Palmeira2015].
However, research suggests that people differentiate between errors made
intentionally—such as those driven by insufficient knowledge or
deliberate bias—and those made unintentionally, with harsher penalties
imposed on advisers perceived to have deliberately undermined the
decision [@Haran2020]. Flattering advice, even when ultimately
unhelpful, may reduce the likelihood of blame because advisees are less
likely to suspect ill intent from an adviser who appears supportive.
Advisees may instead interpret the poor outcome as an unintended error
rather than deliberate misguidance, particularly because they view the
adviser positively and believe the adviser genuinely holds them in high
regard [@Palmeira2015; @Haran2020]. Even when advisees recognize that
flattering advice was intended to boost their confidence, research shows
that people tend to appreciate such misplaced good intentions, as they
strengthen social bonds by making the gift-giver feel socially closer
[@Zhang2012].

Advisers may choose to inflate their advice for different reasons,
depending on their relationship with the advisee and the stakes of the
decision. For instance, advisers might offer flattering advice to
strangers out of politeness, provide inflated encouragement to friends
to offer emotional support [@Dalal2010], or use their recommendations to
motivate individuals with particularly ambitious or competitive
personalities. We recognize that multiple motives may simultaneously
drive this behavior, and their relative influence—as well as advisers'
overall tendency to inflate advice—may vary across situations. However,
in this project, we focus specifically on whether advisers intentionally
inflate their advice and explore the potential consequences of doing so.

One downstream consequence of giving flattering advice is the potential
to reinforce gender differences — specifically, men are more likely to
receive such inflated advice. Prior work has documented that men receive
more aspirational advice and women more risk-averse advice, which may
explain why women are less likely to apply for and ultimately obtain
more rewarding positions [@Kanze2018]. Research on gender discrimination
has focused on the role of unconscious bias [@Greenwald2009;
@Greenwald1995], but interventions that try to reduce this bias have
been largely unsuccessful [@Chang2019; @Paluck2009]. Recent work has
proposed cognitive, rather than unconscious, biases that can lead to
gender discrimination through the formation of false beliefs
[@Hagmann2024]. Here, we propose that one overlooked mechanism driving
gender disparities in career outcomes may lie in the differential
provision of flattering advice, shaped by the advisee's
self-expectations.

Specifically, when advisers aim to make their advice appear flattering,
they may rely less on an objective evaluation of the advisee's
competence and instead anchor their advice on the advisee’s
self-expectations. Drawing on the literature on belief-based utility and
expectation-based reference points [@Koszegi2006; @Koszegi2009], we
suggest that advisers may align their advice with how ambitious or
confident the advisee appears, so that they give optimistic but not
entirely unrealistic recommendations. This dynamic may inadvertently
disadvantage women. Research shows that men generally hold higher
self-expectations or are more likely to display confidence, while women
tend to exhibit more modest self-assessments despite comparable
competence [@Kanze2018]. As a result, advisers may unintentionally offer
men more ambitious and aspirational advice, encouraging them to pursue
higher-rewarding opportunities, while offering women more cautious,
conservative advice that aligns with their lower expressed
self-expectations. This differential treatment, though seemingly
well-intentioned, can have profound downstream consequences. By
tailoring advice to match the advisee's perceived self-expectations,
advisers may subtly nudge men toward risk-taking, leadership roles, or
competitive opportunities, while directing women toward safer, less
ambitious paths — ultimately reinforcing existing gender disparities in
career outcomes.

# Overview of studies

In our experiments, however, advisees have no opportunity to punish the
adviser, and advisers' financial incentives are linked to the decisions
the advisees make. We propose that advisers nonetheless have reason to
provide flattering advise because they recognize the psychological toll
of disappointing news. As a result, they may similarly incur a hedonic
cost for delivering unpleasant information that they anticipate may
distress the advisee. Thus, even in the anonymized context of an online
experiment, we hypothesize (and find) that advisers are relucant to
disappoint advisees and advice is thus biased upward, away from
accuracy. We show that this reduces the disappoint advisees and advice
is thus biased upward, away from accuracy and toward confirming
advisees' priors. We show that this reduces the quality of advice—but
that advisees nonetheless view flattering advisers more favorably.

We present the results of three preregistered experiments in which
participants are paired anonymously as advisers and advisees. Advisers
are incentivized to offer accurate advice, with bonus earnings depending
on the outcome of an advisee's decision. We study a setting in which
advisees can communicate their expectations to an adviser, and
experimentally manipulate whether advisers observe this expectation. Our
design simulates organizational contexts where mentors and managers
often have some sense of an employee's confidence or aspirations and
thus could adjust their advice to avoid disappointing them.
Interpersonal considerations would likely be stronger in situations in
which the adviser and advisee have an existing relationship and when the
advice is delivered face-to-face. Moreover, advisers usually do not
suffer any direct costs when their flattering advice leads to a poor
outcome for the recipient, particularly when it is not clear what would
have happened under a counterfactual.[^1]

[^1]: In some cases, however, repeated interaction might also increase
    incentives for honest feedback. Someone who is known to persistently
    give overoptimistic advice may be viewed as less trustworthy in the
    long run.

Across all three studies, we use a three-stage design. In Stage 1, we
measure the advisee’s performance and their self-expectations. In Stage
2, we show the measured performance to the adviser and ask them to
provide advice on how the advisee can maximize their chances of earning
a bonus. Half of the advisers also receive information about the
advisee’s self-expectations. Importantly, this expectation information
is objectively irrelevant to the task, as the bonus is solely based on
the advisee’s performance, which is already measured. While in real
life, self-expectations might contain useful insights (such as
unobservable skills or motivation), in our setup, self-expectations
serve no informative value. Moreover, flattering advice provided in this
setup cannot be justified as a motivational boost (like a pep talk)
because the performance has already been fixed. Finally, in Stage 3, we
re-invite the advisees from Stage 1, show them the advice they received,
and observe how they make their final decisions.

In Study 1, we ask advisees to upload photos of themselves and advisers
rank them on their attractiveness. Advisers then recommend which rank
the advisee should bet on, receiving an incentive either for the
advisee's accuracy or for being evaluated favorably. Both groups of
advisers recommend a rank that is more attractive than what they
themselves have evaluated the advisee, and those incentivized for
likability further inflate the ranking. Advisees evaluate advisers more
favorably when they recommend betting on a more attractive rank,
including viewing them as more trustworthy.

In Study 2, we nudge advisees into a preference for competing against
either a group of top-performers or a group of low-performers, through
the use of a default option. Advisers who observe the actual performance
of advisees and therefore have the relevant information needed to make a
recommendation nonetheless take into account the initial preference when
recommending which group they should compete against.

In Study 3, we examine a setting in which participants report their
expected performance on a mathematics quiz. Advisers observe the
test-taker's true score and, in one treatment, the score they guessed
they would receive. They then recommend whether the advisee should
compete against a group of high performers or low performers.
Importantly, this "competition" is based on the past performance
observed by the adviser and thus the advisee's expectation does not
provide instrumental information. However, we find that (1) men expect
to perform better than women, given identical performance, and (2)
advisers take these expectations into account. As a result, when
expectations (but not gender) are known to advisers, the advice given to
men and women differs, and men are more often advised to compete against
the group of high performers. Notably, this leads to worse advice for
men than for women.

# Open Science Statement

We report all manipulations, measures, and data exclusion in our
experiments. The preregistration reports, screenshots of all
experimental materials, and the analysis code to replicate all
statistical analyses and figures are available on the Open Science
Framework
(<https://osf.io/8r3d4/?view_only=5ad7bafcd16b4d4ba08bb28b0e2bd02d>).

# Study 1

We first examine if advisers believe that flattering advice will improve
how they are perceived, and if so, whether this perception is true. We
do this by manipulating the incentives for advisers, who receive a bonus
either based on how they are evaluated by an advisee, or a bonus for the
advisee's accuracy.

The study takes place in three stages. First, a group of advisees upload
photos of themselves ("selfies") and are grouped with nine other
participants of the same sex. We then recruit participants of the
opposing sex to rank them from most to least attractive and to provide
advice. Specifically, they advise the participant they ranked as the 7th
most attractive (i.e., 4th least attractive) on what rank they should
bet they were ranked by a larger group of raters. Advisers were randomly
assigned to two treatments, receiving a bonus payment either if the
advisee guessed their rank accurately or if the advisee evaluated the
adviser as likeable as measured by a scale response. We hypothesize that
advisers incentivized to be liked will recommend betting on a lower
rank, i.e. that the advisee is more attractive.

```{r}

study1_time1 <- readRDS(here::here("Data/Study1_time1.Rds"))
study1_time1_200 <- readRDS(here::here("Data/Study1_time1_200.Rds"))
study1_time2 <- readRDS(here::here("Data/Study1_time2.Rds"))
study1_finalRank <- readRDS(here::here("Data/Study1_finalRank.Rds"))
study1_time3 <- readRDS(here::here("Data/Study1_time3.Rds"))

```

## Methods

We recruit 300 participants from Prolific and, after asking demographic
questions, invite them to upload photos of themselves (selfies) to be
rated by other participants on attractiveness. We obtain selfies from
`r scales::comma(dim(study1_time1[study1_time1$gender == 'Male',])[1])`
men and
`r scales::comma(dim(study1_time1[study1_time1$gender == 'Female',])[1])`
women adhering to our instructions (e.g., did not include other people).
In line with our preregistration, we select the first 100 photos from
women to arrive at a gender-balanced sample ($M_{\text{Age}}$ =
`r mean(study1_time1_200$age)` years). Participants are informed that
their selfies will be randomly grouped with those of nine other
participants of their sex and ranked in terms of attractiveness by a
group of new Prolific participants of the opposite sex. They then make
an unincentivized guess of their rank.

Next, we recruit `r scales::comma(dim(study1_time2)[1])` participants
from Prolific for the role of advisers ($M_{\text{Age}}$ =
`r round(mean(study1_time2$age),2)` years;
`r round(mean(study1_time2$gender == "Female")*100, 2)`% Female). We
first collect demographic information, then match them to a group of the
opposite sex. They rank selfies of ten participants from most to least
attractive by entering their rankings into a text box next to each
picture. Because of a limitation with the survey software, we could not
validate that each rank is given only once, and we remove 115
participants who failed to follow instructions and did not provide a
complete ranking.[^2]

[^2]: Moreover, we could not collect data from any participant who did
    not select any participant as the 7th most attractive participant,
    as that image would be shown on subsequent pages.

After submitting their ratings, participants see the photo of the
participant they ranked as the 7th most attractive (i.e., the 4th least
attractive). We remind them of the rank they have just given to that
person and inform them that this participant will be invited back and
can earn a \$1 bonus if they correctly guess their rank. The rank is
determined by the aggregate ratings of all participants who have ranked
this group. Because the participant does not observe the other nine
people in the group, they would depend on the adviser's recommendation
along with their own assessment. We randomly assign advisers to one of
two incentivization schemes. In the "Accuracy" treatment, they receive a
bonus identical to the advisee: \$1 if they guess their rank correctly.
In the "Likeability" treatment, we inform them that the advisee will
rate them on a 5-point Likert scale on how likeable they thought they
are. Each point on the scale would translate to a bonus of 20 cents.
Advisers then select a rank (from 1 to 10) that they recommend the
advisee to bet on.

Finally, we invite participants from Stage 1 who were ranked as the 7th
most attractive participant by at least one adviser (so that they
received advice) back for the follow-up survey. Following our
preregistration, we keep the survey open for 7 days. In total,
`r scales::comma(dim(study1_time3)[1])` participants
(`r scales::comma(dim(study1_time3[study1_time3$gender == 'Male',])[1])`
men,
`r scales::comma(dim(study1_time3[study1_time3$gender == 'Female',])[1])`
women) complete the follow-up survey. We remind them of the selfie they
uploaded in Stage 1 and that a group of 10 selfies, including theirs,
had been rated by other participants from Prolific. They then see the
advice from a randomly selected adviser and make their own guess with a
\$1 incentive for accuracy. Finally, they rate the adviser's likability,
warmth, friendliness, good-naturedness, trustworthiness, and sincerity
on 5-point Likert scales [adapted from @Fiske2007].

## Results

```{r}
#| include: false
library(dplyr)

TargetsUnique <- study1_time2 %>%
  dplyr::select(Target) %>%
  unique()

TargetsUnique <- merge(TargetsUnique,study1_finalRank, by.x = 'Target', by.y = 'QID7_Id')

```

```{r}
#| label: fig-study1advice-old
#| fig-cap: "Advisers were asked to suggest what rank to bet on for someone they ranked as the 7th most attractive out of ten people. When incentivized for accuracy, most of them advised betting on 7, which was based on their evaluation of the recipient's attractiveness. However, when incentivized for likability, many participants gave flattering advice by suggesting betting on higher ranks that imply a higher evaluation of attractiveness."
#| eval: false

library(ggplot2)
library(ggridges)

study1_time2 <- study1_time2 %>% dplyr::mutate(Advice = as.numeric(Advice))

ggplot(study1_time2, aes(x = Advice, y = Condition, fill = Condition)) +
    # Reduce vertical spacing with rel_min_height and scale
    geom_density_ridges(scale = 1.0, rel_min_height = 0.01, alpha = 0.7) +

    # Shading the left side of x = 7 in light grey
    geom_rect(aes(xmin = 7, xmax = Inf, ymin = -Inf, ymax = Inf), fill = "gray95", alpha = 0.01) +

    # Add the dashed line at x = 7
    geom_vline(xintercept = 7, col = 'black', linetype = 'dashed', linewidth = 0.7) +

    # Add the text "Flattering" on the left of the line
    annotate("text", x = 5, y = 0.6, label = "Flattering Ranks", hjust = 1, size = 8, fontface = "bold") +

    # Manual fill colors
    scale_fill_manual(values = c("#2A629A", '#FF7F3E'), labels = c("Accurate", "Likable")) +

    # X-axis adjustments
    scale_x_continuous(name = "Advised Rank", breaks = seq(1, 10)) +

    # Theme adjustments
    theme_bw() +
  coord_flip() +
    theme(panel.grid = element_blank(),
          panel.border = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = "none",
          legend.title = element_blank(),
          legend.key.size = unit(1.2, "lines"),
          legend.text = element_text(size = 12),
          plot.title = element_text(hjust = 0.5, size = 20),
          axis.text.x = element_text(size = 14),  # X-axis text size
          axis.text.y = element_text(size = 14),
          axis.title.x = element_text(size = 16, face = "bold"),  # X-axis title size
          axis.title.y = element_text(size = 16, face = "bold"))

```

```{r}
#| label: fig-study1advice
#| fig-cap: "Participants in Study 2 give advice to the person they rank 7th most attractive out of 10. When incentivized for accuracy, the majority recommends betting on that rank. However, when incentivized for likeability, participants inflate their advice to flatter the recipient as more attractive."

study1_time2 <- study1_time2 %>% dplyr::mutate(Advice = as.numeric(Advice))

study1_time2 %>%
  dplyr::group_by(Condition) %>%
  dplyr::mutate(total_per_condition = dplyr::n()) %>%
  dplyr::group_by(Condition, Advice, total_per_condition) %>%
  dplyr::summarize(n = dplyr::n()) %>%
  dplyr::mutate(pct = ifelse(Condition == "Accurate", -n/total_per_condition, n/total_per_condition),
                Advice = factor(Advice)) %>%
  ggplot(aes(x = pct, y = Advice, fill = Condition)) +
  geom_col() +
  scale_x_continuous("Share of Advice",
                     limits = c(-0.6, 0.6),
                     labels = \(x) scales::percent(abs(x))) +
  scale_y_discrete(limits = rev) +
  # Dashed 'horizontal' line at Advice = 7
  geom_hline(yintercept = 4, col = 'black', linetype = 'dashed', linewidth = 0.7) +
  # Example text annotation (adjust x/y for positioning)
  annotate("text", x = -0.3, y = 9, label = "Accuracy Incentive", hjust = 1, size = 4, fontface = "bold") +
  annotate("text", x = 0.6, y = 9, label = "Likeable Incentive", hjust = 1, size = 4, fontface = "bold") +
  # scale_y_continuous(name = "Advised Rank", breaks = seq(1, 10)) +
  scale_fill_manual(values = c("#2A629A", "#FF7F3E")) +
  papaja::theme_apa() +
  theme(
    legend.position  = "none",
    axis.text.x  = element_text(size = 14),
    axis.text.y  = element_text(size = 14),
    axis.title.x = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 16, face = "bold")
  )
```

We begin by examining the prior beliefs of advisees who uploaded their
selfies. On average, men guess they rank
`r round(mean(study1_time1[study1_time1$gender == "Male",]$Guess),2)` in
their group of 10 and women guess that they rank
`r round(mean(study1_time1[study1_time1$gender == "Female",]$Guess),2)`
(`r papaja::apa_print(t.test(Guess ~ gender, data = study1_time1, var.equal = T))$statistic`).
Notably, people's self-perceptions correlate strongly with the aggregate
ratings of the advisers
(`r round(cor(study1_finalRank$Guess, study1_finalRank$finalRank),2)`,
`r papaja::apa_print(cor.test(study1_finalRank$Guess, study1_finalRank$finalRank))$statistic`).
However, there is substantial heterogeneity in perceptions of
attractiveness. Of the 200 participants,
`r length(unique(study1_time2$Target))` are ranked as 7th most
attractive by at least one adviser. On average, men in this subset
estimate they rank
`r round(mean(TargetsUnique[TargetsUnique$gender == 'Male',]$Guess),2)`th
and women estimate they rank
`r round(mean(TargetsUnique[TargetsUnique$gender == 'Female',]$Guess),2)`th
(`r papaja::apa_print(t.test(Guess ~ gender, data = TargetsUnique, var.equal = T))$statistic`).

Next, we turn our attention to the advisers (see @fig-study1advice). In
the Accuracy condition, those uploading selfies are on average advised
to bet on rank
`r round(mean(study1_time2[study1_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study1_time2[study1_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offer flattering advice.[^3] Importantly, and as predicted, we find that
advisers in the Likeable treatment recommend betting on a lower rank,
communicating that they think the participant in the selfie is more
attractive
(`r round(mean(study1_time2[study1_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study1_time2, var.equal = T))$statistic`).
This suggests that advisers infer flattering someone with pleasant
advice would make the adviser appear more likeable. The distribution
shown in @fig-study1advice shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

[^3]: This shading could be due to concerns of avoiding disappointment.
    However, it could also be that advisers are uncertain about the
    rankings they have given and make a recommendation that combines
    their own belief with a uniform prior. Therefore, our analyses focus
    on the difference between the two conditions.

<!-- David: Making a note here that we should be sure to discuss this at the end -->

<!-- as something for future research. -->

<!-- David: We should also report something on accuracy of the advice given across -->

<!-- the treatments. Even if no difference. -->

```{r}
#| label: tbl-study1regs
#| tbl-cap: "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisers are rated as more trustworthy when they advice lower ranks, but this relationship is only directional."
#| after-caption-space: 0pt

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study1_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study1_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study1_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

```

Finally, we examine whether flattering advice indeed leads to more
positive evaluations of advisers, or whether flattering advice is
dismissed as insencere. Following our preregistration, we average the
ratings on likeability, warmth, friendliness, and good-naturedness to
create a scale of likeability ($\alpha$ =
`r round(cronbach.alpha(study1_time3[, 5:8])$alpha,2)`); and we create a
scale of trustworthiness by averaging the ratings of trustworthiness and
sincerity ($\alpha$ =
`r round(cronbach.alpha(study1_time3[, 9:10])$alpha,2)`).

```{r}
#| label: fig-study1rating
#| fig-cap: "Participants tend to rate advisers who suggest betting on a higher rank (implying greater attractiveness) as more likable and warm."

library(ggplot2)
library(dplyr)
library(tidyr)

study1_time3 <- study1_time3 %>% dplyr::mutate(Advice = as.numeric(Advice))

# Reshape the data to long format
study1_long <- study1_time3 %>%
  pivot_longer(cols = c(likable, warmAvg), 
               names_to = "Evaluation", 
               values_to = "Value")

# Create the plot
ggplot(study1_long, aes(x = Advice, y = Value, color = Evaluation)) +  
  # Add points for each observation
  geom_point(size = 2, alpha = 0.7,
             position = position_jitter()) +

  # Add linear regression lines for each metric
  geom_smooth(method = "lm", se = TRUE, aes(fill = Evaluation), alpha = 0.2, show.legend = FALSE) +

  # Customize axis labels
  scale_y_continuous(name = "Evaluation",limits = c(0.9, 5.1)) +  
  scale_x_continuous(name = "Advised Rank",limits = c(1, 10), breaks = c(1,2,3,4,5,6,7,8,9,10)) +

  # Apply a theme
  theme_bw() +  
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "right",
        plot.title = element_text(hjust = 0.5, size = 20)) +

  # Set custom colors for each metric
  scale_color_manual(values = c("#FF7F3E", "#2A629A"), 
                     labels = c("Likability", "Warm")) + 
  scale_fill_manual(values = c("#FF7F3E", "#2A629A"))

```

```{r}
#| label: fig-study1ratingByGroup
#| fig-cap: "After participants receive advice that suggests betting on a rank that was lower than, equal to, or higher than their own guess, they evaluate advisers based on their likability and warmth. Participants rate advisers who give flattering advice and deliver a high evaluation of their attractiveness as more likable and warm than those whose advice implied a lower evaluation of their attractiveness."

library(dplyr)
library(ggplot2)
library(reshape2)

study1_time3 <- study1_time3 %>% dplyr::mutate(Guess = as.numeric(Guess), Advice = as.numeric(Advice)) %>%
  dplyr::mutate(HorL = ifelse(Guess == Advice,0, ifelse(Guess > Advice, 1,-1))) %>%
  dplyr::mutate(HorL = as.factor(HorL)) %>% 
  dplyr::mutate(bonus = ifelse(bet == `Final Rank`,1,0)) %>%
  dplyr::mutate(accuracy = `Final Rank`-Advice)

ratings <- study1_time3 %>%
  group_by(HorL) %>%
  summarise(like = round(mean(likable),2), warmth = round(mean(warmAvg),2))

rating_table <- matrix(c(ratings$like,ratings$warmth), nrow = 2, byrow = TRUE, dimnames = list( c("Likable",'Warm'),c("Lower than", "Equal to", "Higher than")))

rating_df = as.data.frame.table(rating_table) %>%
  dplyr::rename( 'Condition'='Var1', 'Advice' ='Var2') %>%
  dplyr::mutate(Freq = round(Freq,2))

# Create a data frame similar to your structure
data <- rating_df

# Reshape the data to wide format for the heatmap
heatmap_data <- dcast(data, Condition ~ Advice, value.var = "Freq")

# Melt the data for ggplot
melted_data <- melt(heatmap_data, id.vars = "Condition", variable.name = "Advice", value.name = "Rating")

# Create the heatmap
ggplot(melted_data, aes(x = Advice, y = Condition, fill = Rating)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "darkblue", high = "orange", name = "Rating") +  # Grey gradient from light to dark
  geom_text(aes(label = sprintf("%.2f", Rating)), color = "white", size = 8) +
  theme_minimal() +
  labs(x = "Attractiveness (Advice Relative to Self-Expectations)", y = "Evaluation") +
  theme(
    axis.text.x = element_text(size = 12),  # X-axis text size
          axis.text.y = element_text(size = 12),
          axis.title.x = element_text(size = 12),  # X-axis title size
          axis.title.y = element_text(size = 12))

```

As seen in @fig-study1rating, we advisers who suggested that the advisee
was more attractive (lower rank) were indeed rated as more likeable and
warm (b = -0.109, p \< .05; b = -0.133, p \< .001, respectively; See
Columns 1 and 2 of @tbl-study1regs). When using their own initial guess
as a reference point, advisers who suggest a worse rank are perceived as
less likable and warm (see @fig-study1ratingByGroup). Interestingly,
these benefits are not at the cost of sincerity; advisers who recommend
a more favorable rank are viewed as no less trustworthy (Column 3 of
@tbl-study1regs).

We are not powered to do a comparison across the two experimental groups
and did not preregister such a difference. Indeed, we find no difference
in likeability and warmth across the two treatments
(`r papaja::apa_print(t.test(likable ~ Condition, data = study1_time3, var.equal = T))$statistic`,
and
`r papaja::apa_print(t.test(warmAvg ~ Condition, data = study1_time3, var.equal = T))$statistic`,
respectively). We also assess the quality of advice by measuring the
discrepancy between the advised rank and the correct rank. We find no
difference in accuracy across the advice treatments
(`r papaja::apa_print(t.test(accuracy ~ Condition, data = study1_time3, var.equal = T))$statistic`).

## Discussion

When advice communicates ego-relevant information (here, people's
attractiveness), we find that the advice people give is contingent on
their incentives. Specifically, when they get rewarded for being more
likeable, they recommend that the advisee bet on a more favorable rank
than when they are incentivized for accuracy. People understand that the
advice they give can influence how others perceive them and often use
flattering advice as a tool to boost their likeablity. Importantly,
advisees do not discount flattering advice and instead evaluate people
who advise them to bet on a more attractive rank as warmer and more
likeable. These gains to interpersonal perceptions do not come at the
cost of trustworthiness, even when the advice is inflated.

# Study 2

We next examine whether advisers take into account the expectations of
advisees when providing a recommendation. A group of participants
("Advisees") complete a quiz consisting of ten questions that draw on
ego-relevant domains. They then express a non-binding preference to
compete against a group of high performers or low performers on the same
task, based on their past performance. We nudge participants towards
picking either the low or the high performer group by selecting one of
the two by default. A second group of participants ("Advisers") observe
Advisees' score on the quiz and, in one treatment, also which group they
had selected. We predicted that Advisees who were nudged toward the high
performer group would be more likely to receive advice to compete
against this high group when Advisers observed their non-binding choice.

## Methods

```{r}

study2_time1 <- readRDS(here::here("Data/Study2_time1.Rds"))
study2_ability50 <- read.csv(here::here("Data/Cleaned Data/S2_Ability50_clean.csv"))
study2_time2_giver <- readRDS(here::here("Data/Study2_time2_giver.Rds"))
study2_time2_advice <- readRDS(here::here("Data/Study2_time2_advice.Rds"))
study2_time3 <- readRDS(here::here("Data/Study2_time3.Rds"))

```

In a preliminary stage, we recruited
`r scales::comma(dim(study2_ability50)[1])` participants from Prolific
and gave them five minutes to complete a quiz consisting of ten items.
The quiz included word puzzles, identifying emotions from photos, and
selecting the best responses for hypothetical scenarios. These questions
were adapted from surveys that measure problem-solving, emotional
intelligence, and communication skills---which we pre-tested as being
important in modern society and hence where participants might have a
stake in doing well. Participants received a bonus of five cents for
each correctly answered question. We then ranked them based on their
score and label the 20 participants with the lowest (highest) scores as
the Low (High) Performer Group.

Next, we recruited `r scales::comma(dim(study2_time1)[1])` participants
for the role of Advisees. They completed the same 10-item quiz, also
earning five cents for each correct answer. We then asked them to
express a non-binding preference for whether they would like to compete
against the High or Low Performer Group. We informed them that they
would be invited back at a later date when they could make a binding
decision and they could earn a bonus based on whether their score on
this quiz was equal to or higher than a randomly selected member from
their comparison group. If they picked the High Performer Group and had
an equal or higher score than a randomly selected member of that group,
they would earn a bonus of 50 cents. If they picked the Low Performer
Group and outperformed a randomly selected member of that group, they
could earn a bonus of 20 cents. If their score was lower in this
comparison, they would not earn an additional bonus. We randomized which
of the two groups (High or Low Performers) was selected by default, and
Advisees were free to select the other group. The survey concluded with
basic demographic questions.

We then recruited `r scales::comma(dim(study2_time2_giver)[1])`
participants for the role of Advisers, and the focal part of our
experiment. They were informed of the ability quiz that the Advisees had
completed, how the Low Performer Group and High Performer Groups were
constructed, as well as the choice and incentives for the Advisees. We
then asked Advisers to give advice to ten participants on which group
they should compete against. Advisers were randomly assigned to one of
two treatments. In the "Performance" treatment, they observed only the
score of the other participant. In the "Performance + Expectation"
treatment, they additionally observed the advisee's non-binding choice
of which group to compete against. However, since the outcome was based
solely on the past quiz score and advisers, but not advisees, know the
score, this initial choice was not informative for the purpose of
recommending competing against one group or the other. Advisers could
earn the same bonus as one of the participants they had given advice to
and who returned to make a decision. The survey concluded with basic
demographic questions.

Finally, we invited Advisees back for the follow-up survey. Following
our preregistration, we kept the survey open for 7 days. In total,
`r scales::comma(dim(study2_time3)[1])` Advisees returned. The short
survey reminded them of the task they completed in the previous survey,
informed them that other participants from Prolific had observed their
real score and given them advice on which group to compete against, and,
finally, reminded them of their initial choice of comparison group. They
were not informed of their true score or the score of the groups they
could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

## Results

```{r}
contingency_table <- table(study2_time1$Guess, study2_time1$default)

chi_test <- chisq.test(contingency_table)

```

We begin by examining the performance of the Stage 1 participants.
Because our default treatment takes place after participants completed
the ability quiz, we do not expect a difference in performance across
the Low and High Default treatments. Indeed, the two groups score no
different from one another
(`r papaja::printnum(round(mean(study2_time1[study2_time1$default == "Low",]$Score),2))`
and
`r round(mean(study2_time1[study2_time1$default == "High",]$Score),2)`
for the Low Default and High Default treatments, respectively,
`r papaja::apa_print(t.test(Score ~ default, data = study2_time1, var.equal = T))$statistic`).
As expected, the default manipulation changes their initial choices. In
the "Low Default" treatment, only
`r round(mean(study2_time1[study2_time1$default == "Low",]$Guess),2)*100`%
of participants prefer to compete against the High Performer Group,
compared to
`r round(mean(study2_time1[study2_time1$default == "High",]$Guess),2)*100`%
of participants in the "High Default" treatment ($\chi^2$(1, n = 201) =
36.62, p \< .001).

\linespread{1}

```{r}
#| label: tbl-study2regs
#| tbl-cap: "Advice to compete against the High Performer Group in Study 1. Displaying the non-binding choice of those for whom the Low Performer Group was selected by default makes it less likely that they are advised to compete against the High Performer Group (Column 1). Column 2 restricts the analysis to advisers who observe advisees' initial non-binding choice and controls for advisees' score on the quiz. Standard errors are clustered at the adviser level."
#| after-caption-space: 0pt

study2_advice <- study2_time2_advice %>%
  dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E', 1, 0))

modelsummary::modelsummary(
  list(`All Advice` = lm(Advice ~ ExpShow * Treatment, data = study2_advice), 
       `Expectations Shown` = lm(Advice ~ Treatment + Score, data = study2_advice %>% dplyr::filter(ExpShow == 1))),
  vcov = ~GiverID,
  coef_map = c(
    "TreatmentHigh" = "High Default",
    "ExpShow" = "Expectation Shown",
    "ExpShow:TreatmentHigh" = "Expectation x High Default",
    "Score" = "Score",
    "(Intercept)" = "Constant"
  ),
  stars = TRUE,
  gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0))
) %>%
  tinytable::group_tt(
   j = list("Advice to Compete Against High Performers" = 2:3))

```

We now examine whether advisers take the advisees' expectations into
account. Advisers observe the true performance of the advisee, such that
expectations about performance add no additional information. However,
because participants who score higher are also more likely to pick the
High Performer Group, we use our Default treatment as an instrument that
reflect higher expectations. We show linear probability models for the
decision of the Advisers, which is equal to 1 if they recommend
competing against the High Performer Group and 0 if they recommend the
Low Performer Group. In Column 1 of @tbl-study2regs, we show our main
specificiation with the experimental assignment for Advisees (High
Performer Default vs. Low Performer Default) and Advisers (Expectation
Shown vs. Expectation Hidden), as well as their interaction. Because
each adviser made ten recommendations, we cluster standard errors at the
adviser level. We find that displaying expectations of those in the Low
Performer Default reduces recommendations to compete against the Top
group by 4 percentage points, or about 10% relative to when expectations
are hidden. This decrease is marginally smaller for advisees with a
default choice to compete against the High Performer Group (nudged to
have high expectations, though not necessarily holding them), and whose
advice did not differ based on whether their expectations were shown or
hidden. We reported a linear probability model on advising to compete
against the High Performer group for advisers in the "Expectations"
treatment, controlling for the true performance of the advisee in Column
2 of @tbl-study2regs. We observe a significant main effect of the
default treatment, where a high default increased the likelihood of
receiving flattering advice. It is worth noting that, since only
`r round(mean(study2_time1[study2_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their
default choice, this represents a conservative test of the effect of
high expectations on the likelihood of receiving flattering advice.

```{r}
#| label: fig-study2Advice
#| fig-cap: "Advice given in Study 2. Advisees who were nudged to express high expectations were more likely to receive advice to compete against the High Performer Group."

study2_time2_advice %>%
  mutate(Treatment=factor(Treatment, levels=c('High','Low'), labels = c('High default', 'Low default'))) %>%
  group_by(Condition, Treatment) %>%
  summarize(avg_count = mean(Advice), ci = qnorm(0.975)*sd(Advice)/sqrt(dplyr::n())) %>%
  ggplot2::ggplot(ggplot2::aes(x = Condition, y = avg_count, fill = Treatment)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = avg_count - ci, ymax = avg_count + ci), 
                         position = ggplot2::position_dodge(0.9), linewidth = 0.2, width = 0.1) +
  # ggtitle("Advice: Meeting Expectations vs. Absent Expectations") +
  scale_fill_manual(values = c("#2A629A",'#FF7F3E')) +
  scale_y_continuous("P(Compete Against High Performer)",
                     labels = scales::percent) +
  geom_text(aes(x = Condition, y = 0.02,
                label = paste0(round(avg_count*100), "%")),
            position = position_dodge(0.9),
            color = "white") +
  papaja::theme_apa() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        axis.text.x = element_text(size = 12, face = "bold",color = "black"),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none", # No legend
        plot.title = element_text(hjust = 0.5, size = 16)) +
  scale_x_discrete(labels = c("Expectation Shown", "Expectation Hidden")) +
   geom_text(aes(label=Treatment, y = -0.02), position = position_dodge(0.90), color = "#464747"  )

```

## Discussion

When an adviser knows an outcome that an advisee can only estimate,
learning this estimate should not change the advice that is given.
However, in line with our argument that advisers take into account the
belief utility of the advisee and prefer to avoid disappointing them, we
find that providing information about expectations makes a difference.
Specifically, advisees who expressed a preference for stronger
competitors were more likely to receive flattering advice, encouraging
them to pursue that option. This occurs even though advisers share the
potential costs of recommending a more challenging competition.

```{=html}
<!-- Amanda: I try to avoid using the term "good advice" or bring in risk.
-->
```

# Study 3

```{r}
#| include: false

study3_math50 <- read.csv(here::here("Data/Cleaned Data/S3_Math50_cleaned.csv"))
study3_time1All <- readRDS(here::here("Data/Study3_time1All.Rds"))
study3_time1 <- readRDS(here::here("Data/Study3_time1.Rds"))
study3_time2_giver <- readRDS(here::here("Data/Study3_time2_giver.Rds"))
study3_time2_giver50 <- readRDS(here::here("Data/Study3_time2_giver50.Rds"))
study3_advice <- readRDS(here::here("Data/Study3_time2_advice.Rds"))
study3_time3 <- readRDS(here::here("Data/Study3_time3.Rds"))

```

In Study 3, we examine how advisers who want to avoid disappointment may
inadvertently contribute to gender differences in advice. We ask
participants to estimate their score on a mathematics test, and
anticipate that men will guess higher than women. Moreover, we expect
that advisers take into account the expected score when recommending
whether someone should compete against a group of High or Low
performers. As a result, we hypothesize that, when expectations are
known to advisers, men will be advised to compete against High
Performers more than women, even when advisers do not know gender and
expectations are uninformative.

## Methods

```{r}
#| include: false

subgroup_averages <- data.frame(subgroup = numeric(), avg_score = numeric())
set.seed(807)
# loop through 1000 iterations of randomly selecting 10 participants and computing their average score
for (i in 1:1000) {
  # randomly select 10 participants
  subgroup <- sample(study3_math50$ProlificID, size = 5, replace = FALSE)
  # compute the average score for the selected participants
  subgroup_avg <- mean(study3_math50$Score[study3_math50$ProlificID %in% subgroup])
  # add the subgroup and its average score to the dataframe
  subgroup_averages <- rbind(subgroup_averages, data.frame(subgroup = i, avg_score = subgroup_avg))
}

p5 <- quantile(subgroup_averages$avg_score, probs = 0.05)
p95 <- quantile(subgroup_averages$avg_score, probs = 0.95)

sortedScore <- sort(study3_math50$Score, decreasing = TRUE)
LP_lower <- sortedScore[50]
LP_upper <- sortedScore[31]
HP_lower <- sortedScore[20]
HP_upper <- sortedScore[1]

```

We begin by first recruiting a sample of
`r scales::comma(dim(study3_math50)[1])` participants from Prolific to
complete a 10-question multiple choice mathematics quiz. The questions
are taken from a paper-version of the ASVAB standardized exam, such that
answers are not available online. Participants have five minutes to
answer the quiz and are paid 10 cents for each correctly answered
question. Like in Study 1, we define the top 20 scores as "High
Performers" and the bottom 20 scorers as the "Low Performers." On
average, participants answered `r round(mean(study3_math50$Score),2)`
questions correctly, High Performers scored between `r HP_lower` and
`r HP_upper`, and Low Performers scored between `r LP_lower` and
`r LP_upper`. In order to anchor the expectations of participants in our
main experiment, we simulated 1,000 pairings of groups of five
participants, with the 5<sup>th</sup> percentile of groups scoring an
average of `r p5` and the 95 th percentile scoring an average of
`r p95`. We report these averages to participants in the Low
Expectations and High Expectations treatment, respectively.

We then recruit `r scales::comma(dim(study3_time1All)[1])` participants
for the role of advisees in Stage 1 of our main experiment. To arrive at
a gender-balanced sample, we drop the last two male participants to
complete the survey, ending up with a sample of 500 men and 500 women
($M_{\text{Age}}$ = `r mean(study3_time1$Age)`). Participants complete
the same 10-item mathematics quiz as the earlier participants and are
informed that their performance would affect their bonus earnings in a
follow-up stage to be conducted a few days later. After completing the
quiz, we inform them of the average score of a group of five
participants from the preliminary survey. We randomly assign them to
learn about the 5<sup>th</sup> percentile of groups, which scored
`r p5`. ("Low Expectations" treatment) or the 95<sup>th</sup> percentile
of groups, which scored `r p95` ("High Expectations" treatment).
Participants then make an unincentivized guess about how many questions
they think they answered correctly. The survey concludes with basic
demographic questions.

We then recruit `r scales::comma(dim(study3_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We begin
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 have completed, and inform them of the
average score of all participants in the preliminary study. Advisers
have to recommend whether an advisee should compete against the Low
Performers or the High Performers (we use these terms in the survey). We
anticipate that being told to compete against High Performers is more
flattering and hence being told to compete against the Low Performers
would be disappointing if one had expected to do well. Advisees would
earn a bonus if their score is equal to or higher than that of a
randomly selected member from their chosen group. They earn 50 cents if
they outperform a High Performer and 30 cents if they outperform a Low
Performer.

Advisers are randomly assigned to one of two treatments. In the
"Baseline" treatment, they only observe the score of the advisee on the
mathematics quiz. In the "Expectation" treatment, they observe the score
as well as the advisee's guess for how many questions they answered
correctly. Since the outcome is determined only by the past score on the
quiz, the advisee's guess is not informative for the recommendation,
which can be based on the revealed score. Moreover, in neither treatment
did they receive any demographic information about the advisees.
Participants give recommendations to 10 advisees, which unbeknownst to
them are five men and five women matched to have identical performance
on the test.[^4] They are informed that if their advice is shown to a
participant who returns for the follow-up survey, they will receive the
identical bonus as that participant. The survey then concludes with
basic demographic questions.

[^4]: We made this decision to account for the possibility of gender
    differences in performance.

Finally, we invite participants from Stage 1 back for the follow-up
survey. Following our preregistration, we keep the survey open for seven
days. In total, `r scales::comma(dim(study3_time3)[1])` participants
(`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Female',])[1])`
women) return. The brief survey reminds them of the task they completed
in Stage 1, informs them that other participants from Prolific have
observed their real score and given them advice on which group to
compete against, and finally are reminded of how many questions they
guessed they had answered correctly. Importantly, they are not informed
of their true score or the score of the groups they could compete
against. Participants then observe the advice from a randomly selected
adviser and make their decision.

## Results

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants complete the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups score
no different from one another
(`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study3_time1[study3_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study3_time1, var.equal = T))$statistic`).
As intended, the expectations treatment does affect how well they
thought they performed. Participants in the "Low Expectations" treatment
guess a score of
`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study3_time1[study3_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study3_time1, var.equal = T))$statistic`),
showing that the manipulation is successful, albeit small. Contrary to
our expectations, we find a gender difference in performance: men score
`r round(mean(study3_time1[study3_time1$Gender == "Male",]$count),2)` on
average, while women score
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study3_time1, var.equal = T))$statistic`).
Consistent with this difference, men think they answered more questions
correctly than do women
(`r round(mean(study3_time1[study3_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study3_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We
observe that women underestimate their performance by 0.76 points, while
men do so by only 0.27 points
(`r papaja::apa_print(t.test(overconfidence ~ Gender, data = study3_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`).
Although neither gender is overconfident, men on average are more
confident in their performance than are women, as we had expected.

```{r}
#| label: tbl-study3regs
#| tbl-cap: "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the adviser. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the adviser. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the adviser. All standard errors clustered at the level of the adviser."
#| after-caption-space: 0pt

study3_advice <- study3_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study3_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study3_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study3_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

```

```{r}
#| label: fig-study3genderdiff
#| fig-cap: "Advisers observed real performance and the participants' estimated performance, but not their gender. As a result of their higher expectations, men (blue) were advised to compete against the high performance group more often than women (red). The grey line shows advice absent expectations, which did not differ by gender."

library(dplyr)
library(ggplot2)

# Calculate averages for the performance, grouped by condition and gender
ProbHG <- study3_advice %>%  
  group_by(Condition, Performance, Gender) %>%  
  summarise(avg_hardGroup = mean(Advice), .groups = 'drop')

# Create a data frame for the NONE condition as the baseline
baseline <- ProbHG %>%  
  filter(Condition == "None") %>%  
  select(Performance, baseline_avg = avg_hardGroup)

# Merge with the original data to get comparison values
comparison_data <- ProbHG %>%  
  filter(Condition != "None") %>%  
  left_join(baseline, by = "Performance") %>%  
  mutate(diff = avg_hardGroup - baseline_avg)  # Calculate the difference

# Create a scatter plot comparing conditions with NONE as baseline
ggplot() +
  # Step lines for the NONE condition
  geom_segment(data = baseline,
               aes(x = Performance, 
                   xend = lead(Performance, 1),  # Create horizontal lines to the next performance point
                   y = baseline_avg, 
                   yend = baseline_avg,
                   linetype = "None"), 
               color = "grey70", size = 0.5, 
               na.rm = TRUE,
               show.legend = TRUE) + 
  geom_point(data = comparison_data %>% filter(Gender == "Male"), 
             aes(x = Performance + 0.5, y = avg_hardGroup, shape = Gender, color = Gender), 
             size = 4) +  # Males with an offset to the left
  geom_point(data = comparison_data %>% filter(Gender == "Female"), 
             aes(x = Performance + 0.5, y = avg_hardGroup, shape = Gender, color = Gender), 
             size = 4) +  # Females with an offset to the right
  scale_shape_manual(values = c(16, 17), labels = c("Female", "Male")) +  # Circle for Females, Triangle for Males
  scale_color_manual(values = c( "#FF7F3E","#2A629A"), 
                     labels = c("Female", "Male")) +  
  scale_y_continuous(name = "P(Compete Against High Performance)",
                     labels = scales::percent,
                     limits = c(0, 1.1),
                     breaks = c(0,1)) +  # Set limits for y-axis
  scale_x_continuous(name = "Performance",
                     limits = c(0, 11),expand = c(0, 0), 
                     breaks = seq(0, 10, 1)) +  
  theme_bw() +  
  theme(panel.grid = element_blank(), 
        panel.border = element_blank(), 
        axis.line = element_line(colour = "grey"), 
        legend.position = "bottom", 
        legend.title = element_blank(), 
        legend.key.size = unit(1.2, "lines"), 
        legend.text = element_text(size = 12), 
        plot.title = element_text(hjust = 0.5, size = 20))

```

Next, we examine whether advisers take the advisee's expectations into
account. Column 1 of @tbl-study3regs reports a linear probability model
on advising to compete against the High Performer group for advisers in
the "Expectations" treatment, controlling for the true performance of
the advisee. Because each adviser makes ten recommendations, we cluster
standard errors at the adviser level. However, contrary to our
expectations, we do not see a significant effect of the expectations
treatment. This may be because the induced difference in expectations
was too small.

However, recall that men are more confident in their performance than
are women. Our theory thus predicts that showing expectations should
lead to more flattering advice (i.e., a recommendation to compete
against the High Performer group) for men than for women. We report a
linear probability model with advice to compete against the high group
as the outcome measure, and the advisee's gender, whether expectations
were shown to the adviser, and the interaction of the two in Column 2 of
@tbl-study3regs. As predicted, we find a significant interaction effect:
men are more likely to be advised to compete against the High Performers
when expectations are shown than when they are hidden ($p < 0.001$). We
show this result graphically in @fig-study3genderdiff. As the figure
makes clear, this effect is driven by advice given to those who scored
in the middle of the possible range. Thus, when it is clear that someone
should compete against Low or High performers, advisers are not
deferring to expectations.

To examine the quality of advice, we computed the expected bonus
earnings for someone who follows the recommendations. For example, if an
adviser suggests competing against the High Performer group, we match
the advisee against all 20 members of that group and determined how
often their score matches or exceeds that of the member. We then
multiply this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of
@tbl-study3regs. Displaying expectations leads men to be advised to
compete against the High Performer group more often, and this advice
turns out to be costly: men receive worse advise than do women when
expectations are displayed, but not in their absence.[^5]

[^5]: This analysis was not preregistered, and we note here that the
    reduction in expected earnings is small. However, it is interesting
    that expectations have a negative effect for men who underestimate
    their performance on average. One possibility is that advisers
    suggest the High Performer group more often than is optimal. This is
    consistent with our finding from Study 2, in which participants
    offered flattering advice even when incentivized only for accuracy.

```{r}
#| label: tbl-study3time3
#| tbl-cap: "Column 1 displays the actual bonus based on the performance and whether the participant is primed with high expectations. Column 2 displays the actual bonus based on the performance and gender of the participant. Columns 3 and 4 display the chance of adopting the advice based on the gender of the participants, whether the advisee sees expectations, and whether the advice is to compete against high performers. The former considers only the main effect, while the latter also includes the interactive effect."
#| after-caption-space: 0pt

study3_time3 <- study3_time3 %>% dplyr::mutate(AdviseeCondition = factor(AdviseeCondition, levels = c('Low', 'High'))) %>%
     dplyr::mutate(Adopt = ifelse(Bet == Advice,1,0))

modelsummary::modelsummary(list(`(1)` = lm(bonus ~ AdviseeCondition + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(2)` = lm(bonus ~ Gender + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(3)` = lm(Adopt ~ Gender + GiverCondition + Advice, data = study3_time3),
                                `(4)` = lm(Adopt ~ Gender + GiverCondition + Advice + Gender*GiverCondition + GiverCondition*Advice + Gender*Advice, data = study3_time3)),
                           vcov = c("classical", "classical","classical","classical"),
                            coef_map = c("AdviseeScore" = "Performance",
                                         "AdviseeConditionHigh" = "High Expectation",
                                        "GenderMale" = "Advisee Male",
                                        "GenderMale:GiverCondition" = "Expectation x Male",
                                        "GiverCondition" = "Expectation Shown",
                                        "Advice1" = "Advice: High Performer",
                                        "GenderMale:Advice1" = "Advice: High Performer x Male",
                                        "GiverCondition:Advice1" = "Expectation x Advice: High Performer",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

```

To determine whether flattering advice is truly costly, however, we need
to examine the outcome of the advisees. In particular, they could ignore
flattering advice, recognizing it as such and thus failing to adhere to
it. In line with our prediction, participants who are primed with high
expectations earn less when their expectations are shown to the
advisers, although this result is only directional (see Column 1 of
@tbl-study3time3). Similarly, as shown in Column 2 of @tbl-study3time3,
we also find that male participants earn less when their expectations
are shown. These findings suggest that flattering advice is not without
consequences. Furthermore, our research indicates that because men are
more likely to follow such advice (see Column 4 of @tbl-study3time3),
they compete more as a result of flattering advice.

## Discussion

Our findings suggest that attempts to avoid disappointment can be a
novel source of gender differences in advice. Women underestimate their
mathematics test scores more than men. When these expectations are shown
to advisers, they are more likely to tell men to compete against high
performers than women. Notably, this turns out to be poor advice: men
whose advisers are aware of their expectations receive worse advice. Our
findings suggest that men with the same scores as their female
counterparts end up earning less, although this result is only
directional. This discrepancy may be due to men receiving more favorable
advice, being more likely to follow it, and ultimately facing worse
outcomes.

# General Discussion

Advice has the potential to shape people's career and personal outcomes.
Honest advice, however, may be painful to learn if it falls short of
one's expectations. As prior work notes, this may motivate people to
avoid information and avoid seeking help [@Benabou2022; @Golman2017;
@Jaroszewicz2022]. We present evidence from three experiments that
advisers are also cognizant of this cost. As a result, they present
flattering advice that avoids disappointing the recipient, and correctly
anticipate that this boosts how advisees perceive them. However, this
flattering advice comes at a cost to advisees, who would do worse if
they followed it blindly.

Moreover, a desire to avoid disappointment also means that advisers have
to take into account the expectations of the advisee. We document that
men are more optimistic about their performance than women are. As a
result, they receive more flattering advice and are more likely to be
told to aim higher. Notably, in the context of our experiment, this
turns out to lead to lower bonus compared to more conservative advice.
While we focus on gender groups here, our findings may extend to other
social groups where differences in expectations exist. This dynamic may
create a reinforcing cycle — individuals with higher confidence are more
likely to receive encouragement and positive reinforcement, further
boosting their confidence. In contrast, individuals with lower
confidence have fewer opportunities to receive such encouragement,
despite potentially needing it the most.

<!--# I tried to replace "bad advice". -->

Our findings have implications for organizational practice, where
mentoring and advice-giving often consider employees' expectations. On
one hand, offering flattering advice may enhance relationships between
mentors and mentees, enhancing trust and communication, which are
crucial for collaboration and long-term organizational success. However,
flattering advice may also lead to talent misallocation, as employees
are more likely to be encouraged to pursue career paths aligned with
their stated preferences and less likely to be recommended to explore
alternative directions that might better match their potential.
Additionally, when advisers hold significantly more power than advisees,
interpersonal concerns may diminish [@Schaerer2018], resulting in advice
that is more blunt or even harsh. As a result, advisees may seek
flattering advice from peers or subordinates instead, missing out on
valuable expertise from their supervisors. Lastly, we identify
flattering advice as a novel source of gender bias, as women may receive
less ambitious advice due to lower expressed confidence. Organizations
could reduce this bias by calibrating employees' expectations to curb
overconfidence and/or compensate for a lack of confidence.

Participants in our experiment were paired online and remained
anonymous. Even so, we document this desire to avoid disappointment.
Given that communications mediated by computer tends to be less
emotionally rich [@Kiesler1985], and that anonymity reduces the
psychological burden of responsibility for the advisee's feelings, as it
cannot be traced back to the adviser. We anticipate that advice would be
more flattering in face-to-face communication and without anonymity.

Moreover, existing relationships might make it even harder for advisers
to be fully honest. Advisers tend to feel more concerned and emotionally
attached to how friends, family, or colleagues feel compared to how a
stranger would feel, which may increase their motivation to provide
flattering advice. Conversetly, advisees also tend to value the
emotional support embedded in advice from their friends, even when it
lacks informational utility. Moreover, people generally prefer their
close ones to demonstrate benevolence toward them rather than apply
impartial integrity during interactions [@Levine2020]. In response to
these expectations, advisers might intentionally inflate their advice to
strengthen their bond with the advisee and protect them from discomfort.

The closeness of the relationship may moderate the tendency to provide
flattering advice. On one hand, advisers may prioritize the advisee’s
best interests above other considerations when offering advice to
someone they have a close and strong relationship with, making them more
willing to provide the most helpful advice, even if it is unflattering.
Advisers may also feel more confident that a close relationship is
resilient enough to withstand any immediate disappointment the advisee
may experience if the advice falls short of their self-expectations. On
the other hand, closer relationships may also strengthed the adviser’s
motivation to protect the advisee’s feelings, increasing the likelihood
of offering flattering advice. Future research could explore how
relationship closeness influences the balance between providing
flattering versus candid advice.

In our experiments, participants only received advice once and did not
evaluate the adviser after observing the outcome of their decision. For
example, advice that leads to bad outcomes may undermine the
interpersonal benefits of flattery. Alternatively, people may still like
the flattering advice and not fault the adviser for the bad outcome.
Moreover, our setting involved only a single piece of advice on one
task. Future research could examine whether people return to those who
gave them flattering advice, or if they prefer someone who gave them the
honest (but unpleasant) truth.

Advice has long been studied from the perspective of the receiver
[@Rader2017; @yaniv2004receiving; @bonaccio2006Review; @Milyavsky2024].
Similarly, research on belief utility has examined how recipients
respond to the valence of the information they receive. Here, we show
that advisers, too, take into account the psychological impact of the
information they convey. They may have even greater motivations to avoid
conveying bad news, because they incur the interpersonal costs of
delivering unfavorable information without reaping the benefits from
helping someone make a better choice.

# References

::: {#refs}
:::